---
title: "Psilocybin Project 1"
subtitle: "Differential Expression"
author: "Kennedi Todd"
date: "10/14/2024"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

# Setup
## Working directory
```{r setup}
knitr::opts_knit$set(root.dir = ".")
```

## Libraries
```{r libraries, message=FALSE, warning=FALSE}
library(edgeR)      # to load .rds obj
library(ggrepel)    # geom_text_repel()
library(limma)      # to load .rds obj
library(pheatmap)   # pheatmap()
library(tidyverse)  # ggplot(), %>%, pivot_wider()
```

## User defined variables
```{r set_variables}
# Output directories and file paths
out <- "../results/sex_specific_filtering_but_both_sexes_comparison/"
out2 <- "filtered_samples_sex_specific"
model <- "DEGs_with_hbb/"

# Read metadata
meta <- readRDS("../rObjects/meta.rds")

# Colors and levels
treatment_order <- levels(meta$treatment)
treatment_colors <- c("lightblue", "blue")

sex_order <- levels(meta$sex)
sex_colors <- c("green", "green4")

age_order <- levels(meta$age)
age_colors <- c("pink", "purple")

timepoint_order <- levels(meta$timepoint_days)
timepoint_colors <- c("gold","red3")

group_order <- levels(meta$group)
group_colors <- c("#f55d3f","#f5ae3f","#EFF964","#59C751","#60ABF6","#ffaee0",
                  "#9D54DE","#bbbbbb")
remove(meta)
```

## Load data
```{r read_DGEList}
dge.filtered.norm <- readRDS(paste0("../rObjects/",
                                    out2,
                                    "_DGEList_filtered_normalized.rds"))
dim(dge.filtered.norm)
head(colnames(dge.filtered.norm))
```

# Voom
## Design matrix
```{r design_matrix}
design <- ""

if (out2 == "filtered_samples_sex_specific" && out == "../results/sex_specific_filtering_but_both_sexes_comparison/") {
  print("both sexes analysis with hbb")
  group <- interaction(dge.filtered.norm$samples$group)
  sex <- factor(dge.filtered.norm$samples$sex)
  hbb <- as.numeric(dge.filtered.norm$samples$hbb_log2cpm)
  design <- model.matrix(~ 0 + group + sex + hbb)
  design.names <- gsub("group","", colnames(design))
  colnames(design) <- design.names
} else if (grepl("sex_specific", out2) && (model == "DEGs_with_hbb/")) {
  print("sex specific analysis with hbb")
  group2 <- interaction(dge.filtered.norm$samples$group2)
  hbb <- as.numeric(dge.filtered.norm$samples$hbb_log2cpm)
  design <- model.matrix(~ 0 + group2 + hbb)
  design.names <- gsub("group2","", colnames(design))
  colnames(design) <- design.names
} else if (grepl("sex_specific", out2) && (model == "DEGs/")) {
  print("sex specific analysis without hbb")
  group2 <- factor(dge.filtered.norm$samples$group2)
  design <- model.matrix(~ 0 + group2)
  design.names <- gsub("group2","", colnames(design))
  colnames(design) <- design.names  
} else if (grepl("both_sexes", out2) && (model == "DEGs/")) {
  print("both sexes analysis without hbb")
  group <- interaction(dge.filtered.norm$samples$group)
  sex <- factor(dge.filtered.norm$samples$sex)
  design <- model.matrix(~ 0 + group + sex)
  design.names <- gsub("group","", colnames(design))
  colnames(design) <- design.names
} else if (grepl("both_sexes", out2) && (model == "DEGs_with_hbb/")) {
  print("both sexes analysis with hbb")
  group <- interaction(dge.filtered.norm$samples$group)
  sex <- factor(dge.filtered.norm$samples$sex)
  hbb <- as.numeric(dge.filtered.norm$samples$hbb_log2cpm)
  design <- model.matrix(~ 0 + group + sex + hbb)
  design.names <- gsub("group","", colnames(design))
  colnames(design) <- design.names
}

head(design)
```

## Contrasts
```{r contrasts}
myContrasts <- ""

if (out2 == "filtered_samples_sex_specific" && out == "../results/sex_specific_filtering_but_both_sexes_comparison/") {
    myContrasts <- c(
    # E - S comparisons
    "E.2D.Y - S.2D.Y", "E.21D.Y - S.21D.Y",
    "E.2D.O - S.2D.O", "E.21D.O - S.21D.O",
  
    # 21D - 2D comparisons
    "E.21D.Y - E.2D.Y", "S.21D.Y - S.2D.Y",
    "E.21D.O - E.2D.O", "S.21D.O - S.2D.O",
  
    # O - Y comparisons
    "E.2D.O - E.2D.Y", "S.2D.O - S.2D.Y",
    "E.21D.O - E.21D.Y", "S.21D.O - S.21D.Y")
} else if (out2 == "filtered_samples_sex_specific" | out2 == "all_samples_sex_specific") {
  myContrasts <- c(
    # E - S comparisons
    "E.2D.Y.F - S.2D.Y.F", "E.2D.Y.M - S.2D.Y.M",
    "E.2D.O.F - S.2D.O.F", "E.2D.O.M - S.2D.O.M",
    "E.21D.Y.F - S.21D.Y.F", "E.21D.Y.M - S.21D.Y.M",
    "E.21D.O.F - S.21D.O.F", "E.21D.O.M - S.21D.O.M",
  
    # 21D - 2D comparisons
    "E.21D.Y.F - E.2D.Y.F", "E.21D.Y.M - E.2D.Y.M",
    "E.21D.O.F - E.2D.O.F", "E.21D.O.M - E.2D.O.M",
    "S.21D.Y.F - S.2D.Y.F", "S.21D.Y.M - S.2D.Y.M",
    "S.21D.O.F - S.2D.O.F", "S.21D.O.M - S.2D.O.M",
  
    # O - Y comparisons
    "E.2D.O.F - E.2D.Y.F", "E.2D.O.M - E.2D.Y.M",
    "S.2D.O.F - S.2D.Y.F", "S.2D.O.M - S.2D.Y.M",
    "E.21D.O.F - E.21D.Y.F", "E.21D.O.M - E.21D.Y.M",
    "S.21D.O.F - S.21D.Y.F", "S.21D.O.M - S.21D.Y.M",
  
    # F - M comparisons
    "E.2D.Y.F - E.2D.Y.M", "E.2D.O.F - E.2D.O.M",
    "S.2D.Y.F - S.2D.Y.M", "S.2D.O.F - S.2D.O.M",
    "E.21D.Y.F - E.21D.Y.M", "E.21D.O.F - E.21D.O.M",
    "S.21D.Y.F - S.21D.Y.M", "S.21D.O.F - S.21D.O.M"
  )
} else if (out2 == "filtered_samples_both_sexes" | out2 == "all_samples_both_sexes") {
  myContrasts <- c(
    # E - S comparisons
    "E.2D.Y - S.2D.Y", "E.21D.Y - S.21D.Y",
    "E.2D.O - S.2D.O", "E.21D.O - S.21D.O",
  
    # 21D - 2D comparisons
    "E.21D.Y - E.2D.Y", "S.21D.Y - S.2D.Y",
    "E.21D.O - E.2D.O", "S.21D.O - S.2D.O",
  
    # O - Y comparisons
    "E.2D.O - E.2D.Y", "S.2D.O - S.2D.Y",
    "E.21D.O - E.21D.Y", "S.21D.O - S.21D.Y")
}

head(myContrasts)
```

## Quality weights
```{r voom}
# save
path <- paste0(out, "postfiltering_QC/voom_gene_mean_var_weights.pdf")
pdf(path, width = 6, height = 4)

# voom transform counts
v <- voomWithQualityWeights(dge.filtered.norm,
                            design,
                            plot = TRUE)
dev.off()

# condense gene info
genes <- v$genes
genes <- genes %>% select(all_of(c("gene_name_unique","gene_name",
                                   "gene_id","seqnames","start","end",
                                   "strand","exonic_length")))
v$genes <- genes

# fits linear model for each gene given a series of arrays
fit <- lmFit(v, design)

# contrast design for differential expression
contrasts <- makeContrasts(
  contrasts = myContrasts,
  levels = design.names)
head(contrasts)

# save contrast names
allComparisons <- colnames(contrasts)

# save
path <- paste0(out, "postfiltering_QC/voom_gene_final_mean_var.pdf")
pdf(path, width = 6, height = 4)

# run contrast analysis
vfit <- contrasts.fit(fit, contrasts = contrasts)

# Compute differential expression based on the empirical Bayes moderation of the
# standard errors towards a common value.
veBayesFit <- eBayes(vfit)
plotSA(veBayesFit, main = "Final Model: Mean-variance Trend")
dev.off()

# Cleanup
remove(fit,vfit,genes,design)
```

# Voom MDS Plots
## Sex
```{r mds_postfiltering_sex, warning=FALSE}
# loop through dimensions and top genes
names <- v$targets$sample
dims <- list(c(1,2),c(2,3),c(3,4))
topGenes <- c(100,500)

for (dim in dims) {
  
  for (numGenes in topGenes) {
      
    mds <- plotMDS(v, 
                   labels = names, 
                   top = numGenes,
                   plot = FALSE,
                   dim.plot = dim)

    toplot <- data.frame(DimX = mds$x, 
                         DimY = mds$y,
                         Sample = rownames(mds$distance.matrix.squared),
                         sex = v$targets$sex)
    
    x.lab <- paste0(mds$axislabel," ",mds$dim.plot[1]," (",
                    round(mds$var.explained[mds$dim.plot[1]] * 100, digits = 2),
                    "%)")
    y.lab <- paste0(mds$axislabel," ",mds$dim.plot[2]," (",
                    round(mds$var.explained[mds$dim.plot[2]] * 100, digits = 2),
                    "%)")
    
    p <- ggplot(toplot, aes(x = DimX, y = DimY, color = sex, label = Sample)) + 
      geom_point(size = 3) +
      scale_color_manual(values = sex_colors) +
      theme_bw() +
      geom_text(aes(label = Sample), size=4, vjust = -1) +
      xlab(x.lab) + ylab(y.lab) + 
      ggtitle(paste0("MDS Filtered: Top ", numGenes, ", Log2(CPM)"))
    print(p)
    
    # Set export path
    dimTitle <- as.character(dim)
    path <- paste0(out, "postfiltering_QC/MDS_top_", numGenes, "_genes/MDS_filtered_top_", 
                   numGenes, "_", "dim", as.character(dim)[1], "&", as.character(dim[2]), 
                   "_colored_by_sex.pdf")
    
    # Save
    pdf(path, width = 8, height = 8)
    print(p)
    dev.off()
  }
}

remove(data)
```

## Group
```{r mds_postfiltering_group}
# loop through dimensions and top genes
names <- v$targets$sample
dims <- list(c(1,2),c(2,3),c(3,4))
topGenes <- c(100,500)

for (dim in dims) {
  
  for (numGenes in topGenes) {
      
    mds <- plotMDS(v, 
                   labels = names, 
                   top = numGenes,
                   plot = FALSE,
                   dim.plot = dim)

    toplot <- data.frame(DimX = mds$x, 
                         DimY = mds$y,
                         Sample = rownames(mds$distance.matrix.squared),
                         group = v$targets$group)
    
    x.lab <- paste0(mds$axislabel," ",mds$dim.plot[1]," (",
                    round(mds$var.explained[mds$dim.plot[1]] * 100, digits = 2),
                    "%)")
    y.lab <- paste0(mds$axislabel," ",mds$dim.plot[2]," (",
                    round(mds$var.explained[mds$dim.plot[2]] * 100, digits = 2),
                    "%)")
    
    p <- ggplot(toplot, aes(x = DimX, y = DimY, color = group, label = Sample)) + 
      geom_point(size = 3) +
      scale_color_manual(values = group_colors) +
      theme_bw() +
      geom_text(aes(label = Sample), size=4, vjust = -1) +
      xlab(x.lab) + ylab(y.lab) + 
      ggtitle(paste0("MDS Filtered: Top ", numGenes, ", Log2(CPM)"))
    print(p)
    
    # save
    # Set export path
    dimTitle <- as.character(dim)
    path <- paste0(out, "postfiltering_QC/MDS_top_", numGenes, "_genes/MDS_filtered_top_", 
                   numGenes, "_", "dim", as.character(dim)[1], "&", as.character(dim[2]), 
                   "_colored_by_group.pdf")
    
    # Save
    pdf(path, width = 8, height = 8)
    print(p)
    dev.off()
  }
}
```

## Time point
```{r mds_postfiltering_timepoint}
# loop through dimensions and top genes
names <- v$targets$sample
dims <- list(c(1,2),c(2,3),c(3,4))
topGenes <- c(100,500)

for (dim in dims) {
  
  for (numGenes in topGenes) {
      
    mds <- plotMDS(v, 
                   labels = names, 
                   top = numGenes,
                   plot = FALSE,
                   dim.plot = dim)

    toplot <- data.frame(DimX = mds$x, 
                         DimY = mds$y,
                         Sample = rownames(mds$distance.matrix.squared),
                         timepoint = v$targets$timepoint_days)
    
    x.lab <- paste0(mds$axislabel," ",mds$dim.plot[1]," (",
                    round(mds$var.explained[mds$dim.plot[1]] * 100, digits = 2),
                    "%)")
    y.lab <- paste0(mds$axislabel," ",mds$dim.plot[2]," (",
                    round(mds$var.explained[mds$dim.plot[2]] * 100, digits = 2),
                    "%)")
    
    p <- ggplot(toplot, aes(x = DimX, y = DimY, color = timepoint, label = Sample)) + 
      geom_point(size = 3) +
      scale_color_manual(values = timepoint_colors) +
      theme_bw() +
      geom_text(aes(label = Sample), size=4, vjust = -1) +
      xlab(x.lab) + ylab(y.lab) + 
      ggtitle(paste0("MDS Filtered: Top ", numGenes, ", Log2(CPM)"))
    print(p)
    
    # Set export path
    dimTitle <- as.character(dim)
    path <- paste0(out, "postfiltering_QC/MDS_top_", numGenes, "_genes/MDS_filtered_top_", 
                   numGenes, "_", "dim", as.character(dim)[1], "&", as.character(dim[2]), 
                   "_colored_by_timepoint.pdf")
    
    # Save
    pdf(path, width = 8, height = 8)
    print(p)
    dev.off()
  }
}
```

## Treatment
```{r mds_postfiltering_treatment}
# loop through dimensions and top genes
names <- v$targets$sample
dims <- list(c(1,2),c(2,3),c(3,4))
topGenes <- c(100,500)

for (dim in dims) {
  
  for (numGenes in topGenes) {
      
    mds <- plotMDS(v, 
                   labels = names, 
                   top = numGenes,
                   plot = FALSE,
                   dim.plot = dim)

    toplot <- data.frame(DimX = mds$x, 
                         DimY = mds$y,
                         Sample = rownames(mds$distance.matrix.squared),
                         treatment = v$targets$treatment)
    
    x.lab <- paste0(mds$axislabel," ",mds$dim.plot[1]," (",
                    round(mds$var.explained[mds$dim.plot[1]] * 100, digits = 2),
                    "%)")
    y.lab <- paste0(mds$axislabel," ",mds$dim.plot[2]," (",
                    round(mds$var.explained[mds$dim.plot[2]] * 100, digits = 2),
                    "%)")
    
    p <- ggplot(toplot, aes(x = DimX, y = DimY, color = treatment, label = Sample)) + 
      geom_point(size = 3) +
      scale_color_manual(values = treatment_colors) +
      theme_bw() +
      geom_text(aes(label = Sample), size=4, vjust = -1) +
      xlab(x.lab) + ylab(y.lab) + 
      ggtitle(paste0("MDS Filtered: Top ", numGenes, ", Log2(CPM)"))
    print(p)
    
    # Set export path
    dimTitle <- as.character(dim)
    path <- paste0(out, "postfiltering_QC/MDS_top_", numGenes, "_genes/MDS_filtered_top_", 
                   numGenes, "_", "dim", as.character(dim)[1], "&", as.character(dim[2]), 
                   "_colored_by_treatment.pdf")
    
    # Save
    pdf(path, width = 8, height = 8)
    print(p)
    dev.off()
  }
}
```

## Age
```{r mds_postfiltering_RIN}
# loop through dimensions and top genes
names <- v$targets$sample
dims <- list(c(1,2),c(2,3),c(3,4))
topGenes <- c(100,500)

for (dim in dims) {
  
  for (numGenes in topGenes) {
      
    mds <- plotMDS(v, 
                   labels = names, 
                   top = numGenes,
                   plot = FALSE,
                   dim.plot = dim)

    toplot <- data.frame(DimX = mds$x, 
                         DimY = mds$y,
                         Sample = rownames(mds$distance.matrix.squared),
                         age = v$targets$age)
    
    x.lab <- paste0(mds$axislabel," ",mds$dim.plot[1]," (",
                    round(mds$var.explained[mds$dim.plot[1]] * 100, digits = 2),
                    "%)")
    y.lab <- paste0(mds$axislabel," ",mds$dim.plot[2]," (",
                    round(mds$var.explained[mds$dim.plot[2]] * 100, digits = 2),
                    "%)")
    
    p <- ggplot(toplot, aes(x = DimX, y = DimY, color = age, label = Sample)) + 
      geom_point(size = 3) +
      scale_color_manual(values = age_colors) +
      theme_bw() +
      geom_text(aes(label = Sample), size=4, vjust = -1) +
      xlab(x.lab) + ylab(y.lab) + 
      ggtitle(paste0("MDS Filtered: Top ", numGenes, ", Log2(CPM)"))
    print(p)
    
    # Set export path
    dimTitle <- as.character(dim)
    path <- paste0(out, "postfiltering_QC/MDS_top_", numGenes, "_genes/MDS_filtered_top_", 
                   numGenes, "_", "dim", as.character(dim)[1], "&", as.character(dim[2]), 
                   "_colored_by_age.pdf")
    
    # Save
    pdf(path, width = 8, height = 8)
    print(p)
    dev.off()
  }
}
```

# Differential Expressioin
## Number of DEGs
Identify number of differentially expressed genes.
```{r decide_tests}
# set thresholds
pval <- 0.05
my.lfc <- 0

# run tests
sumTable <- summary(decideTests(
  veBayesFit,
  adjust.method = "BH",
  p.value = pval,
  lfc = my.lfc
  ))
print(paste0(" FDRq < ", pval, ", LFC = ", my.lfc))
sumTable

# reformat table
sumTable <- as.data.frame(sumTable)
colnames(sumTable) <- c("direction", "comparison", "count")
sumTable <- sumTable |> 
  pivot_wider(names_from = comparison, values_from = count)
colnames(sumTable) <- gsub(" - ", "_vs_", colnames(sumTable))

# save table
path <- paste0(out, model, "DEG_tables/DEG_comparison_FDRq_", 
               format(pval, nsmall = 2), "_LFC_", format(my.lfc, nsmall = 2), 
               ".tsv")
write.table(sumTable, 
            path, 
            quote = FALSE, 
            sep = "\t")
```

## DEG heatmap both sexes
```{r deg_heatmap_both_sexes}
# reformat before plotting
df <- as.data.frame(sumTable)
df <- df[!df$direction == "NotSig",]
rownames(df) <- df$direction
df$direction <- NULL
df <- t(df)

# set heatmap colors and names
meta <- data.frame(comparison = c(rep("E. coli vs Saline", 4),
                                  rep("21 days vs 2 days", 4),
                                  rep("Old vs Young", 4)))
rownames(meta) <- rownames(df)
paletteLength <- 100
myColor <- colorRampPalette(c("white","#f0eb9e","darkgreen"))(paletteLength)
ann_colors = list(comparison = c(`E. coli vs Saline` = "cornflowerblue",
                                 `21 days vs 2 days` = "firebrick2",
                                 `Old vs Young` = "gold"))

# save
path <- paste0(out, model, "/DEG_tables/DEG_comparison_FDRq_",
               format(pval, nsmall = 2),
               "_LFC_", format(my.lfc, nsmall = 2),".pdf")
pdf(path, width = 8, height = 9)

# plot
pheatmap::pheatmap(df,
                   main = paste0("FDRq < ", pval, ", |LFC| > ", my.lfc),
                   treeheight_row = 0,
                   treeheight_col = 0,
                   color = myColor,
                   cluster_rows = FALSE,
                   annotation_row = meta,
                   annotation_colors = ann_colors,
                   display_numbers = round(df, digits = 0),
                   fontsize_number = 12,
                   number_color = "black")
```

## DEG heatmap sex specific
```{r deg_heatmap_sex_specific}
# reformat before plotting
df <- as.data.frame(sumTable)
df <- df[!df$direction == "NotSig",]
rownames(df) <- df$direction
df$direction <- NULL
df <- t(df)

# set heatmap colors and names
meta <- data.frame(comparison = c(rep("E. coli vs Saline", 8),
                                  rep("21 days vs 2 days", 8),
                                  rep("Old vs Young", 8),
                                  rep("Female vs Male", 8)))
rownames(meta) <- rownames(df)
paletteLength <- 100
myColor <- colorRampPalette(c("white","#f0eb9e","darkgreen"))(paletteLength)
ann_colors = list(comparison = c(`E. coli vs Saline` = "cornflowerblue",
                                 `21 days vs 2 days` = "firebrick2",
                                 `Old vs Young` = "gold",
                                 `Female vs Male` = "chartreuse2"))

# save
path <- paste0(out, model, "/DEG_tables/DEG_comparison_FDRq_",
               format(pval, nsmall = 2),
               "_LFC_", format(my.lfc, nsmall = 2),".pdf")
pdf(path, width = 8, height = 12)

# plot
pheatmap::pheatmap(df,
                   main = paste0("FDRq < ", pval, ", |LFC| > ", my.lfc),
                   treeheight_row = 0,
                   treeheight_col = 0,
                   color = myColor,
                   cluster_rows = FALSE,
                   annotation_row = meta,
                   annotation_colors = ann_colors,
                   display_numbers = round(df, digits = 0),
                   fontsize_number = 12,
                   number_color = "black")
```

## DEG heatmap timepoint and sex specific
```{r}
# reformat before plotting
df <- as.data.frame(sumTable)
df <- df[!df$direction == "NotSig",]
rownames(df) <- df$direction
df$direction <- NULL
df <- t(df)

# set heatmap colors and names
meta <- data.frame(treatment = c(rep("low dose", 2),
                                rep("high dose", 2),
                                rep("other", 3)),
                   sex = c(rep(c("female","male"),2), rep("both",3)))
rownames(meta) <- rownames(df)
paletteLength <- 100
myColor <- colorRampPalette(c("white","#f0eb9e","darkgreen"))(paletteLength)
ann_colors = list(treatment = c(`low dose` = "cornflowerblue", 
                               `high dose` = "firebrick2",
                               `other` = "gray"),
                  sex = c(female = 'chartreuse2',
                          male = 'purple',
                          both = "gold"))

# save
path <- paste0(out, model, "/DEG_tables/DEG_comparison_FDRq_",
               format(pval, nsmall = 2),
               "_LFC_", format(my.lfc, nsmall = 2),".pdf")
pdf(path, width = 8, height = 9)

# plot
pheatmap::pheatmap(df,
                   main = paste0("FDRq < ", pval, ", |LFC| > ", my.lfc),
                   treeheight_row = 0,
                   treeheight_col = 0,
                   color = myColor,
                   cluster_rows = FALSE,
                   annotation_row = meta,
                   annotation_colors = ann_colors,
                   display_numbers = round(df, digits = 0),
                   fontsize_number = 12,
                   number_color = "black")
```

## Output DEG tables
```{r output_DEG_tables}
coef <- 1

for (i in allComparisons) {
  # title 
  title <- gsub(" ", "_", i)
  title <- gsub("-","vs", title)
  
  
  # p < 1, log2fc > 0 ----------------------------------------------------------
  vTopTableAll <-
    topTable(
      veBayesFit, 
      coef = coef,  
      n = Inf, 
      p.value = 1,
      lfc = 0,
      confint = TRUE
    )
  #output as txt file
  path <- paste(out, model, "DEG_tables/", title, "_FDRq_1.00_LFC_0.00.tsv", sep = "") 
  write.table(
    vTopTableAll,
    path,
    sep = "\t",
    row.names = FALSE,
    quote = FALSE
  )
  
  # p < 0.05 -------------------------------------------------------------------
  vTopTable1 <-
    topTable( 
      veBayesFit,  
      coef = coef,  
      n = Inf, 
      p.value = 0.05,
      lfc = 0,
      confint = TRUE
    )
  path <- paste(out, model, "DEG_tables/", title, "_FDRq_0.05_LFC_0.00.tsv", sep = "") 
  write.table(
    vTopTable1,
    path,
    sep = "\t",
    row.names = FALSE,
    quote = FALSE
  )
  
  # increment ------------------------------------------------------------------
  coef <- coef + 1
}
```

## Volcano
```{r volcano_plot, message=FALSE, eval=FALSE}
# set variables
thresh <- 0.05
my.lfc <- 0

# get file list
files <- list.files(paste0(out, model,"DEG_tables/"))
keep <- grep("_1.00_LFC_0.00.tsv", files)
files <- files[keep]
files <- paste0(out, model, "DEG_tables/", files)

for (i in files) {
  
  # read DEG file
  data <- read.delim(i,sep = "\t")
  
  # assign colors
  color_values <- vector()
  max <- nrow(data)
  for(row in 1:max){
    if (data$adj.P.Val[row] < thresh){
      if (data$logFC [row] > my.lfc){
        color_values <- c(color_values, 1) # 1 when logFC > my.lfc and FDRq < thresh
      } else if (data$logFC[row] < -my.lfc){
        color_values <- c(color_values, 2) # 2 when logFC < -my.lfc and FDRq < thresh
      } else {
        color_values <- c(color_values, 3) # 3 when logFC between -my.lfc and +my.lfc and FDRq < thresh
      }
    } else {
      color_values <- c(color_values, 3) # 3 when FDRq >= thresh
    }
  }
  data$color_adjpval <- factor(color_values)
  
 # comparison name
 comparison <- gsub(paste0(out, model, "DEG_tables/"), "", i)
 comparison <- gsub("_FDRq_1.00_LFC_0.00.tsv","",comparison)
    
 # plot only if there are DEGs with adj.P.Val < thresh
 num <- data[data$adj.P.Val < thresh,]
 num <- nrow(num)
 if(num != 0) {
   
    # subset genes to label
    up <- data[data$color_adjpval == 1,]
    up.sig <- up[order(up$adj.P.Val),][1:15,]
    up.lfc <- up[order(up$logFC, decreasing = TRUE),][1:15,]
    up30 <- rbind(up.sig,up.lfc)
    up30 <- up30[!duplicated(up30$gene_name_unique),]
    down <- data[data$color_adjpval == 2,]
    down.sig <- down[order(down$adj.P.Val),][1:15,]
    down.lfc <- down[order(down$logFC, decreasing = FALSE),][1:15,]
    down30 <- rbind(down.sig,down.lfc)
    down30 <- down30[!duplicated(down30$gene_name_unique),]
      
    # set manual colors
    if (!1 %in% unique(data$color_adjpval)) {
      my_colors <- c("blue","gray")
    } else if (!2 %in% unique(data$color_adjpval)) {
      my_colors <- c("red","gray")
    } else if (!1 %in% unique(data$color_adjpval) && !2 %in% unique(data$color_adjpval)) {
      my_colors <- c("gray")
    } else {
      my_colors <- c("red","blue","gray")
    }
      
    # set significance threshold
    hadjpval <- (-log10(max(data$P.Value[data$adj.P.Val < thresh], na.rm=TRUE)))

    # plot
    p <-
      ggplot(data = data, 
             aes(x = logFC,  # x-axis is logFC
                 y = -log10(P.Value),  # y-axis will be -log10 of P.Value
                 color = color_adjpval)) +  # color is based on factored color column
      geom_point(alpha = 0.8, size = 2) +  # create scatterplot, alpha makes points transparent
      theme_bw() +  # set color theme
      theme(legend.position = "none") +  # no legend
      scale_color_manual(values = my_colors) +  # set factor colors
      labs(
        title = "", # no main title
        x = expression(log[2](FC)), # x-axis title
         y = expression(-log[10] ~ "(" ~ italic("p") ~ "-value)") # y-axis title
      ) +
      theme(axis.title.x = element_text(size = 15),
            axis.text.x = element_text(size = 15)) +
      theme(axis.title.y = element_text(size = 15),
            axis.text.y = element_text(size = 15)) +
      theme(plot.title = element_text(size = 15)) +
      geom_hline(yintercept = hadjpval,  #  horizontal line
                         colour = "#000000",
                         linetype = "dashed") +
      geom_vline(xintercept = -my.lfc,  #  vertical line
                         colour = "#000000",
                         linetype = "dashed") +
      geom_vline(xintercept = my.lfc,  #  vertical line
                         colour = "#000000",
                         linetype = "dashed") +
      ggtitle(paste0(comparison, ", adj.P.Val < ", thresh, ", LFC = ", my.lfc)) +
      geom_text_repel(data = up30,
                      aes(x = logFC, y= -log10(P.Value), label = gene_name), 
                      size = 5,
                      color = "maroon", 
                      fontface="italic",
                      max.overlaps = getOption("ggrepel.max.overlaps", default = 30)
                      ) +
      geom_text_repel(data = down30,
                      aes(x = logFC, y= -log10(P.Value), label = gene_name), 
                      color = "navyblue", 
                      size = 5,
                      fontface="italic",
                      max.overlaps = getOption("ggrepel.max.overlaps", default = 30)
                      ) +
      ylim(c(0,15))
     p
      
    # save
    path <- paste0(out, model, "volcano/", comparison, "_FDRq_", 
                   format(thresh, nsmall = 2), "_LFC_",
                   format(my.lfc, nsmall = 2), "_volcano.pdf")
    pdf(path, height = 8, width = 8)
    print(p)
    dev.off()
  }
} # end loop through variables
```

## Single volcano
```{r}
# set variables
thresh <- 0.1
my.lfc <- 0.2

# get file list
files <- list.files(paste0(out, model,"DEG_tables/"))
keep <- grep("_1.00_LFC_0.00.tsv", files)
files <- files[keep]
files <- paste0(out, model, "DEG_tables/", files)
files <- files[grepl(pattern = "DEG_tables/S\\.", x = files)]
files <- paste0(out, model, "DEG_tables/L.8h.M_vs_S.8h.M_FDRq_1.00_LFC_0.00.tsv")

for (i in files) {
  
  # read DEG file
  data <- read.delim(i,sep = "\t")
  
  # assign colors
  color_values <- vector()
  max <- nrow(data)
  for(row in 1:max){
    if (data$adj.P.Val[row] < thresh){
      if (data$logFC [row] > my.lfc){
        color_values <- c(color_values, 1) # 1 when logFC > my.lfc and FDRq < thresh
      } else if (data$logFC[row] < -my.lfc){
        color_values <- c(color_values, 2) # 2 when logFC < -my.lfc and FDRq < thresh
      } else {
        color_values <- c(color_values, 3) # 3 when logFC between -my.lfc and +my.lfc and FDRq < thresh
      }
    } else {
      color_values <- c(color_values, 3) # 3 when FDRq >= thresh
    }
  }
  data$color_adjpval <- factor(color_values)
  
 # comparison name
 comparison <- gsub(paste0(out, model, "DEG_tables/"), "", i)
 comparison <- gsub("_FDRq_1.00_LFC_0.00.tsv","",comparison)
    
 # plot only if there are DEGs with adj.P.Val < thresh
 num <- data[data$adj.P.Val < thresh,]
 num <- nrow(num)
 if(num != 0) {
   
    # subset genes to label
    up <- data[data$color_adjpval == 1,]
    up.sig <- up[order(up$adj.P.Val),][1:15,]
    up.lfc <- up[order(up$logFC, decreasing = TRUE),][1:15,]
    up30 <- rbind(up.sig,up.lfc)
    up30 <- up30[!duplicated(up30$gene_name_unique),]
    down <- data[data$color_adjpval == 2,]
    down.sig <- down[order(down$adj.P.Val),][1:15,]
    down.lfc <- down[order(down$logFC, decreasing = FALSE),][1:15,]
    down30 <- rbind(down.sig,down.lfc)
    down30 <- down30[!duplicated(down30$gene_name_unique),]
      
    # set manual colors
    if (!1 %in% unique(data$color_adjpval)) {
      my_colors <- c("blue","gray")
    } else if (!2 %in% unique(data$color_adjpval)) {
      my_colors <- c("red","gray")
    } else if (!1 %in% unique(data$color_adjpval) && !2 %in% unique(data$color_adjpval)) {
      my_colors <- c("gray")
    } else {
      my_colors <- c("red","blue","gray")
    }
      
    # set significance threshold
    hadjpval <- (-log10(max(data$adj.P.Val[data$adj.P.Val < thresh], na.rm=TRUE)))

    # plot
    p <-
      ggplot(data = data, 
             aes(x = logFC,  # x-axis is logFC
                 y = -log10(adj.P.Val),  # y-axis will be -log10 of adj.P.Val
                 color = color_adjpval)) +  # color is based on factored color column
      geom_point(alpha = 0.8, size = 2) +  # create scatterplot, alpha makes points transparent
      theme_bw() +  # set color theme
      theme(legend.position = "none") +  # no legend
      scale_color_manual(values = my_colors) +  # set factor colors
      labs(
        title = "", # no main title
        x = expression(log[2](FC)), # x-axis title
         y = expression(-log[10] ~ "(adjsuted p-value)") # y-axis title
      ) +
      theme(axis.title.x = element_text(size = 15),
            axis.text.x = element_text(size = 15)) +
      theme(axis.title.y = element_text(size = 15),
            axis.text.y = element_text(size = 15)) +
      theme(plot.title = element_text(size = 15)) +
      geom_hline(yintercept = hadjpval,  #  horizontal line
                         colour = "#000000",
                         linetype = "dashed") +
      geom_vline(xintercept = -my.lfc,  #  vertical line
                         colour = "#000000",
                         linetype = "dashed") +
      geom_vline(xintercept = my.lfc,  #  vertical line
                         colour = "#000000",
                         linetype = "dashed") +
      ggtitle(paste0(comparison, ", adj.P.Val < ", thresh, ", LFC = ", my.lfc)) +
      geom_text_repel(data = up30,
                      aes(x = logFC, y= -log10(adj.P.Val), label = gene_name), 
                      size = 5,
                      color = "maroon", 
                      fontface="italic",
                      max.overlaps = getOption("ggrepel.max.overlaps", default = 30)
                      ) +
      geom_text_repel(data = down30,
                      aes(x = logFC, y= -log10(adj.P.Val), label = gene_name), 
                      color = "navyblue", 
                      size = 5,
                      fontface="italic",
                      max.overlaps = getOption("ggrepel.max.overlaps", default = 30)
                      )
     p
      
    # save
    path <- paste0(out, model, "volcano/sex_linked_removed_", comparison, "_FDRq_", 
                   format(thresh, nsmall = 2), "_LFC_",
                   format(my.lfc, nsmall = 2), "_volcano.pdf")
    pdf(path, height = 8, width = 8)
    print(p)
    dev.off()
  }
} # end loop through variables
```

## Metascape input
```{r}
# set variables
thresh <- 0.05
my.lfc <- 0

# get file list
files <- list.files(paste0(out, model,"DEG_tables/"))
keep <- grep("_1.00_LFC_0.00.tsv", files)
files <- paste0(out, model, "DEG_tables/", files[keep])

# intitalize table for command line version of metascape
msbio.df <- data.frame()

# loop through DEG files
for (i in 1:length(files)) {
  # read table
  data <- read.table(files[i], header = TRUE, sep = "\t")
  
  # filter based on adjusted p-value
  data <- data[data$adj.P.Val < thresh,]
  
  # create up-regulated gene list
  up <- data[data$logFC > my.lfc,]
  up <- up$gene_name
  
  # create down-regulated gene list
  down <- data[data$logFC < -my.lfc,]
  down <- down$gene_name
  
 # get the comparison name
 comparison <- gsub(paste0(out, model, "DEG_tables/"), "", files[i])
 comparison <- gsub("_FDRq_1.00_LFC_0.00.tsv","",comparison)
 
 # make two filenames for up and downregulated genes
 # include the comparison, adjusted p-value threshold, and logFC threshold
 up.filename <- paste0(out, model, "metascape_input/", comparison,
                       "_upregulated_FDRq_", format(thresh, nsmall = 2), 
                       "_LFC_", format(my.lfc, nsmall = 2), ".txt")
 down.filename <- paste0(out, model, "metascape_input/", comparison,
                         "_downregulated_FDRq_", format(thresh, nsmall = 2), 
                         "_LFC_", format(my.lfc, nsmall = 2), ".txt")
 
 
  # save the up and down-reulated gene lists
  write.table(x = up,
              file = up.filename,
              quote = FALSE,
              row.names = FALSE,
              col.names = FALSE)
  write.table(x = down,
              file = down.filename,
              quote = FALSE,
              row.names = FALSE,
              col.names = FALSE)
    
  # job for MSBio
  # if there are more than 10 up-regulated genes, add paths in json format for msbio job
  if (length(up) > 10) {
      msbio <- jsonlite::toJSON(list(
        input = gsub(paste0(out, model, "metascape_input/"),
                     "/data/metascape_input/",
                     up.filename),
        output = paste0("/data/metascape_output/",
                        comparison,
                        "_upregulated_", 
                        format(thresh, nsmall = 2), 
                        "_LFC_", format(my.lfc, nsmall = 2)),
        single = TRUE))
      msbio <- gsub("\\[","",msbio)
      msbio <- gsub("\\]","",msbio)
      msbio.df <- rbind(msbio.df, msbio)
  }
  # if there are more than 10 down-regulated genes, add paths in json format for msbio job
  if (length(down) > 10) {
      msbio <- jsonlite::toJSON(list(
        input = gsub(paste0(out, model, "metascape_input/"),
                     "/data/metascape_input/",
                     down.filename),
        output = paste0("/data/metascape_output/",
                        comparison,
                        "_downregulated_", 
                        format(thresh, nsmall = 2),
                        "_LFC_", 
                        format(my.lfc, nsmall = 2)),
        single = TRUE))
      msbio <- gsub("\\[","",msbio)
      msbio <- gsub("\\]","",msbio)
      msbio.df <- rbind(msbio.df, msbio)
  }
}

# save msbio job
write.table(x = msbio.df,
            file = paste0(out, model, "metascape_input/msbio.job"),
            quote = FALSE,
            row.names = FALSE,
            col.names = FALSE)
```
