---
title: "E. coli Mice bulkRNAseq"
subtitle: "Figures"
author: "Kennedi Todd"
date: "02/23/2024"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

# Setup
## Working directory
```{r setup}
knitr::opts_knit$set(root.dir = ".")
```

## Libraries
```{r libraries, message=FALSE, warning=FALSE}
library(plotly)
library(tidyr)
library(ComplexUpset)
library(UpSetR)
library(gridExtra)   # grid.arrange()
library(ggrepel)
```

## Functions
```{r get_exclusiver_intersection_genes}
get_exclusive_intersection_genes <- function(upset_data, list_input) {
  # Initialize a list to store the results
  results <- list()
  
  # Get the names of the columns (gene sets)
  set_names <- colnames(upset_data)
  
  # For each row of the UpSet matrix, extract the exclusive intersection
  for (i in 1:nrow(upset_data)) {
    # Check which sets (columns) are "active" (i.e., 1) for the current row
    active_sets <- set_names[upset_data[i, ] == 1]
    
    if (length(active_sets) > 0) {
      # Find the intersection of genes in the active sets
      intersected_genes <- Reduce(intersect, list_input[active_sets])
      
      # Find genes that are present in any of the other sets (not part of this combination)
      other_sets <- set_names[upset_data[i, ] == 0]
      other_genes <- Reduce(union, list_input[other_sets])
      
      # Subtract the genes that are in the other sets to make the intersection exclusive
      exclusive_genes <- setdiff(intersected_genes, other_genes)
      
      # Create a label for the current intersection (e.g., "Set1 & Set2")
      intersection_label <- paste(active_sets, collapse = " & ")
      
      # Store the number of genes and the exclusive gene list
      results[[intersection_label]] <- list(
        total_genes = length(exclusive_genes),
        gene_list = paste(exclusive_genes, collapse = ", ")
      )
    }
  }
  
  return(results)
}
```

# Heatmap 1
- Created an upset plot with 4 genes lists
  - E.2D.Y vs S.2D.Y up-regulated
  - E.2D.Y vs S.2D.Y down-regulated
  - E.2D.O vs S.2D.O up-regulated
  - E.2D.O vs S.2D.O down-regulated
- There were 99 overlapping genes for the up-regulated gene lists
- Create a heatmap looking at the immune and endothelial overlapping genes
```{r}
# save path
out <- "../results/both_sexes/"

# genes categories
h1_categories <- read.delim2("../refs/heatmap_gene_categories.tsv")

# heatmap 1 genes
h1_genes <- read.delim2(file = "../refs/heatmap1_genes.tsv")

# add category info
h1_genes <- dplyr::left_join(x = h1_genes, y = h1_categories)
remove(h1_categories)
rownames(h1_genes) <- h1_genes$gene_name

# get DEG files
files <- list.files(paste0(out, "DEGs_with_hbb/DEG_tables/"))
files <- files[grep("E.2D.+_vs_S.2D.+_1.00_LFC_0.00.tsv", files)]

# init
h1_lfc <- data.frame(row.names = h1_genes$gene_name)

for (i in seq_along(files)) {
  # read table
  data <- read.delim(
    file = paste0(out, "DEGs_with_hbb/DEG_tables/", files[i]),
    sep = "\t", check.names = FALSE
  )

  # comparison name
  comparison <- sub("_FDRq_1.00_LFC_0.00\\.tsv$", "", files[i])

  # subset rows and columns
  data <- data[data$gene_name %in% h1_genes$gene_name, c("gene_name","logFC", "adj.P.Val")]
  rownames(data) <- data$gene_name
  data <- data[match(h1_genes$gene_name, rownames(data)), , drop = FALSE]

  # ensure numeric (in case it was read as character)
  data$logFC <- round(as.numeric(data$logFC), 4)
  data$adj.P.Val <- as.numeric(data$adj.P.Val)

  # assign logFC column
  h1_lfc[[comparison]] <- data$logFC
}
remove(data)

# rename columns
colnames(h1_lfc) <- c("E. coli Old", "E. coli Young")
h1_lfc <- h1_lfc[,c(2,1)]

# set heatmap colors
paletteLength <- 100
h1_color_scale <- colorRampPalette(c("cornflowerblue", "white", "red"))(paletteLength)
h1_breaks <- c(seq(min(h1_lfc), 0, length.out = ceiling(paletteLength/2) + 1), 
              seq(max(h1_lfc)/paletteLength, max(h1_lfc), length.out = floor(paletteLength/2)))
h1_meta <- data.frame(category = h1_genes$category)
rownames(h1_meta) <- rownames(h1_genes)
h1_ann_colors = list(category = c(immune = "chartreuse2", endothelial = "magenta"))

# save
pdf(paste0(out, "custom_figures/heatmap1.pdf"), width = 5, height = 6)
pheatmap::pheatmap(h1_lfc,
                   main = "Overlapping Up-regulated Genes",
                   cluster_rows = FALSE,
                   cluster_cols = FALSE,
                   angle_col = 0,
                   annotation_colors = h1_ann_colors,
                   annotation_row = h1_meta,
                   color = h1_color_scale,
                   breaks = h1_breaks)

# cleanup
remove(h1_genes, h1_lfc, h1_meta, h1_ann_colors)
```

# Heatmap 2
- Taking some of the DEGs from...
  - E.2D.Y vs S.2D.Y up-regulated
  - E.2D.Y vs S.2D.Y down-regulated
  - E.2D.O vs S.2D.O up-regulated
  - E.2D.O vs S.2D.O down-regulated
```{r}
# save path
out <- "../results/both_sexes/"

# genes categories
h2_categories <- read.delim2("../refs/heatmap_gene_categories.tsv")

# heatmap 1 genes
h2_genes <- read.delim2(file = "../refs/heatmap2_genes.tsv")

# add category info
h2_genes <- dplyr::left_join(x = h2_genes, y = h2_categories)
remove(h2_categories)
rownames(h2_genes) <- h2_genes$gene_name

# get DEG files
files <- list.files(paste0(out, "DEGs_with_hbb/DEG_tables/"))
files <- files[grep("E.2D.+_vs_S.2D.+_1.00_LFC_0.00.tsv", files)]

# init
h2_lfc <- data.frame(row.names = h2_genes$gene_name)

for (i in seq_along(files)) {
  # read table
  data <- read.delim(
    file = paste0(out, "DEGs_with_hbb/DEG_tables/", files[i]),
    sep = "\t", check.names = FALSE
  )

  # comparison name
  comparison <- sub("_FDRq_1.00_LFC_0.00\\.tsv$", "", files[i])

  # subset rows and columns
  data <- data[data$gene_name %in% h2_genes$gene_name, c("gene_name","logFC", "adj.P.Val")]
  rownames(data) <- data$gene_name
  data <- data[match(h2_genes$gene_name, rownames(data)), , drop = FALSE]

  # ensure numeric (in case it was read as character)
  data$logFC <- round(as.numeric(data$logFC), 4)
  data$adj.P.Val <- as.numeric(data$adj.P.Val)

  # assign logFC column
  h2_lfc[[comparison]] <- data$logFC
}
remove(data)

# rename columns
colnames(h2_lfc) <- c("E. coli Old", "E. coli Young")
h2_lfc <- h2_lfc[,c(2,1)]

# set heatmap colors
paletteLength <- 100
h2_color_scale <- colorRampPalette(c("cornflowerblue", "white", "red"))(paletteLength)
h2_breaks <- c(seq(min(h2_lfc), 0, length.out = ceiling(paletteLength/2) + 1), 
              seq(max(h2_lfc)/paletteLength, max(h2_lfc), length.out = floor(paletteLength/2)))
h2_meta <- data.frame(category = h2_genes$category)
rownames(h2_meta) <- rownames(h2_genes)
h2_ann_colors = list(category = c(immune = "chartreuse2", endothelial = "magenta",
                                  proliferative = "yellow", metabolism = "cyan"))

# save
pdf(paste0(out, "custom_figures/heatmap2.pdf"), width = 7, height = 6)
pheatmap::pheatmap(h2_lfc,
                   main = "E. coli 2 Days DEGs",
                   cluster_rows = FALSE,
                   cluster_cols = FALSE,
                   angle_col = 0,
                   annotation_colors = h2_ann_colors,
                   annotation_row = h2_meta,
                   color = h2_color_scale,
                   breaks = h2_breaks)
```

# Heatmaps
## 2 days with pval
- sex specific filtering but both sexes comparison \
- FDRq < 0.05 and LFC=0 \
- Heatmap for 2-day timepoints \
```{r}
# save path
out <- "../results/both_sexes/"

# genes of interest
goi_2D <- c("S100a8","Lcn2","S100a9","Ctla2a","Lrg1","Top2a","Pkn3","Igsf6",
            "Srarp","Cd52","Edn1","Vil1","Rhoh","Tmem252","Vwf","Myct1",
            "Clec14a","Lyz1","Fermt1","Gpr160","Tnfsf10","Slc38a5","Ctss",
            "Ly6a","C1qa","C1qb","C1qc","Csf3r","Mpeg1","Vim","Pecam1",
            "Adgre1","Itgam","Hif3a","Sgk1","Smox","Mfsd2a","Ly6c1","Tcim")

# get DEG files
files <- list.files(paste0(out, "DEGs_with_hbb/DEG_tables/"))
files <- files[grep("E.2D.+_vs_S.2D.+_1.00_LFC_0.00.tsv", files)]

# init
lfc.df <- data.frame(row.names = goi_2D)
adjpval.df <- data.frame(row.names = goi_2D)

for (i in seq_along(files)) {
  # read table
  data <- read.delim(
    file = paste0(out, "DEGs_with_hbb/DEG_tables/", files[i]),
    sep = "\t", check.names = FALSE
  )

  # comparison name
  comparison <- sub("_FDRq_1.00_LFC_0.00\\.tsv$", "", files[i])

  # filter + order to goi_2D
  data <- data[data$gene_name %in% goi_2D, c("gene_name","logFC", "adj.P.Val")]
  rownames(data) <- data$gene_name
  data <- data[match(goi_2D, rownames(data)), , drop = FALSE]

  # ensure numeric (in case it was read as character)
  data$logFC <- round(as.numeric(data$logFC), 4)
  data$adj.P.Val <- as.numeric(data$adj.P.Val)

  # assign just the logFC column
  lfc.df[[comparison]] <- data$logFC
  adjpval.df[[comparison]] <- data$adj.P.Val
}

# set heatmap colors and names
paletteLength <- 100
myColor <- colorRampPalette(c("cornflowerblue", "white", "red"))(paletteLength)
myBreaks <- c(seq(min(lfc.df), 0, length.out = ceiling(paletteLength/2) + 1), 
              seq(max(lfc.df)/paletteLength, max(lfc.df), length.out = floor(paletteLength/2)))

# save
pdf(paste0(out, "custom_figures/2D_logFC_adjPVal_heatmap.pdf"), width = 5, height = 10)


pheatmap::pheatmap(lfc.df,
                   main = "logFC and adj.P.Val for 2 Day Time Points",
                   cluster_rows = FALSE,
                   cluster_cols = FALSE,
                   annotation_colors = ann_colors,
                   color = myColor,
                   breaks = myBreaks,
                   display_numbers = round(adjpval.df, digits = 4),
                   fontsize_number = 12,
                   number_color = "black",
                   )
```

## 2 days w/out pval
```{r}
# save path
out <- "../results/both_sexes/"

# genes of interest
goi_2D <- c("S100a8","Lcn2","S100a9","Ctla2a","Lrg1","Top2a","Pkn3","Igsf6",
            "Srarp","Cd52","Edn1","Vil1","Rhoh","Tmem252","Vwf","Myct1",
            "Clec14a","Lyz1","Fermt1","Gpr160","Tnfsf10","Slc38a5","Ctss",
            "Ly6a","C1qa","C1qb","C1qc","Csf3r","Mpeg1","Vim","Pecam1",
            "Adgre1","Itgam","Hif3a","Sgk1","Smox","Mfsd2a","Ly6c1","Tcim")

# get DEG files
files <- list.files(paste0(out, "DEGs_with_hbb/DEG_tables/"))
files <- files[grep("E.2D.+_vs_S.2D.+_1.00_LFC_0.00.tsv", files)]

# init
lfc.df <- data.frame(row.names = goi_2D)
adjpval.df <- data.frame(row.names = goi_2D)

for (i in seq_along(files)) {
  # read table
  data <- read.delim(
    file = paste0(out, "DEGs_with_hbb/DEG_tables/", files[i]),
    sep = "\t", check.names = FALSE
  )

  # comparison name
  comparison <- sub("_FDRq_1.00_LFC_0.00\\.tsv$", "", files[i])

  # filter + order to goi_2D
  data <- data[data$gene_name %in% goi_2D, c("gene_name","logFC", "adj.P.Val")]
  rownames(data) <- data$gene_name
  data <- data[match(goi_2D, rownames(data)), , drop = FALSE]

  # ensure numeric (in case it was read as character)
  data$logFC <- round(as.numeric(data$logFC), 4)
  data$adj.P.Val <- as.numeric(data$adj.P.Val)

  # assign just the logFC column
  lfc.df[[comparison]] <- data$logFC
  adjpval.df[[comparison]] <- data$adj.P.Val
}

# set heatmap colors and names
paletteLength <- 100
myColor <- colorRampPalette(c("cornflowerblue", "white", "red"))(paletteLength)
myBreaks <- c(seq(min(lfc.df), 0, length.out = ceiling(paletteLength/2) + 1), 
              seq(max(lfc.df)/paletteLength, max(lfc.df), length.out = floor(paletteLength/2)))

# save
pdf(paste0(out, "custom_figures/2D_logFC_heatmap.pdf"), width = 5, height = 10)


pheatmap::pheatmap(lfc.df,
                   main = "logFC for 2 Day Time Points",
                   cluster_rows = FALSE,
                   cluster_cols = FALSE,
                   annotation_colors = ann_colors,
                   color = myColor,
                   breaks = myBreaks)
```

## 21 days w/ pval
```{r}
# save path
out <- "../results/both_sexes/"

# genes of interest
goi_21D <- c("Hif3a", "Tlr13", "Xdh", "Skap1", "Ccdc88a", "Kcnb2", "Grik1", 
             "Adcy8", "H2-K1", "Higd1b", "Atp10a", "C1qtnf4", "Ppp3r1", "Gab3", 
             "Il1rapl1", "Rasgrp2", "Camk1", "Hs3st5", "Syn3", "Pde4b", "Dlg2", 
             "Ece2", "Slc22a4", "Pola2", "Dgki")

# get DEG files
files <- list.files(paste0(out, "DEGs_with_hbb/DEG_tables/"))
files <- files[grep("E.21D.+_vs_S.21D.+_1.00_LFC_0.00.tsv", files)]

# init
lfc.df <- data.frame(row.names = goi_21D)
adjpval.df <- data.frame(row.names = goi_21D)

for (i in seq_along(files)) {
  # read table
  data <- read.delim(
    file = paste0(out, "DEGs_with_hbb/DEG_tables/", files[i]),
    sep = "\t", check.names = FALSE
  )

  # comparison name
  comparison <- sub("_FDRq_1.00_LFC_0.00\\.tsv$", "", files[i])

  # filter + order to goi_2D
  data <- data[data$gene_name %in% goi_21D, c("gene_name","logFC", "adj.P.Val")]
  rownames(data) <- data$gene_name
  data <- data[match(goi_21D, rownames(data)), , drop = FALSE]

  # ensure numeric (in case it was read as character)
  data$logFC <- round(as.numeric(data$logFC), 4)
  data$adj.P.Val <- as.numeric(data$adj.P.Val)

  # assign just the logFC column
  lfc.df[[comparison]] <- data$logFC
  adjpval.df[[comparison]] <- data$adj.P.Val
}

# set heatmap colors and names
paletteLength <- 100
myColor <- colorRampPalette(c("cornflowerblue", "white", "red"))(paletteLength)
myBreaks <- c(seq(min(lfc.df), 0, length.out = ceiling(paletteLength/2) + 1), 
              seq(max(lfc.df)/paletteLength, max(lfc.df), length.out = floor(paletteLength/2)))

# save
pdf(paste0(out, "custom_figures/21D_logFC_adjPVal_heatmap.pdf"), width = 5, height = 10)


pheatmap::pheatmap(lfc.df,
                   main = "logFC and adj.P.Val for 21 Day Time Points",
                   cluster_rows = FALSE,
                   cluster_cols = FALSE,
                   annotation_colors = ann_colors,
                   color = myColor,
                   breaks = myBreaks,
                   display_numbers = round(adjpval.df, digits = 4),
                   fontsize_number = 12,
                   number_color = "black",
                   )
```


## 21 days w/out pval
```{r}
# save path
out <- "../results/both_sexes/"

# genes of interest
goi_21D <- c("Hif3a", "Tlr13", "Xdh", "Skap1", "Ccdc88a", "Kcnb2", "Grik1", 
             "Adcy8", "H2-K1", "Higd1b", "Atp10a", "C1qtnf4", "Ppp3r1", "Gab3", 
             "Il1rapl1", "Rasgrp2", "Camk1", "Hs3st5", "Syn3", "Pde4b", "Dlg2", 
             "Ece2", "Slc22a4", "Pola2", "Dgki")

# get DEG files
files <- list.files(paste0(out, "DEGs_with_hbb/DEG_tables/"))
files <- files[grep("E.21D.+_vs_S.21D.+_1.00_LFC_0.00.tsv", files)]

# init
lfc.df <- data.frame(row.names = goi_21D)
adjpval.df <- data.frame(row.names = goi_21D)

for (i in seq_along(files)) {
  # read table
  data <- read.delim(
    file = paste0(out, "DEGs_with_hbb/DEG_tables/", files[i]),
    sep = "\t", check.names = FALSE
  )

  # comparison name
  comparison <- sub("_FDRq_1.00_LFC_0.00\\.tsv$", "", files[i])

  # filter + order to goi_2D
  data <- data[data$gene_name %in% goi_21D, c("gene_name","logFC", "adj.P.Val")]
  rownames(data) <- data$gene_name
  data <- data[match(goi_21D, rownames(data)), , drop = FALSE]

  # ensure numeric (in case it was read as character)
  data$logFC <- round(as.numeric(data$logFC), 4)
  data$adj.P.Val <- as.numeric(data$adj.P.Val)

  # assign just the logFC column
  lfc.df[[comparison]] <- data$logFC
  adjpval.df[[comparison]] <- data$adj.P.Val
}

# set heatmap colors and names
paletteLength <- 100
myColor <- colorRampPalette(c("cornflowerblue", "white", "red"))(paletteLength)
myBreaks <- c(seq(min(lfc.df), 0, length.out = ceiling(paletteLength/2) + 1), 
              seq(max(lfc.df)/paletteLength, max(lfc.df), length.out = floor(paletteLength/2)))

# save
pdf(paste0(out, "custom_figures/21D_logFC_heatmap.pdf"), width = 5, height = 10)


pheatmap::pheatmap(lfc.df,
                   main = "logFC for 21 Day Time Points",
                   cluster_rows = FALSE,
                   cluster_cols = FALSE,
                   annotation_colors = ann_colors,
                   color = myColor,
                   breaks = myBreaks)
```


# Upset filtering comparison
## E.coli vs saline, 2 days, both sexes
### Plot
```{r}
# set variables
thresh <- 0.05
myLfc <- 0
model <- "DEGs_with_hbb/"

# read tables
E2DY_vs_S2DY <- read.delim(
  paste0("../results/both_sexes/", 
         model, "DEG_tables/E.2D.Y_vs_S.2D.Y_FDRq_1.00_LFC_0.00.tsv"))
E2DO_vs_S2DO <- read.delim(
  paste0("../results/filtered_samples_both_sexes/", 
         model, "DEG_tables/E.2D.O_vs_S.2D.O_FDRq_1.00_LFC_0.00.tsv"))
rownames(E2DY_vs_S2DY) <- E2DY_vs_S2DY$gene_name_unique
rownames(E2DO_vs_S2DO) <- E2DO_vs_S2DO$gene_name_unique

# filter on pval
E2DY_vs_S2DY <- E2DY_vs_S2DY[E2DY_vs_S2DY$adj.P.Val < thresh,]
E2DO_vs_S2DO <- E2DO_vs_S2DO[E2DO_vs_S2DO$adj.P.Val < thresh,]

# Subset lists
E2DY_vs_S2DY_up <- subset(E2DY_vs_S2DY$gene_name_unique, E2DY_vs_S2DY$logFC > myLfc)
E2DY_vs_S2DY_down <- subset(E2DY_vs_S2DY$gene_name_unique, E2DY_vs_S2DY$logFC < -myLfc)
E2DO_vs_S2DO_up <- subset(E2DO_vs_S2DO$gene_name_unique, E2DO_vs_S2DO$logFC > myLfc)
E2DO_vs_S2DO_down <- subset(E2DO_vs_S2DO$gene_name_unique, E2DO_vs_S2DO$logFC < -myLfc)
list_input <- list("E.2D.O vs S.2D.O up-regulated" = E2DO_vs_S2DO_up,
                   "E.2D.Y vs S.2D.Y up-regulated" = E2DY_vs_S2DY_up,
                   "E.2D.O vs S.2D.O down-regulated" = E2DO_vs_S2DO_down,
                   "E.2D.Y vs S.2D.Y down-regulated" = E2DY_vs_S2DY_down)
data <- fromList(list_input)
  
# store names
names <- colnames(data)
  
# plot
upset_gene <- ComplexUpset::upset(data, 
				  names,
				  set_sizes=(
					upset_set_size()
					+ geom_text(aes(label=..count..), hjust=1.1, stat='count')
					+ expand_limits(y=400)),
				  queries = list(upset_query("E.2D.O vs S.2D.O up-regulated", fill = "red"),
								 upset_query("E.2D.Y vs S.2D.Y up-regulated", fill = "red"),
								 upset_query("E.2D.O vs S.2D.O down-regulated", fill = "blue"),
								 upset_query("E.2D.Y vs S.2D.Y down-regulated", fill = "blue")),
				  base_annotations = list('Intersection size' = (
					intersection_size(bar_number_threshold=1, width=0.5)
					+ scale_y_continuous(expand=expansion(mult=c(0, 0.05)),limits = c(0,1000)) # space on top
					+ theme(
						  # hide grid lines
						  panel.grid.major=element_blank(),
						  panel.grid.minor=element_blank(),
						  # show axis lines
						  axis.line=element_line(colour='black')))),
				  stripes = upset_stripes(
					geom=geom_segment(size=12),  # make the stripes larger
					colors=c('grey95', 'white')),
				  # to prevent connectors from getting the colorured
				  # use `fill` instead of `color`, together with `shape='circle filled'`
				  matrix = intersection_matrix(
					geom=geom_point(
					  shape='circle filled',
					  size=3,
					  stroke=0.45)),
				  sort_sets=FALSE,
				  sort_intersections='descending'
				)

upset_gene <- upset_gene + ggtitle(paste0("E.2D vs S.2D, FDRq < ", format(thresh, nsmall = 2),
                                          ", |LFC| > ", format(myLfc, nsmall = 2)))

# save
pdf(paste0("../results/both_sexes/",
           model, "upset/E.2D_vs_S.2D_both_sexes_FDRq_", 
           format(thresh, nsmall = 2), "_LFC_", format(myLfc, nsmall = 2), "_upset.pdf"), 
    height = 6, width = 10)
upset_gene
```

### Extract
```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = paste0("../results/both_sexes/",
                          model, "upset/E.2D_vs_S.2D_both_sexes_FDRq_", 
                          format(thresh, nsmall = 2), "_LFC_", format(myLfc, nsmall = 2), 
                          "_results.tsv"),
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

## E.coli vs saline, 21 days, both sexes
### Plot
```{r}
# set variables
thresh <- 0.05
myLfc <- 0
model <- "DEGs_with_hbb/"

# read tables
E21DY_vs_S21DY <- read.delim(
  paste0("../results/both_sexes/", 
         model, "DEG_tables/E.21D.Y_vs_S.21D.Y_FDRq_1.00_LFC_0.00.tsv"))
E21DO_vs_S21DO <- read.delim(
  paste0("../results/filtered_samples_both_sexes/", 
         model, "DEG_tables/E.21D.O_vs_S.21D.O_FDRq_1.00_LFC_0.00.tsv"))
rownames(E21DY_vs_S21DY) <- E21DY_vs_S21DY$gene_name_unique
rownames(E21DO_vs_S21DO) <- E21DO_vs_S21DO$gene_name_unique

# filter on pval
E21DY_vs_S21DY <- E21DY_vs_S21DY[E21DY_vs_S21DY$adj.P.Val < thresh,]
E21DO_vs_S21DO <- E21DO_vs_S21DO[E21DO_vs_S21DO$adj.P.Val < thresh,]

# Subset lists
E21DY_vs_S21DY_up <- subset(E21DY_vs_S21DY$gene_name_unique, E21DY_vs_S21DY$logFC > myLfc)
E21DY_vs_S21DY_down <- subset(E21DY_vs_S21DY$gene_name_unique, E21DY_vs_S21DY$logFC < -myLfc)
E21DO_vs_S21DO_up <- subset(E21DO_vs_S21DO$gene_name_unique, E21DO_vs_S21DO$logFC > myLfc)
E21DO_vs_S21DO_down <- subset(E21DO_vs_S21DO$gene_name_unique, E21DO_vs_S21DO$logFC < -myLfc)
list_input <- list("E.21D.O vs S.21D.O up-regulated" = E21DO_vs_S21DO_up,
                   "E.21D.Y vs S.21D.Y up-regulated" = E21DY_vs_S21DY_up,
                   "E.21D.O vs S.21D.O down-regulated" = E21DO_vs_S21DO_down,
                   "E.21D.Y vs S.21D.Y down-regulated" = E21DY_vs_S21DY_down)
data <- fromList(list_input)
  
# store names
names <- colnames(data)
  
# plot
upset_gene <- ComplexUpset::upset(data, 
				  names,
				  set_sizes=(
					upset_set_size()
					+ geom_text(aes(label=..count..), hjust=1.1, stat='count')
					+ expand_limits(y=400)),
				  queries = list(upset_query("E.21D.O vs S.21D.O up-regulated", fill = "red"),
								 upset_query("E.21D.Y vs S.21D.Y up-regulated", fill = "red"),
								 upset_query("E.21D.O vs S.21D.O down-regulated", fill = "blue"),
								 upset_query("E.21D.Y vs S.21D.Y down-regulated", fill = "blue")),
				  base_annotations = list('Intersection size' = (
					intersection_size(bar_number_threshold=1, width=0.5)
					+ scale_y_continuous(expand=expansion(mult=c(0, 0.05)),limits = c(0,1000)) # space on top
					+ theme(
						  # hide grid lines
						  panel.grid.major=element_blank(),
						  panel.grid.minor=element_blank(),
						  # show axis lines
						  axis.line=element_line(colour='black')))),
				  stripes = upset_stripes(
					geom=geom_segment(size=12),  # make the stripes larger
					colors=c('grey95', 'white')),
				  # to prevent connectors from getting the colorured
				  # use `fill` instead of `color`, together with `shape='circle filled'`
				  matrix = intersection_matrix(
					geom=geom_point(
					  shape='circle filled',
					  size=3,
					  stroke=0.45)),
				  sort_sets=FALSE,
				  sort_intersections='descending'
				)

upset_gene <- upset_gene + ggtitle(paste0("E.21D vs S.21D, FDRq < ", format(thresh, nsmall = 2),
                                          ", |LFC| > ", format(myLfc, nsmall = 2)))

# save
pdf(paste0("../results/both_sexes/",
           model, "upset/E.21D_vs_S.21D_both_sexes_FDRq_", 
           format(thresh, nsmall = 2), "_LFC_", format(myLfc, nsmall = 2), "_upset.pdf"), 
    height = 6, width = 10)
upset_gene
```

### Extract
```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = paste0("../results/both_sexes/",
                          model, "upset/E.21D_vs_S.21D_both_sexes_FDRq_", 
                          format(thresh, nsmall = 2), "_LFC_", format(myLfc, nsmall = 2), 
                          "_results.tsv"),
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```


## E.coli vs saline, both sexes
### Plot
```{r}
# set variables
thresh <- 0.05
myLfc <- 0
model <- "DEGs_with_hbb/"

# read tables
E2DY_vs_S2DY <- read.delim(
  paste0("../results/both_sexes/", 
         model, "DEG_tables/E.2D.Y_vs_S.2D.Y_FDRq_1.00_LFC_0.00.tsv"))
E2DO_vs_S2DO <- read.delim(
  paste0("../results/filtered_samples_both_sexes/", 
         model, "DEG_tables/E.2D.O_vs_S.2D.O_FDRq_1.00_LFC_0.00.tsv"))
E21DY_vs_S21DY <- read.delim(
  paste0("../results/both_sexes/", 
         model, "DEG_tables/E.21D.Y_vs_S.21D.Y_FDRq_1.00_LFC_0.00.tsv"))
E21DO_vs_S21DO <- read.delim(
  paste0("../results/filtered_samples_both_sexes/", 
         model, "DEG_tables/E.21D.O_vs_S.21D.O_FDRq_1.00_LFC_0.00.tsv"))

rownames(E2DY_vs_S2DY) <- E2DY_vs_S2DY$gene_name_unique
rownames(E2DO_vs_S2DO) <- E2DO_vs_S2DO$gene_name_unique
rownames(E21DY_vs_S21DY) <- E21DY_vs_S21DY$gene_name_unique
rownames(E21DO_vs_S21DO) <- E21DO_vs_S21DO$gene_name_unique

# filter on pval
E2DY_vs_S2DY <- E2DY_vs_S2DY[E2DY_vs_S2DY$adj.P.Val < thresh,]
E2DO_vs_S2DO <- E2DO_vs_S2DO[E2DO_vs_S2DO$adj.P.Val < thresh,]
E21DY_vs_S21DY <- E21DY_vs_S21DY[E21DY_vs_S21DY$adj.P.Val < thresh,]
E21DO_vs_S21DO <- E21DO_vs_S21DO[E21DO_vs_S21DO$adj.P.Val < thresh,]

# Subset lists
E2DY_vs_S2DY_up <- subset(E2DY_vs_S2DY$gene_name_unique, E2DY_vs_S2DY$logFC > myLfc)
E2DY_vs_S2DY_down <- subset(E2DY_vs_S2DY$gene_name_unique, E2DY_vs_S2DY$logFC < -myLfc)
E2DO_vs_S2DO_up <- subset(E2DO_vs_S2DO$gene_name_unique, E2DO_vs_S2DO$logFC > myLfc)
E2DO_vs_S2DO_down <- subset(E2DO_vs_S2DO$gene_name_unique, E2DO_vs_S2DO$logFC < -myLfc)
E21DY_vs_S21DY_up <- subset(E21DY_vs_S21DY$gene_name_unique, E21DY_vs_S21DY$logFC > myLfc)
E21DY_vs_S21DY_down <- subset(E21DY_vs_S21DY$gene_name_unique, E21DY_vs_S21DY$logFC < -myLfc)
E21DO_vs_S21DO_up <- subset(E21DO_vs_S21DO$gene_name_unique, E21DO_vs_S21DO$logFC > myLfc)
E21DO_vs_S21DO_down <- subset(E21DO_vs_S21DO$gene_name_unique, E21DO_vs_S21DO$logFC < -myLfc)
list_input <- list("E.21D.O vs S.21D.O up-regulated" = E21DO_vs_S21DO_up,
                   "E.21D.Y vs S.21D.Y up-regulated" = E21DY_vs_S21DY_up,
                   "E.21D.O vs S.21D.O down-regulated" = E21DO_vs_S21DO_down,
                   "E.21D.Y vs S.21D.Y down-regulated" = E21DY_vs_S21DY_down,
                   "E.2D.O vs S.2D.O up-regulated" = E2DO_vs_S2DO_up,
                   "E.2D.Y vs S.2D.Y up-regulated" = E2DY_vs_S2DY_up,
                   "E.2D.O vs S.2D.O down-regulated" = E2DO_vs_S2DO_down,
                   "E.2D.Y vs S.2D.Y down-regulated" = E2DY_vs_S2DY_down)
data <- fromList(list_input)
  
# store names
names <- colnames(data)
  
# plot
upset_gene <- ComplexUpset::upset(data, 
				  names,
				  set_sizes=(
					upset_set_size()
					+ geom_text(aes(label=..count..), hjust=1.1, stat='count')
					+ expand_limits(y=400)),
				  queries = list(upset_query("E.21D.O vs S.21D.O up-regulated", fill = "red"),
								 upset_query("E.21D.Y vs S.21D.Y up-regulated", fill = "red"),
								 upset_query("E.21D.O vs S.21D.O down-regulated", fill = "blue"),
								 upset_query("E.21D.Y vs S.21D.Y down-regulated", fill = "blue"),
								 upset_query("E.2D.O vs S.2D.O up-regulated", fill = "red"),
								 upset_query("E.2D.Y vs S.2D.Y up-regulated", fill = "red"),
								 upset_query("E.2D.O vs S.2D.O down-regulated", fill = "blue"),
								 upset_query("E.2D.Y vs S.2D.Y down-regulated", fill = "blue")),
				  base_annotations = list('Intersection size' = (
					intersection_size(bar_number_threshold=1, width=0.5)
					+ scale_y_continuous(expand=expansion(mult=c(0, 0.05)),limits = c(0,250)) # space on top
					+ theme(
						  # hide grid lines
						  panel.grid.major=element_blank(),
						  panel.grid.minor=element_blank(),
						  # show axis lines
						  axis.line=element_line(colour='black')))),
				  stripes = upset_stripes(
					geom=geom_segment(size=12),  # make the stripes larger
					colors=c('grey95', 'white')),
				  # to prevent connectors from getting the colorured
				  # use `fill` instead of `color`, together with `shape='circle filled'`
				  matrix = intersection_matrix(
					geom=geom_point(
					  shape='circle filled',
					  size=3,
					  stroke=0.45)),
				  sort_sets=FALSE,
				  sort_intersections='descending'
				)

upset_gene <- upset_gene + ggtitle(paste0("E. coli vs Saline, FDRq < ", format(thresh, nsmall = 2),
                                          ", |LFC| > ", format(myLfc, nsmall = 2)))

# save
pdf(paste0("../results/both_sexes/",
           model, "upset/E_vs_S_both_sexes_FDRq_", 
           format(thresh, nsmall = 2), "_LFC_", format(myLfc, nsmall = 2), "_upset.pdf"), 
    height = 6, width = 10)
upset_gene
```

### Extract
```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = paste0("../results/both_sexes/",
                          model, "upset/E.2D_vs_S.2D_both_sexes_FDRq_", 
                          format(thresh, nsmall = 2), "_LFC_", format(myLfc, nsmall = 2), 
                          "_results.tsv"),
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```
## E.2D.O vs S.2D.O
### Plot
```{r}
# set variables
thresh <- 0.05
myLfc <- 0
model <- "DEGs_with_hbb/"

# read tables
ss_filt <- read.delim(
  paste0("../results/both_sexes/", 
         model, "DEG_tables/E.2D.O_vs_S.2D.O_FDRq_1.00_LFC_0.00.tsv"))
bs_filt <- read.delim(
  paste0("../results/filtered_samples_both_sexes/", 
         model, "DEG_tables/E.2D.O_vs_S.2D.O_FDRq_1.00_LFC_0.00.tsv"))
rownames(ss_filt) <- ss_filt$gene_name_unique
rownames(bs_filt) <- bs_filt$gene_name_unique

# filter on pval
ss_filt <- ss_filt[ss_filt$adj.P.Val < thresh,]
bs_filt <- bs_filt[bs_filt$adj.P.Val < thresh,]

# Subset lists
ss_filt.up <- subset(ss_filt$gene_name_unique, ss_filt$logFC > myLfc)
ss_filt.down <- subset(ss_filt$gene_name_unique, ss_filt$logFC < -myLfc)
bs_filt.up <- subset(bs_filt$gene_name_unique, bs_filt$logFC > myLfc)
bs_filt.down <- subset(bs_filt$gene_name_unique, bs_filt$logFC < -myLfc)
list_input <- list("both sexes filtering, up-regulated" = bs_filt.up,
                   "sex specific filtering, up-regulated" = ss_filt.up,
                   "both sexes filtering, down-regulated" = bs_filt.down,
                   "sex specific filtering, down-regulated" = ss_filt.down)
data <- fromList(list_input)
  
# store names
names <- colnames(data)
  
# plot
upset_gene <- ComplexUpset::upset(data, 
				  names,
				  set_sizes=(
					upset_set_size()
					+ geom_text(aes(label=..count..), hjust=1.1, stat='count')
					+ expand_limits(y=400)),
				  queries = list(upset_query("both sexes filtering, up-regulated", fill = "red"),
								 upset_query("sex specific filtering, up-regulated", fill = "red"),
								 upset_query("8 hr Low Dose Psilocybin vs Saline Down-regulated", fill = "blue"),
								 upset_query("both sexes filtering, down-regulated", fill = "blue"),
								 upset_query("sex specific filtering, down-regulated", fill = "blue")),
				  base_annotations = list('Intersection size' = (
					intersection_size(bar_number_threshold=1, width=0.5)
					+ scale_y_continuous(expand=expansion(mult=c(0, 0.05)),limits = c(0,350)) # space on top
					+ theme(
						  # hide grid lines
						  panel.grid.major=element_blank(),
						  panel.grid.minor=element_blank(),
						  # show axis lines
						  axis.line=element_line(colour='black')))),
				  stripes = upset_stripes(
					geom=geom_segment(size=12),  # make the stripes larger
					colors=c('grey95', 'white')),
				  # to prevent connectors from getting the colorured
				  # use `fill` instead of `color`, together with `shape='circle filled'`
				  matrix = intersection_matrix(
					geom=geom_point(
					  shape='circle filled',
					  size=3,
					  stroke=0.45)),
				  sort_sets=FALSE,
				  sort_intersections='descending'
				)

upset_gene <- upset_gene + ggtitle(paste0("E.2D.O vs S.2D.O, FDRq < ", format(thresh, nsmall = 2),
                                          ", |LFC| > ", format(myLfc, nsmall = 2)))

# save
pdf(paste0("../results/filtering_comparison/E.2D.O_vs_S.2D.O_FDRq_", format(thresh, nsmall = 2), 
           "_LFC_", format(myLfc, nsmall = 2), "_upset.pdf"), 
    height = 6, width = 10)
upset_gene
```

### Extract
```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = paste0("../results/filtering_comparison/E.2D.O_vs_S.2D.O_FDRq_", 
                          format(thresh, nsmall = 2), "_LFC_", format(myLfc, nsmall = 2), 
                          "_results.tsv"),
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```


## E.21D.Y vs S.21D.Y
### Plot
```{r}
# set variables
thresh <- 0.05
myLfc <- 0
model <- "DEGs_with_hbb/"

# read tables
ss_filt <- read.delim(
  paste0("../results/both_sexes/", 
         model, "DEG_tables/E.21D.Y_vs_S.21D.Y_FDRq_1.00_LFC_0.00.tsv"))
bs_filt <- read.delim(
  paste0("../results/filtered_samples_both_sexes/", 
         model, "DEG_tables/E.21D.Y_vs_S.21D.Y_FDRq_1.00_LFC_0.00.tsv"))
rownames(ss_filt) <- ss_filt$gene_name_unique
rownames(bs_filt) <- bs_filt$gene_name_unique

# filter on pval
ss_filt <- ss_filt[ss_filt$adj.P.Val < thresh,]
bs_filt <- bs_filt[bs_filt$adj.P.Val < thresh,]

# Subset lists
ss_filt.up <- subset(ss_filt$gene_name_unique, ss_filt$logFC > myLfc)
ss_filt.down <- subset(ss_filt$gene_name_unique, ss_filt$logFC < -myLfc)
bs_filt.up <- subset(bs_filt$gene_name_unique, bs_filt$logFC > myLfc)
bs_filt.down <- subset(bs_filt$gene_name_unique, bs_filt$logFC < -myLfc)
list_input <- list("both sexes filtering, up-regulated" = bs_filt.up,
                   "sex specific filtering, up-regulated" = ss_filt.up,
                   "both sexes filtering, down-regulated" = bs_filt.down,
                   "sex specific filtering, down-regulated" = ss_filt.down)
data <- fromList(list_input)
  
# store names
names <- colnames(data)
  
# plot
upset_gene <- ComplexUpset::upset(data, 
				  names,
				  set_sizes=(
					upset_set_size()
					+ geom_text(aes(label=..count..), hjust=1.1, stat='count')
					+ expand_limits(y=400)),
				  queries = list(upset_query("both sexes filtering, up-regulated", fill = "red"),
								 upset_query("sex specific filtering, up-regulated", fill = "red"),
								 upset_query("8 hr Low Dose Psilocybin vs Saline Down-regulated", fill = "blue"),
								 upset_query("both sexes filtering, down-regulated", fill = "blue"),
								 upset_query("sex specific filtering, down-regulated", fill = "blue")),
				  base_annotations = list('Intersection size' = (
					intersection_size(bar_number_threshold=1, width=0.5)
					+ scale_y_continuous(expand=expansion(mult=c(0, 0.05)),limits = c(0,350)) # space on top
					+ theme(
						  # hide grid lines
						  panel.grid.major=element_blank(),
						  panel.grid.minor=element_blank(),
						  # show axis lines
						  axis.line=element_line(colour='black')))),
				  stripes = upset_stripes(
					geom=geom_segment(size=12),  # make the stripes larger
					colors=c('grey95', 'white')),
				  # to prevent connectors from getting the colorured
				  # use `fill` instead of `color`, together with `shape='circle filled'`
				  matrix = intersection_matrix(
					geom=geom_point(
					  shape='circle filled',
					  size=3,
					  stroke=0.45)),
				  sort_sets=FALSE,
				  sort_intersections='descending'
				)

upset_gene <- upset_gene + ggtitle(paste0("E.21D.Y vs S.21D.Y, FDRq < ", format(thresh, nsmall = 2),
                                          ", |LFC| > ", format(myLfc, nsmall = 2)))

# save
pdf(paste0("../results/filtering_comparison/E.21D.Y_vs_S.21D.Y_FDRq_", format(thresh, nsmall = 2), 
           "_LFC_", format(myLfc, nsmall = 2), "_upset.pdf"), 
    height = 6, width = 10)
upset_gene
```

### Extract
```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = paste0("../results/filtering_comparison/E.21D.Y_vs_S.21D.Y_FDRq_", 
                          format(thresh, nsmall = 2), "_LFC_", format(myLfc, nsmall = 2), 
                          "_results.tsv"),
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```


## E.21D.O vs S.21D.O
### Plot
```{r}
# set variables
thresh <- 0.05
myLfc <- 0
model <- "DEGs_with_hbb/"

# read tables
ss_filt <- read.delim(
  paste0("../results/both_sexes/", 
         model, "DEG_tables/E.21D.O_vs_S.21D.O_FDRq_1.00_LFC_0.00.tsv"))
bs_filt <- read.delim(
  paste0("../results/filtered_samples_both_sexes/", 
         model, "DEG_tables/E.21D.O_vs_S.21D.O_FDRq_1.00_LFC_0.00.tsv"))
rownames(ss_filt) <- ss_filt$gene_name_unique
rownames(bs_filt) <- bs_filt$gene_name_unique

# filter on pval
ss_filt <- ss_filt[ss_filt$adj.P.Val < thresh,]
bs_filt <- bs_filt[bs_filt$adj.P.Val < thresh,]

# Subset lists
ss_filt.up <- subset(ss_filt$gene_name_unique, ss_filt$logFC > myLfc)
ss_filt.down <- subset(ss_filt$gene_name_unique, ss_filt$logFC < -myLfc)
bs_filt.up <- subset(bs_filt$gene_name_unique, bs_filt$logFC > myLfc)
bs_filt.down <- subset(bs_filt$gene_name_unique, bs_filt$logFC < -myLfc)
list_input <- list("both sexes filtering, up-regulated" = bs_filt.up,
                   "sex specific filtering, up-regulated" = ss_filt.up,
                   "both sexes filtering, down-regulated" = bs_filt.down,
                   "sex specific filtering, down-regulated" = ss_filt.down)
data <- fromList(list_input)
  
# store names
names <- colnames(data)
  
# plot
upset_gene <- ComplexUpset::upset(data, 
				  names,
				  set_sizes=(
					upset_set_size()
					+ geom_text(aes(label=..count..), hjust=1.1, stat='count')
					+ expand_limits(y=400)),
				  queries = list(upset_query("both sexes filtering, up-regulated", fill = "red"),
								 upset_query("sex specific filtering, up-regulated", fill = "red"),
								 upset_query("8 hr Low Dose Psilocybin vs Saline Down-regulated", fill = "blue"),
								 upset_query("both sexes filtering, down-regulated", fill = "blue"),
								 upset_query("sex specific filtering, down-regulated", fill = "blue")),
				  base_annotations = list('Intersection size' = (
					intersection_size(bar_number_threshold=1, width=0.5)
					+ scale_y_continuous(expand=expansion(mult=c(0, 0.05)),limits = c(0,350)) # space on top
					+ theme(
						  # hide grid lines
						  panel.grid.major=element_blank(),
						  panel.grid.minor=element_blank(),
						  # show axis lines
						  axis.line=element_line(colour='black')))),
				  stripes = upset_stripes(
					geom=geom_segment(size=12),  # make the stripes larger
					colors=c('grey95', 'white')),
				  # to prevent connectors from getting the colorured
				  # use `fill` instead of `color`, together with `shape='circle filled'`
				  matrix = intersection_matrix(
					geom=geom_point(
					  shape='circle filled',
					  size=3,
					  stroke=0.45)),
				  sort_sets=FALSE,
				  sort_intersections='descending'
				)

upset_gene <- upset_gene + ggtitle(paste0("E.21D.O vs S.21D.O, FDRq < ", format(thresh, nsmall = 2),
                                          ", |LFC| > ", format(myLfc, nsmall = 2)))

# save
pdf(paste0("../results/filtering_comparison/E.21D.O_vs_S.21D.O_FDRq_", format(thresh, nsmall = 2), 
           "_LFC_", format(myLfc, nsmall = 2), "_upset.pdf"), 
    height = 6, width = 10)
upset_gene
```

### Extract
```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = paste0("../results/filtering_comparison/E.21D.O_vs_S.21D.O_FDRq_", 
                          format(thresh, nsmall = 2), "_LFC_", format(myLfc, nsmall = 2), 
                          "_results.tsv"),
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

# Custom volcanos
## E.2D.Y vs S.2D.Y
```{r}
# custom genes to label
goi <- c("S100a8", "S100a9", "Lcn2", "Mpeg1", "Csf3r", "Adgre1", "Vim", "Pecam1",
         "Itgam", "Sgk1", "Nfkbia", "Acot11", "Smox", "Mfsd2a")

# Output directories and file paths
out <- "../results/both_sexes/"
model <- "DEGs_with_hbb/"

# set variables
thresh <- 0.05
my.lfc <- 0

# get file list
file <- paste0(out, model, "DEG_tables/E.2D.Y_vs_S.2D.Y_FDRq_1.00_LFC_0.00.tsv")

# read DEG file
data <- read.delim(file, sep = "\t")

# subset genes to label
labels <- data[data$gene_name %in% goi,]
up <- labels[labels$logFC > 0,]
down <- labels[labels$logFC < 0,]
  
# assign colors based on thresh
color_values <- vector()
max <- nrow(data)
for(row in 1:max){
  if (data$adj.P.Val[row] < thresh){
    if (data$logFC [row] > my.lfc){
      color_values <- c(color_values, 1) # 1 when logFC > my.lfc and FDRq < thresh
    } else if (data$logFC[row] < -my.lfc){
      color_values <- c(color_values, 2) # 2 when logFC < -my.lfc and FDRq < thresh
    } else {
      color_values <- c(color_values, 3) # 3 when logFC between -my.lfc and +my.lfc and FDRq < thresh
    }
  } else {
    color_values <- c(color_values, 3) # 3 when FDRq >= thresh
  }
}
data$color_adjpval <- color_values

# manual colors for goi
data$color_adjpval[data$gene_name %in% up$gene_name] <- 4
data$color_adjpval[data$gene_name %in% down$gene_name] <- 5
data$color_adjpval <- factor(data$color_adjpval)
  
# comparison name
comparison <- gsub(paste0(out, model, "DEG_tables/"), "", file)
comparison <- gsub("_FDRq_1.00_LFC_0.00.tsv","",comparison)

# set manual colors
my_colors <- c("#ffdcdb","#e6eefc","lightgray","red","blue")
 
# set significance threshold
hadjpval <- (-log10(max(data$P.Value[data$adj.P.Val < thresh], na.rm=TRUE)))

# plot
p <-
  ggplot(data = data, 
         aes(x = logFC,  # x-axis is logFC
             y = -log10(P.Value),  # y-axis will be -log10 of P.Value
             color = color_adjpval)) +  # color is based on factored color column
  geom_point(alpha = 0.8, size = 2) +  # create scatterplot, alpha makes points transparent
  theme_bw() +  # set color theme
  theme(legend.position = "none") +  # no legend
  scale_color_manual(values = my_colors) +  # set factor colors
  labs(
    title = "", # no main title
    x = expression(log[2](FC)), # x-axis title
     y = expression(-log[10] ~ "(" ~ italic("p") ~ "-value)") # y-axis title
  ) +
  theme(axis.title.x = element_text(size = 15),
        axis.text.x = element_text(size = 15)) +
  theme(axis.title.y = element_text(size = 15),
        axis.text.y = element_text(size = 15)) +
  theme(plot.title = element_text(size = 15)) +
  geom_hline(yintercept = hadjpval,  #  horizontal line
                     colour = "#000000",
                     linetype = "dashed") +
  geom_vline(xintercept = -my.lfc,  #  vertical line
                     colour = "#000000",
                     linetype = "dashed") +
  geom_vline(xintercept = my.lfc,  #  vertical line
                     colour = "#000000",
                     linetype = "dashed") +
  ggtitle(paste0(comparison, ", adj.P.Val < ", thresh, ", LFC = ", my.lfc)) +
  geom_text_repel(data = up,
                  aes(x = logFC, y= -log10(P.Value), label = gene_name), 
                  size = 5,
                  color = "maroon", 
                  fontface="italic",
                  max.overlaps = getOption("ggrepel.max.overlaps", default = 30)
                  ) +
  geom_text_repel(data = down,
                  aes(x = logFC, y= -log10(P.Value), label = gene_name), 
                  color = "navyblue", 
                  size = 5,
                  fontface="italic",
                  max.overlaps = getOption("ggrepel.max.overlaps", default = 30)
                  ) +
  ylim(c(0,15))
 p
  
# save
path <- paste0(out, "custom_figures/", comparison, "_FDRq_", 
               format(thresh, nsmall = 2), "_LFC_",
               format(my.lfc, nsmall = 2), "_volcano.pdf")
pdf(path, height = 8, width = 8)
p
dev.off()
```

## E.2D.O vs S.2D.O
```{r}
# custom genes to label
goi <- c("S100a8", "S100a9", "Lcn2", "Mpeg1", "Csf3r", "Adgre1", "C1qa", "C1qb", 
         "C1qc", "Top2a", "Tnfsf10", "Sqle", "Hmgcr", "Insig1")

# Output directories and file paths
out <- "../results/both_sexes/"
model <- "DEGs_with_hbb/"

# set variables
thresh <- 0.05
my.lfc <- 0

# get file list
file <- paste0(out, model, "DEG_tables/E.2D.O_vs_S.2D.O_FDRq_1.00_LFC_0.00.tsv")

# read DEG file
data <- read.delim(file, sep = "\t")

# subset genes to label
labels <- data[data$gene_name %in% goi,]
up <- labels[labels$logFC > 0,]
down <- labels[labels$logFC < 0,]
  
# assign colors based on thresh
color_values <- vector()
max <- nrow(data)
for(row in 1:max){
  if (data$adj.P.Val[row] < thresh){
    if (data$logFC [row] > my.lfc){
      color_values <- c(color_values, 1) # 1 when logFC > my.lfc and FDRq < thresh
    } else if (data$logFC[row] < -my.lfc){
      color_values <- c(color_values, 2) # 2 when logFC < -my.lfc and FDRq < thresh
    } else {
      color_values <- c(color_values, 3) # 3 when logFC between -my.lfc and +my.lfc and FDRq < thresh
    }
  } else {
    color_values <- c(color_values, 3) # 3 when FDRq >= thresh
  }
}
data$color_adjpval <- color_values

# manual colors for goi
data$color_adjpval[data$gene_name %in% up$gene_name] <- 4
data$color_adjpval[data$gene_name %in% down$gene_name] <- 5
data$color_adjpval <- factor(data$color_adjpval)
  
# comparison name
comparison <- gsub(paste0(out, model, "DEG_tables/"), "", file)
comparison <- gsub("_FDRq_1.00_LFC_0.00.tsv","",comparison)

# set manual colors
my_colors <- c("#ffdcdb","#e6eefc","lightgray","red","blue")
 
# set significance threshold
hadjpval <- (-log10(max(data$P.Value[data$adj.P.Val < thresh], na.rm=TRUE)))

# plot
p <-
  ggplot(data = data, 
         aes(x = logFC,  # x-axis is logFC
             y = -log10(P.Value),  # y-axis will be -log10 of P.Value
             color = color_adjpval)) +  # color is based on factored color column
  geom_point(alpha = 0.8, size = 2) +  # create scatterplot, alpha makes points transparent
  theme_bw() +  # set color theme
  theme(legend.position = "none") +  # no legend
  scale_color_manual(values = my_colors) +  # set factor colors
  labs(
    title = "", # no main title
    x = expression(log[2](FC)), # x-axis title
     y = expression(-log[10] ~ "(" ~ italic("p") ~ "-value)") # y-axis title
  ) +
  theme(axis.title.x = element_text(size = 15),
        axis.text.x = element_text(size = 15)) +
  theme(axis.title.y = element_text(size = 15),
        axis.text.y = element_text(size = 15)) +
  theme(plot.title = element_text(size = 15)) +
  geom_hline(yintercept = hadjpval,  #  horizontal line
                     colour = "#000000",
                     linetype = "dashed") +
  geom_vline(xintercept = -my.lfc,  #  vertical line
                     colour = "#000000",
                     linetype = "dashed") +
  geom_vline(xintercept = my.lfc,  #  vertical line
                     colour = "#000000",
                     linetype = "dashed") +
  ggtitle(paste0(comparison, ", adj.P.Val < ", thresh, ", LFC = ", my.lfc)) +
  geom_text_repel(data = up,
                  aes(x = logFC, y= -log10(P.Value), label = gene_name), 
                  size = 5,
                  color = "maroon", 
                  fontface="italic",
                  max.overlaps = getOption("ggrepel.max.overlaps", default = 30)
                  ) +
  geom_text_repel(data = down,
                  aes(x = logFC, y= -log10(P.Value), label = gene_name), 
                  color = "navyblue", 
                  size = 5,
                  fontface="italic",
                  max.overlaps = getOption("ggrepel.max.overlaps", default = 30)
                  ) +
  ylim(c(0,15))
 p
  
# save
path <- paste0(out, "custom_figures/", comparison, "_FDRq_", 
               format(thresh, nsmall = 2), "_LFC_",
               format(my.lfc, nsmall = 2), "_volcano.pdf")
pdf(path, height = 8, width = 8)
p
dev.off()
```
# '
# '
# '
# Upset Both Sexes
## Low dose
### Plot
```{r}
# set variables
thresh <- 0.1
myLfc <- 0.2
out <- "../results/all_samples_both_sexes_analysis/"
out2 <- "all_samples_both_sexes_analysis"
model <- "DEGs_with_hbb/"

# paths
prefix <- paste0(out, model, "DEG_tables/")
suffix <- "_FDRq_1.00_LFC_0.00.tsv"

# read tables
low.8h <- read.delim(paste0(out, model, 
                            "DEG_tables/L.8h_vs_S.8h_FDRq_1.00_LFC_0.00.tsv"))
low.24h <- read.delim(paste0(out, model, 
                             "DEG_tables/L.24h_vs_S.24h_FDRq_1.00_LFC_0.00.tsv"))
low.7d <- read.delim(paste0(out, model, 
                            "DEG_tables/L.7d_vs_S.7d_FDRq_1.00_LFC_0.00.tsv"))
rownames(low.8h) <- low.8h$gene_name_unique
rownames(low.24h) <- low.24h$gene_name_unique
rownames(low.7d) <- low.7d$gene_name_unique

# filter on pval
low.8h <- low.8h[low.8h$adj.P.Val < thresh,]
low.24h <- low.24h[low.24h$adj.P.Val < thresh,]
low.7d <- low.7d[low.7d$adj.P.Val < thresh,]

# Subset lists
low.8h.up <- subset(low.8h$gene_name_unique, low.8h$logFC > myLfc)
low.8h.down <- subset(low.8h$gene_name_unique, low.8h$logFC < -myLfc)
low.24h.up <- subset(low.24h$gene_name_unique, low.24h$logFC > myLfc)
low.24h.down <- subset(low.24h$gene_name_unique, low.24h$logFC < -myLfc)
low.7d.up <- subset(low.7d$gene_name_unique, low.7d$logFC > myLfc)
low.7d.down <- subset(low.7d$gene_name_unique, low.7d$logFC < -myLfc)
list_input <- list("7 day Low Dose Psilocybin vs Saline Up-regulated" = low.7d.up,
                   "24 hr Low Dose Psilocybin vs Saline Up-regulated" = low.24h.up,
                   "8 hr Low Dose Psilocybin vs Saline Up-regulated" = low.8h.up,
                   "7 day Low Dose Psilocybin vs Saline Down-regulated" = low.7d.down,
                   "24 hr Low Dose Psilocybin vs Saline Down-regulated" = low.24h.down,
                   "8 hr Low Dose Psilocybin vs Saline Down-regulated" = low.8h.down)
data <- fromList(list_input)
  
# store names
names <- colnames(data)
  
# plot
upset_gene <- ComplexUpset::upset(data, 
				  names,
				  set_sizes=(
					upset_set_size()
					+ geom_text(aes(label=..count..), hjust=1.1, stat='count')
					+ expand_limits(y=400)),
				  queries = list(upset_query("8 hr Low Dose Psilocybin vs Saline Up-regulated", fill = "red"),
								 upset_query("24 hr Low Dose Psilocybin vs Saline Up-regulated", fill = "red"),
								 upset_query("7 day Low Dose Psilocybin vs Saline Up-regulated", fill = "red"),
								 upset_query("8 hr Low Dose Psilocybin vs Saline Down-regulated", fill = "blue"),
								 upset_query("24 hr Low Dose Psilocybin vs Saline Down-regulated", fill = "blue"),
								 upset_query("7 day Low Dose Psilocybin vs Saline Down-regulated", fill = "blue")),
				  base_annotations = list('Intersection size' = (
					intersection_size(bar_number_threshold=1, width=0.5)
					+ scale_y_continuous(expand=expansion(mult=c(0, 0.05)),limits = c(0,250)) # space on top
					+ theme(
						  # hide grid lines
						  panel.grid.major=element_blank(),
						  panel.grid.minor=element_blank(),
						  # show axis lines
						  axis.line=element_line(colour='black')))),
				  stripes = upset_stripes(
					geom=geom_segment(size=12),  # make the stripes larger
					colors=c('grey95', 'white')),
				  # to prevent connectors from getting the colorured
				  # use `fill` instead of `color`, together with `shape='circle filled'`
				  matrix = intersection_matrix(
					geom=geom_point(
					  shape='circle filled',
					  size=3,
					  stroke=0.45)),
				  sort_sets=FALSE,
				  sort_intersections='descending'
				)

upset_gene <- upset_gene + ggtitle(paste0("L Comparison, FDRq < ", format(thresh, nsmall = 2),
                                          ", |LFC| > ", format(myLfc, nsmall = 2)))

# save
pdf(paste0(out, model, "/upset/low_dose_comparison_FDRq_", format(thresh, nsmall = 2), 
           "_LFC_", format(myLfc, nsmall = 2), ".pdf"), 
    height = 6, width = 10)
upset_gene
```

### Extract
```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = paste0(out, model, "/upset/low_dose_comparison_results.tsv"),
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

## High dose
### Plot
```{r}
thresh <- 0.1
myLfc <- 0.2
out <- "../results/all_samples_both_sexes_analysis/"
out2 <- "all_samples_both_sexes_analysis"
model <- "DEGs_with_hbb/"

# paths
prefix <- paste0(out, model, "DEG_tables/")
suffix <- "_FDRq_1.00_LFC_0.00.tsv"

# read tables
high.8h <- read.delim(paste0(out, model, 
                            "DEG_tables/H.8h_vs_S.8h_FDRq_1.00_LFC_0.00.tsv"))
high.24h <- read.delim(paste0(out, model, 
                             "DEG_tables/H.24h_vs_S.24h_FDRq_1.00_LFC_0.00.tsv"))
high.7d <- read.delim(paste0(out, model, 
                            "DEG_tables/H.7d_vs_S.7d_FDRq_1.00_LFC_0.00.tsv"))
rownames(high.8h) <- high.8h$gene_name_unique
rownames(high.24h) <- high.24h$gene_name_unique
rownames(high.7d) <- high.7d$gene_name_unique

# filter on pval
high.8h <- high.8h[high.8h$adj.P.Val < thresh,]
high.24h <- high.24h[high.24h$adj.P.Val < thresh,]
high.7d <- high.7d[high.7d$adj.P.Val < thresh,]

# Subset lists
high.8h.up <- subset(high.8h$gene_name_unique, high.8h$logFC > myLfc)
high.8h.down <- subset(high.8h$gene_name_unique, high.8h$logFC < -myLfc)
high.24h.up <- subset(high.24h$gene_name_unique, high.24h$logFC > myLfc)
high.24h.down <- subset(high.24h$gene_name_unique, high.24h$logFC < -myLfc)
high.7d.up <- subset(high.7d$gene_name_unique, high.7d$logFC > myLfc)
high.7d.down <- subset(high.7d$gene_name_unique, high.7d$logFC < -myLfc)
list_input <- list("7 day High Dose Psilocybin vs Saline Up-regulated" = high.7d.up,
                   "24 hr High Dose Psilocybin vs Saline Up-regulated" = high.24h.up,
                   "8 hr High Dose Psilocybin vs Saline Up-regulated" = high.8h.up,
                   "7 day High Dose Psilocybin vs Saline Down-regulated" = high.7d.down,
                   "24 hr High Dose Psilocybin vs Saline Down-regulated" = high.24h.down,
                   "8 hr High Dose Psilocybin vs Saline Down-regulated" = high.8h.down)
data <- fromList(list_input)
  
# store names
names <- colnames(data)
  
# plot
upset_gene <- ComplexUpset::upset(data, 
				  names,
				  set_sizes=(
					upset_set_size()
					+ geom_text(aes(label=..count..), hjust=1.1, stat='count')
					+ expand_limits(y=750)),
				  queries = list(upset_query("8 hr High Dose Psilocybin vs Saline Up-regulated", fill = "red"),
								 upset_query("24 hr High Dose Psilocybin vs Saline Up-regulated", fill = "red"),
								 upset_query("7 day High Dose Psilocybin vs Saline Up-regulated", fill = "red"),
								 upset_query("8 hr High Dose Psilocybin vs Saline Down-regulated", fill = "blue"),
								 upset_query("24 hr High Dose Psilocybin vs Saline Down-regulated", fill = "blue"),
								 upset_query("7 day High Dose Psilocybin vs Saline Down-regulated", fill = "blue")),
				  base_annotations = list('Intersection size' = (
					intersection_size(bar_number_threshold=1, width=0.5)
					+ scale_y_continuous(expand=expansion(mult=c(0, 0.05)),limits = c(0,750)) # space on top
					+ theme(
						  # hide grid lines
						  panel.grid.major=element_blank(),
						  panel.grid.minor=element_blank(),
						  # show axis lines
						  axis.line=element_line(colour='black')))),
				  stripes = upset_stripes(
					geom=geom_segment(size=12),  # make the stripes larger
					colors=c('grey95', 'white')),
				  # to prevent connectors from getting the colorured
				  # use `fill` instead of `color`, together with `shape='circle filled'`
				  matrix = intersection_matrix(
					geom=geom_point(
					  shape='circle filled',
					  size=3,
					  stroke=0.45)),
				  sort_sets=FALSE,
				  sort_intersections='descending'
				)

upset_gene <- upset_gene + ggtitle(paste0("H Comparison, FDRq < ", format(thresh, nsmall = 2),
                                          ", |LFC| > ", format(myLfc, nsmall = 2)))

# save
pdf(paste0(out, model, "/upset/high_dose_comparison_FDRq_", format(thresh, nsmall = 2), 
           "_LFC_", format(myLfc, nsmall = 2), ".pdf"), 
    height = 6, width = 10)
upset_gene
```

### Extract
```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = paste0(out, model, "/upset/high_dose_comparison_results.tsv"),
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

## 8 hours
### Plot
```{r}
# set var
thresh <- 0.1
myLfc <- 0.2
out <- "../results/all_samples_both_sexes_analysis/"
out2 <- "all_samples_both_sexes_analysis"
model <- "DEGs_with_hbb/"

# paths
prefix <- paste0(out, model, "DEG_tables/")
suffix <- "_FDRq_1.00_LFC_0.00.tsv"

# read tables
l.8h <- read.delim(paste0(out, model, "DEG_tables/L.8h_vs_S.8h_FDRq_1.00_LFC_0.00.tsv"))
h.8h <- read.delim(paste0(out, model, "DEG_tables/H.8h_vs_S.8h_FDRq_1.00_LFC_0.00.tsv"))
rownames(l.8h) <- l.8h$gene_name_unique
rownames(h.8h) <- h.8h$gene_name_unique

# filter on pval
l.8h <- l.8h[l.8h$adj.P.Val < thresh,]
h.8h <- h.8h[h.8h$adj.P.Val < thresh,]

# Subset lists
l.8h.up <- subset(l.8h$gene_name_unique, l.8h$logFC > myLfc)
l.8h.down <- subset(l.8h$gene_name_unique, l.8h$logFC < -myLfc)
h.8h.up <- subset(h.8h$gene_name_unique, h.8h$logFC > myLfc)
h.8h.down <- subset(h.8h$gene_name_unique, h.8h$logFC < -myLfc)
list_input <- list("8 hr High Psilocybin vs Saline Up-regulated" = h.8h.up,
                   "8 hr Low Psilocybin vs Saline Up-regulated" = l.8h.up,
                   "8 hr High Psilocybin vs Saline Down-regulated" = h.8h.down,
                   "8 hr Low Psilocybin vs Saline Down-regulated" = l.8h.down)
data <- fromList(list_input)
  
# store names
names <- colnames(data)
  
# plot
upset_gene <- ComplexUpset::upset(data, 
				  names,
				  set_sizes=(
					upset_set_size()
					+ geom_text(aes(label=..count..), hjust=1.1, stat='count')
					+ expand_limits(y=50)),
				  queries = list(
				    upset_query("8 hr High Psilocybin vs Saline Up-regulated", fill = "red"),
				    upset_query("8 hr Low Psilocybin vs Saline Up-regulated", fill = "red"),
				    upset_query("8 hr High Psilocybin vs Saline Down-regulated", fill = "blue"),
				    upset_query("8 hr Low Psilocybin vs Saline Down-regulated", fill = "blue")),
				  base_annotations = list('Intersection size' = (
					intersection_size(bar_number_threshold=1, width=0.5)
					+ scale_y_continuous(expand=expansion(mult=c(0, 0.05)),limits = c(0,50)) # space on top
					+ theme(
						  # hide grid lines
						  panel.grid.major=element_blank(),
						  panel.grid.minor=element_blank(),
						  # show axis lines
						  axis.line=element_line(colour='black')))),
				  stripes = upset_stripes(
					geom=geom_segment(size=12),  # make the stripes larger
					colors=c('grey95', 'white')),
				  # to prevent connectors from getting the colorured
				  # use `fill` instead of `color`, together with `shape='circle filled'`
				  matrix = intersection_matrix(
					geom=geom_point(
					  shape='circle filled',
					  size=3,
					  stroke=0.45)),
				  sort_sets=FALSE,
				  sort_intersections='descending'
				)

upset_gene <- upset_gene + ggtitle(paste0("8 Hours, FDRq < ", format(thresh, nsmall = 2),
                                          ", |LFC| > ", format(myLfc, nsmall = 2)))

# save
pdf(paste0(out, model, "/upset/8_hours_comparison_FDRq_", format(thresh, nsmall = 2), 
           "_LFC_", format(myLfc, nsmall = 2), ".pdf"), 
    height = 6, width = 15)
upset_gene
```

### Extract
```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = paste0(out, model, "/upset/8_hours_comparison_results.tsv"),
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

## 24 hours
### Plot
```{r}
# set var
thresh <- 0.1
myLfc <- 0.2
out <- "../results/all_samples_both_sexes_analysis/"
out2 <- "all_samples_both_sexes_analysis"
model <- "DEGs_with_hbb/"

# paths
prefix <- paste0(out, model, "DEG_tables/")
suffix <- "_FDRq_1.00.tsv"

# read tables
l.24h <- read.delim(paste0(out, model, "DEG_tables/L.24h_vs_S.24h_FDRq_1.00_LFC_0.00.tsv"))
h.24h <- read.delim(paste0(out, model, "DEG_tables/H.24h_vs_S.24h_FDRq_1.00_LFC_0.00.tsv"))
rownames(l.24h) <- l.24h$gene_name_unique
rownames(h.24h) <- h.24h$gene_name_unique

# filter on pval
l.24h <- l.24h[l.24h$adj.P.Val < thresh,]
h.24h <- h.24h[h.24h$adj.P.Val < thresh,]

# Subset lists
l.24h.up <- subset(l.24h$gene_name_unique, l.24h$logFC > myLfc)
l.24h.down <- subset(l.24h$gene_name_unique, l.24h$logFC < -myLfc)
h.24h.up <- subset(h.24h$gene_name_unique, h.24h$logFC > myLfc)
h.24h.down <- subset(h.24h$gene_name_unique, h.24h$logFC < -myLfc)
list_input <- list("24 hr High Psilocybin vs Saline Up-regulated" = h.24h.up,
                   "24 hr Low Psilocybin vs Saline Up-regulated" = l.24h.up,
                   "24 hr High Psilocybin vs Saline Down-regulated" = h.24h.down,
                   "24 hr Low Psilocybin vs Saline Down-regulated" = l.24h.down)
data <- fromList(list_input)
  
# store names
names <- colnames(data)
  
# plot
upset_gene <- ComplexUpset::upset(data, 
				  names,
				  set_sizes=(
					upset_set_size()
					+ geom_text(aes(label=..count..), hjust=1.1, stat='count')
					+ expand_limits(y=750)),
				  queries = list(
				    upset_query("24 hr High Psilocybin vs Saline Up-regulated", fill = "red"),
				    upset_query("24 hr Low Psilocybin vs Saline Up-regulated", fill = "red"),
						upset_query("24 hr High Psilocybin vs Saline Down-regulated", fill = "blue"),
				    upset_query("24 hr Low Psilocybin vs Saline Down-regulated", fill = "blue")),
				  base_annotations = list('Intersection size' = (
					intersection_size(bar_number_threshold=1, width=0.5)
					+ scale_y_continuous(expand=expansion(mult=c(0, 0.05)),limits = c(0,750)) # space on top
					+ theme(
						  # hide grid lines
						  panel.grid.major=element_blank(),
						  panel.grid.minor=element_blank(),
						  # show axis lines
						  axis.line=element_line(colour='black')))),
				  stripes = upset_stripes(
					geom=geom_segment(size=12),  # make the stripes larger
					colors=c('grey95', 'white')),
				  # to prevent connectors from getting the colorured
				  # use `fill` instead of `color`, together with `shape='circle filled'`
				  matrix = intersection_matrix(
					geom=geom_point(
					  shape='circle filled',
					  size=3,
					  stroke=0.45)),
				  sort_sets=FALSE,
				  sort_intersections='descending'
				)

upset_gene <- upset_gene + ggtitle(paste0("24 Hours, FDRq < ", format(thresh, nsmall = 2),
                                          ", |LFC| > ", format(myLfc, nsmall = 2)))

# save
pdf(paste0(out, model, "/upset/24_hours_comparison_FDRq_", format(thresh, nsmall = 2), 
           "_LFC_", format(myLfc, nsmall = 2), ".pdf"), 
    height = 6, width = 16)
upset_gene
```

### Extract
```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = paste0(out, model, "/upset/24_hours_comparison_results.tsv"),
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

## 7 days
### Plot
```{r}
# set var
thresh <- 0.1
myLfc <- 0.2
out <- "../results/all_samples_both_sexes_analysis/"
out2 <- "all_samples_both_sexes_analysis"
model <- "DEGs_with_hbb/"

# paths
prefix <- paste0(out, model, "DEG_tables/")
suffix <- "_FDRq_1.00.tsv"

# read tables
l.7d <- read.delim(paste0(out, model, "DEG_tables/L.7d_vs_S.7d_FDRq_1.00_LFC_0.00.tsv"))
h.7d <- read.delim(paste0(out, model, "DEG_tables/H.7d_vs_S.7d_FDRq_1.00_LFC_0.00.tsv"))
rownames(l.7d) <- l.7d$gene_name_unique
rownames(h.7d) <- h.7d$gene_name_unique

# filter on pval
l.7d <- l.7d[l.7d$adj.P.Val < thresh,]
h.7d <- h.7d[h.7d$adj.P.Val < thresh,]

# Subset lists
l.7d.up <- subset(l.7d$gene_name_unique, l.7d$logFC > myLfc)
l.7d.down <- subset(l.7d$gene_name_unique, l.7d$logFC < -myLfc)
h.7d.up <- subset(h.7d$gene_name_unique, h.7d$logFC > myLfc)
h.7d.down <- subset(h.7d$gene_name_unique, h.7d$logFC < -myLfc)
list_input <- list("7 days High Psilocybin vs Saline Up-regulated" = h.7d.up,
                   "7 days Low Psilocybin vs Saline Up-regulated" = l.7d.up,
                   "7 days High Psilocybin vs Saline Down-regulated" = h.7d.down,
                   "7 days Low Psilocybin vs Saline Down-regulated" = l.7d.down)
data <- fromList(list_input)
  
# store names
names <- colnames(data)
  
# plot
upset_gene <- ComplexUpset::upset(data, 
				  names,
				  set_sizes=(
					upset_set_size()
					+ geom_text(aes(label=..count..), hjust=1.1, stat='count')
					+ expand_limits(y=200)),
				  queries = list(
				    upset_query("7 days High Psilocybin vs Saline Up-regulated", fill = "red"),
				    upset_query("7 days Low Psilocybin vs Saline Up-regulated", fill = "red"),
						upset_query("7 days High Psilocybin vs Saline Down-regulated", fill = "blue"),
				    upset_query("7 days Low Psilocybin vs Saline Down-regulated", fill = "blue")),
				  base_annotations = list('Intersection size' = (
					intersection_size(bar_number_threshold=1, width=0.5)
					+ scale_y_continuous(expand=expansion(mult=c(0, 0.05)),limits = c(0,200)) # space on top
					+ theme(
						  # hide grid lines
						  panel.grid.major=element_blank(),
						  panel.grid.minor=element_blank(),
						  # show axis lines
						  axis.line=element_line(colour='black')))),
				  stripes = upset_stripes(
					geom=geom_segment(size=12),  # make the stripes larger
					colors=c('grey95', 'white')),
				  # to prevent connectors from getting the colorured
				  # use `fill` instead of `color`, together with `shape='circle filled'`
				  matrix = intersection_matrix(
					geom=geom_point(
					  shape='circle filled',
					  size=3,
					  stroke=0.45)),
				  sort_sets=FALSE,
				  sort_intersections='descending'
				)

upset_gene <- upset_gene + ggtitle(paste0("7 Days, FDRq < ", format(thresh, nsmall = 2),
                                          ", |LFC| > ", format(myLfc, nsmall = 2)))

# save
pdf(paste0(out, model, "/upset/7_days_comparison_FDRq_", format(thresh, nsmall = 2), 
           "_LFC_", format(myLfc, nsmall = 2), ".pdf"), 
    height = 6, width = 16)
upset_gene
```

### Extract
```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = paste0(out, model, "/upset/7_days_comparison_results.tsv"),
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

# Upset Sex Specific
## Low dose
```{r}
# set variables
out <- "../results/all_samples_sex_specific_combat_analysis/"
out2 <- "all_samples_sex_specific_combat_analysis"
model <- "DEGs_with_hbb/"

# get file list
files <- list.files(paste0(out, model, "metascape_input"))
files <- files[grepl(pattern = "^L.+", x = files)]

# loop through files and create variable for each gene list
for (i in files) {
  genes <- readLines(paste0(out, model, "metascape_input/", i))
  genes <- genes[genes != ""]
  var_name <- gsub(pattern = "regulated_FDRq_0.10_LFC_0.20.txt",
                   replacement = "",
                   x = i)
  assign(x = var_name, value = genes)
}

# format in a list
list_input <- list(
  "L.7d.F vs S.7d.F down-regulated" = L.7d.F_vs_S.7d.F_down,
  "L.7d.M vs S.7d.M down-regulated" = L.7d.M_vs_S.7d.M_down,
  "L.24h.F vs S.24h.F down-regulated" = L.24h.F_vs_S.24h.F_down,
  "L.24h.M vs S.24h.M down-regulated" = L.24h.M_vs_S.24h.M_down,
  "L.8h.F vs S.8h.F down-regulated" = L.8h.F_vs_S.8h.F_down,
  "L.8h.M vs S.8h.M down-regulated" = L.8h.M_vs_S.8h.M_down,
  "L.7d.F vs S.7d.F up-regulated" = L.7d.F_vs_S.7d.F_up,
  "L.7d.M vs S.7d.M up-regulated" = L.7d.M_vs_S.7d.M_up,
  "L.24h.F vs S.24h.F up-regulated" = L.24h.F_vs_S.24h.F_up,
  "L.24h.M vs S.24h.M up-regulated" = L.24h.M_vs_S.24h.M_up,
  "L.8h.F vs S.8h.F up-regulated" = L.8h.F_vs_S.8h.F_up,
  "L.8h.M vs S.8h.M up-regulated" = L.8h.M_vs_S.8h.M_up
)
data <- UpSetR::fromList(list_input)

# You will specify the male and female groups by their group number
group_mapping <- data.frame(
  Group_Number = 1:12,  # Group numbers (1-8)
  Gene_Set_Name = sort(colnames(data))  # Sorted gene set names
)
group_mapping

# Assign shapes based on group numbers
male_groups <- c("3","4","7","8","11","12")  # Male group numbers
female_groups <- c("1","2","5","6","9","10")  # Female group numbers

# Assign colors and shapes based on specific group numbers
upregulated_groups <- c("2","4","6","8","10","12")
downregulated_groups <- c("1","3","5","7","9","11")

# Create the basic UpSet plot (no dynamic colors or shapes) to extract n_points
dummy_upset <- ComplexUpset::upset(
  data, 
  colnames(data),  # Use column names (gene sets)
  set_sizes = ComplexUpset::upset_set_size(),  # Display set sizes
  sort_sets = FALSE,
  
  # Plot the intersection matrix with points (default shapes)
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(size = 3)  # Simple points without dynamic aesthetics
  )
)

# Ensure n_points matches the actual data
n_points <- nrow(dummy_upset$data)
print(paste("Number of points being plotted:", n_points))  # Should be 300

# Create empty vectors for colors and shapes
color_vector <- rep(NA, n_points)  # To store colors
shape_vector <- rep(NA, n_points)  # To store shapes

# Assign shapes and colors based on group numbers
for (i in 1:n_points) {
  group_value <- as.character(dummy_upset$data$group[i])
  
  # Assign shapes
  if (group_value %in% male_groups) {
    shape_vector[i] <- 22  # Filled square for Male comparisons
  } else if (group_value %in% female_groups) {
    shape_vector[i] <- 21  # Circle for Female comparisons
  }
  
  # Assign colors based on group numbers, only if the point is involved in an intersection
  if (dummy_upset$data$value[i]) {  # If the point is part of an intersection
    if (group_value %in% upregulated_groups) {
      color_vector[i] <- "red"  # Red for Up-regulated
    } else if (group_value %in% downregulated_groups) {
      color_vector[i] <- "blue"  # Blue for Down-regulated
    }
  } else {
    color_vector[i] <- NA  # No fill for points not involved in intersections
  }
}

# Ensure both vectors are of correct length
if (length(shape_vector) != n_points || length(color_vector) != n_points) {
  stop("Error: shape_vector or color_vector length does not match the number of plotted points.")
}

# Print the color_vector to ensure correctness before plotting
print(unique(color_vector))  # Ensure this contains only valid colors or NA

# Reintroduce dynamic fill color for filled shapes
upset_gene <- ComplexUpset::upset(
  data, 
  colnames(data),
  set_sizes = (
    ComplexUpset::upset_set_size() +
    geom_text(aes(label = ..count..), hjust = 1.1, stat = 'count') +
    expand_limits(y = 1000)
  ),
  
  # Prevent automatic sorting of sets
  sort_sets = FALSE,
  
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(
      ggplot2::aes(shape = factor(shape_vector), fill = color_vector),  # Fill shapes with dynamic colors
      size = 3,
      stroke = 0.45  # Keeps the outline black
    )
  ) + ggplot2::scale_shape_manual(
    name = "Sex",
    labels = c("Male", "Female"),  # Legend labels
    values = c(21, 22) # Circle=21, Filled square=22
    ) + ggplot2::scale_fill_identity(na.translate = FALSE),  # Use the exact colors in color_vector and remove NA fill
  
    base_annotations = list(
    'Intersection size' = (
      intersection_size(bar_number_threshold = 1, width = 0.5) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, 800)) +
      theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = 'black')
      )
    )
  ),
)

# Add title
upset_gene <- upset_gene + 
  ggtitle("Low Dose Comparison, FDRq < 0.10, |LFC| > 0.20")

# Remove legend
upset_gene <- upset_gene + theme(legend.position = "None")

# Save temp without legend
pdf(file = paste0(out, model, "/upset/low_dose_comparison_FDRq_0.10_LFC_0.20.pdf"), 
    height = 7, 
    width = 12)
upset_gene
```


```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = paste0(out, model, "/upset/low_dose_comparison_results.tsv"),
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

## High dose
```{r}
# set variables
out <- "../results/all_samples_sex_specific_combat_analysis/"
out2 <- "all_samples_sex_specific_combat_analysis"
model <- "DEGs_with_hbb/"

# get file list
files <- list.files(paste0(out, model, "metascape_input"))
files <- files[grepl(pattern = "^H.+", x = files)]

# loop through files and create variable for each gene list
for (i in files) {
  genes <- readLines(paste0(out, model, "metascape_input/", i))
  genes <- genes[genes != ""]
  var_name <- gsub(pattern = "regulated_FDRq_0.10_LFC_0.20.txt",
                   replacement = "",
                   x = i)
  assign(x = var_name, value = genes)
}

# format in a list
list_input <- list(
  "H.7d.F vs S.7d.F down-regulated" = H.7d.F_vs_S.7d.F_down,
  "H.7d.M vs S.7d.M down-regulated" = H.7d.M_vs_S.7d.M_down,
  "H.24h.F vs S.24h.F down-regulated" = H.24h.F_vs_S.24h.F_down,
  "H.24h.M vs S.24h.M down-regulated" = H.24h.M_vs_S.24h.M_down,
  "H.8h.F vs S.8h.F down-regulated" = H.8h.F_vs_S.8h.F_down,
  "H.8h.M vs S.8h.M down-regulated" = H.8h.M_vs_S.8h.M_down,
  "H.7d.F vs S.7d.F up-regulated" = H.7d.F_vs_S.7d.F_up,
  "H.7d.M vs S.7d.M up-regulated" = H.7d.M_vs_S.7d.M_up,
  "H.24h.F vs S.24h.F up-regulated" = H.24h.F_vs_S.24h.F_up,
  "H.24h.M vs S.24h.M up-regulated" = H.24h.M_vs_S.24h.M_up,
  "H.8h.F vs S.8h.F up-regulated" = H.8h.F_vs_S.8h.F_up,
  "H.8h.M vs S.8h.M up-regulated" = H.8h.M_vs_S.8h.M_up
)
data <- UpSetR::fromList(list_input)

# You will specify the male and female groups by their group number
group_mapping <- data.frame(
  Group_Number = 1:12,  # Group numbers (1-8)
  Gene_Set_Name = sort(colnames(data))  # Sorted gene set names
)
group_mapping

# Assign shapes based on group numbers
male_groups <- c("3","4","7","8","11","12")  # Male group numbers
female_groups <- c("1","2","5","6","9","10")  # Female group numbers

# Assign colors and shapes based on specific group numbers
upregulated_groups <- c("2","4","6","8","10","12")
downregulated_groups <- c("1","3","5","7","9","11")

# Create the basic UpSet plot (no dynamic colors or shapes) to extract n_points
dummy_upset <- ComplexUpset::upset(
  data, 
  colnames(data),  # Use column names (gene sets)
  set_sizes = ComplexUpset::upset_set_size(),  # Display set sizes
  sort_sets = FALSE,
  
  # Plot the intersection matrix with points (default shapes)
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(size = 3)  # Simple points without dynamic aesthetics
  )
)

# Ensure n_points matches the actual data
n_points <- nrow(dummy_upset$data)
print(paste("Number of points being plotted:", n_points))  # Should be 300

# Create empty vectors for colors and shapes
color_vector <- rep(NA, n_points)  # To store colors
shape_vector <- rep(NA, n_points)  # To store shapes

# Assign shapes and colors based on group numbers
for (i in 1:n_points) {
  group_value <- as.character(dummy_upset$data$group[i])
  
  # Assign shapes
  if (group_value %in% male_groups) {
    shape_vector[i] <- 22  # Filled square for Male comparisons
  } else if (group_value %in% female_groups) {
    shape_vector[i] <- 21  # Circle for Female comparisons
  }
  
  # Assign colors based on group numbers, only if the point is involved in an intersection
  if (dummy_upset$data$value[i]) {  # If the point is part of an intersection
    if (group_value %in% upregulated_groups) {
      color_vector[i] <- "red"  # Red for Up-regulated
    } else if (group_value %in% downregulated_groups) {
      color_vector[i] <- "blue"  # Blue for Down-regulated
    }
  } else {
    color_vector[i] <- NA  # No fill for points not involved in intersections
  }
}

# Ensure both vectors are of correct length
if (length(shape_vector) != n_points || length(color_vector) != n_points) {
  stop("Error: shape_vector or color_vector length does not match the number of plotted points.")
}

# Print the color_vector to ensure correctness before plotting
print(unique(color_vector))  # Ensure this contains only valid colors or NA

# Reintroduce dynamic fill color for filled shapes
upset_gene <- ComplexUpset::upset(
  data, 
  colnames(data),
  set_sizes = (
    ComplexUpset::upset_set_size() +
    geom_text(aes(label = ..count..), hjust = 1.1, stat = 'count') +
    expand_limits(y = 600)
  ),
  
  # Prevent automatic sorting of sets
  sort_sets = FALSE,
  
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(
      ggplot2::aes(shape = factor(shape_vector), fill = color_vector),  # Fill shapes with dynamic colors
      size = 3,
      stroke = 0.45  # Keeps the outline black
    )
  ) + ggplot2::scale_shape_manual(
    name = "Sex",
    labels = c("Male", "Female"),  # Legend labels
    values = c(21, 22) # Circle=21, Filled square=22
    ) + ggplot2::scale_fill_identity(na.translate = FALSE),  # Use the exact colors in color_vector and remove NA fill
  
    base_annotations = list(
    'Intersection size' = (
      intersection_size(bar_number_threshold = 1, width = 0.5) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, 400)) +
      theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = 'black')
      )
    )
  ),
)

# Add title
upset_gene <- upset_gene + 
  ggtitle("High Dose Comparison, FDRq < 0.10, |LFC| > 0.20")

# Remove legend
upset_gene <- upset_gene + theme(legend.position = "None")

# Save temp without legend
pdf(file = paste0(out, model, "/upset/high_dose_comparison_FDRq_0.10_LFC_0.20.pdf"), 
    height = 7, 
    width = 14)
upset_gene
dev.off()
```


```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = paste0(out, model, "/upset/high_dose_comparison_results.tsv"),
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

## 8 hours
```{r}
# set variables
out <- "../results/all_samples_sex_specific_combat_analysis/"
out2 <- "all_samples_sex_specific_combat_analysis"
model <- "DEGs_with_hbb/"

# get file list
files <- list.files(paste0(out, model, "metascape_input"))
files <- files[grepl(pattern = "^[LH]\\.8h.+", x = files)]

# loop through files and create variable for each gene list
for (i in files) {
  genes <- readLines(paste0(out, model, "metascape_input/", i))
  genes <- genes[genes != ""]
  var_name <- gsub(pattern = "regulated_FDRq_0.10_LFC_0.20.txt",
                   replacement = "",
                   x = i)
  assign(x = var_name, value = genes)
}

# format in a list
list_input <- list(
  "H.8h.F vs S.8h.F down-regulated" = H.8h.F_vs_S.8h.F_down,
  "H.8h.M vs S.8h.M down-regulated" = H.8h.M_vs_S.8h.M_down,
  "L.8h.F vs S.8h.F down-regulated" = L.8h.F_vs_S.8h.F_down,
  "L.8h.M vs S.8h.M down-regulated" = L.8h.M_vs_S.8h.M_down,
  "H.8h.F vs S.8h.F up-regulated" = H.8h.F_vs_S.8h.F_up,
  "H.8h.M vs S.8h.M up-regulated" = H.8h.M_vs_S.8h.M_up,
  "L.8h.F vs S.8h.F up-regulated" = L.8h.F_vs_S.8h.F_up,
  "L.8h.M vs S.8h.M up-regulated" = L.8h.M_vs_S.8h.M_up
)
data <- UpSetR::fromList(list_input)

# You will specify the male and female groups by their group number
group_mapping <- data.frame(
  Group_Number = 1:8,  # Group numbers (1-8)
  Gene_Set_Name = sort(colnames(data))  # Sorted gene set names
)
group_mapping

# Assign shapes based on group numbers
male_groups <- c("3","4","7","8")  # Male group numbers
female_groups <- c("1","2","5","6")  # Female group numbers

# Assign colors and shapes based on specific group numbers
upregulated_groups <- c("2","4","6","8")
downregulated_groups <- c("1","3","5","7")

# Create the basic UpSet plot (no dynamic colors or shapes) to extract n_points
dummy_upset <- ComplexUpset::upset(
  data, 
  colnames(data),  # Use column names (gene sets)
  set_sizes = ComplexUpset::upset_set_size(),  # Display set sizes
  sort_sets = FALSE,
  
  # Plot the intersection matrix with points (default shapes)
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(size = 3)  # Simple points without dynamic aesthetics
  )
)

# Ensure n_points matches the actual data
n_points <- nrow(dummy_upset$data)
print(paste("Number of points being plotted:", n_points))  # Should be 300

# Create empty vectors for colors and shapes
color_vector <- rep(NA, n_points)  # To store colors
shape_vector <- rep(NA, n_points)  # To store shapes

# Assign shapes and colors based on group numbers
for (i in 1:n_points) {
  group_value <- as.character(dummy_upset$data$group[i])
  
  # Assign shapes
  if (group_value %in% male_groups) {
    shape_vector[i] <- 22  # Filled square for Male comparisons
  } else if (group_value %in% female_groups) {
    shape_vector[i] <- 21  # Circle for Female comparisons
  }
  
  # Assign colors based on group numbers, only if the point is involved in an intersection
  if (dummy_upset$data$value[i]) {  # If the point is part of an intersection
    if (group_value %in% upregulated_groups) {
      color_vector[i] <- "red"  # Red for Up-regulated
    } else if (group_value %in% downregulated_groups) {
      color_vector[i] <- "blue"  # Blue for Down-regulated
    }
  } else {
    color_vector[i] <- NA  # No fill for points not involved in intersections
  }
}

# Ensure both vectors are of correct length
if (length(shape_vector) != n_points || length(color_vector) != n_points) {
  stop("Error: shape_vector or color_vector length does not match the number of plotted points.")
}

# Print the color_vector to ensure correctness before plotting
print(unique(color_vector))  # Ensure this contains only valid colors or NA

# Reintroduce dynamic fill color for filled shapes
upset_gene <- ComplexUpset::upset(
  data, 
  colnames(data),
  set_sizes = (
    ComplexUpset::upset_set_size() +
    geom_text(aes(label = ..count..), hjust = 1.1, stat = 'count') +
    expand_limits(y = 800)
  ),
  
  # Prevent automatic sorting of sets
  sort_sets = FALSE,
  
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(
      ggplot2::aes(shape = factor(shape_vector), fill = color_vector),  # Fill shapes with dynamic colors
      size = 3,
      stroke = 0.45  # Keeps the outline black
    )
  ) + ggplot2::scale_shape_manual(
    name = "Sex",
    labels = c("Male", "Female"),  # Legend labels
    values = c(21, 22) # Circle=21, Filled square=22
    ) + ggplot2::scale_fill_identity(na.translate = FALSE),  # Use the exact colors in color_vector and remove NA fill
  
    base_annotations = list(
    'Intersection size' = (
      intersection_size(bar_number_threshold = 1, width = 0.5) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, 600)) +
      theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = 'black')
      )
    )
  ),
)

# Add title
upset_gene <- upset_gene + 
  ggtitle("8 Hours Comparison, FDRq < 0.10, |LFC| > 0.20")

# Remove legend
upset_gene <- upset_gene + theme(legend.position = "None")

# Save temp without legend
pdf(file = paste0(out, model, "/upset/8_hours_comparison_FDRq_0.10_LFC_0.20.pdf"), 
    height = 7, 
    width = 10)
upset_gene
```


```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = paste0(out, model, "/upset/8_hours_comparison_results.tsv"),
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

## 24 hours
```{r}
# set variables
out <- "../results/all_samples_sex_specific_combat_analysis/"
out2 <- "all_samples_sex_specific_combat_analysis"
model <- "DEGs_with_hbb/"

# get file list
files <- list.files(paste0(out, model, "metascape_input"))
files <- files[grepl(pattern = "^[LH]\\.24h.+", x = files)]

# loop through files and create variable for each gene list
for (i in files) {
  genes <- readLines(paste0(out, model, "metascape_input/", i))
  genes <- genes[genes != ""]
  var_name <- gsub(pattern = "regulated_FDRq_0.10_LFC_0.20.txt",
                   replacement = "",
                   x = i)
  assign(x = var_name, value = genes)
}

# format in a list
list_input <- list(
  "H.24h.F vs S.24h.F down-regulated" = H.24h.F_vs_S.24h.F_down,
  "H.24h.M vs S.24h.M down-regulated" = H.24h.M_vs_S.24h.M_down,
  "L.24h.F vs S.24h.F down-regulated" = L.24h.F_vs_S.24h.F_down,
  "L.24h.M vs S.24h.M down-regulated" = L.24h.M_vs_S.24h.M_down,
  "H.24h.F vs S.24h.F up-regulated" = H.24h.F_vs_S.24h.F_up,
  "H.24h.M vs S.24h.M up-regulated" = H.24h.M_vs_S.24h.M_up,
  "L.24h.F vs S.24h.F up-regulated" = L.24h.F_vs_S.24h.F_up,
  "L.24h.M vs S.24h.M up-regulated" = L.24h.M_vs_S.24h.M_up
)
data <- UpSetR::fromList(list_input)

# You will specify the male and female groups by their group number
group_mapping <- data.frame(
  Group_Number = 1:8,  # Group numbers (1-8)
  Gene_Set_Name = sort(colnames(data))  # Sorted gene set names
)
group_mapping

# Assign shapes based on group numbers
male_groups <- c("3","4","7","8")  # Male group numbers
female_groups <- c("1","2","5","6")  # Female group numbers

# Assign colors and shapes based on specific group numbers
upregulated_groups <- c("2","4","6","8")
downregulated_groups <- c("1","3","5","7")

# Create the basic UpSet plot (no dynamic colors or shapes) to extract n_points
dummy_upset <- ComplexUpset::upset(
  data, 
  colnames(data),  # Use column names (gene sets)
  set_sizes = ComplexUpset::upset_set_size(),  # Display set sizes
  sort_sets = FALSE,
  
  # Plot the intersection matrix with points (default shapes)
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(size = 3)  # Simple points without dynamic aesthetics
  )
)

# Ensure n_points matches the actual data
n_points <- nrow(dummy_upset$data)
print(paste("Number of points being plotted:", n_points))  # Should be 300

# Create empty vectors for colors and shapes
color_vector <- rep(NA, n_points)  # To store colors
shape_vector <- rep(NA, n_points)  # To store shapes

# Assign shapes and colors based on group numbers
for (i in 1:n_points) {
  group_value <- as.character(dummy_upset$data$group[i])
  
  # Assign shapes
  if (group_value %in% male_groups) {
    shape_vector[i] <- 22  # Filled square for Male comparisons
  } else if (group_value %in% female_groups) {
    shape_vector[i] <- 21  # Circle for Female comparisons
  }
  
  # Assign colors based on group numbers, only if the point is involved in an intersection
  if (dummy_upset$data$value[i]) {  # If the point is part of an intersection
    if (group_value %in% upregulated_groups) {
      color_vector[i] <- "red"  # Red for Up-regulated
    } else if (group_value %in% downregulated_groups) {
      color_vector[i] <- "blue"  # Blue for Down-regulated
    }
  } else {
    color_vector[i] <- NA  # No fill for points not involved in intersections
  }
}

# Ensure both vectors are of correct length
if (length(shape_vector) != n_points || length(color_vector) != n_points) {
  stop("Error: shape_vector or color_vector length does not match the number of plotted points.")
}

# Print the color_vector to ensure correctness before plotting
print(unique(color_vector))  # Ensure this contains only valid colors or NA

# Reintroduce dynamic fill color for filled shapes
upset_gene <- ComplexUpset::upset(
  data, 
  colnames(data),
  set_sizes = (
    ComplexUpset::upset_set_size() +
    geom_text(aes(label = ..count..), hjust = 1.1, stat = 'count') +
    expand_limits(y = 1000)
  ),
  
  # Prevent automatic sorting of sets
  sort_sets = FALSE,
  
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(
      ggplot2::aes(shape = factor(shape_vector), fill = color_vector),  # Fill shapes with dynamic colors
      size = 3,
      stroke = 0.45  # Keeps the outline black
    )
  ) + ggplot2::scale_shape_manual(
    name = "Sex",
    labels = c("Male", "Female"),  # Legend labels
    values = c(21, 22) # Circle=21, Filled square=22
    ) + ggplot2::scale_fill_identity(na.translate = FALSE),  # Use the exact colors in color_vector and remove NA fill
  
    base_annotations = list(
    'Intersection size' = (
      intersection_size(bar_number_threshold = 1, width = 0.5) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, 800)) +
      theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = 'black')
      )
    )
  ),
)

# Add title
upset_gene <- upset_gene + 
  ggtitle("24 Hours Comparison, FDRq < 0.10, |LFC| > 0.20")

# Remove legend
upset_gene <- upset_gene + theme(legend.position = "None")

# Save temp without legend
pdf(file = paste0(out, model, "/upset/24_hours_comparison_FDRq_0.10_LFC_0.20.pdf"), 
    height = 7, 
    width = 11)
upset_gene
```


```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = paste0(out, model, "/upset/24_hours_comparison_results.tsv"),
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

## 7 days
```{r}
# set variables
out <- "../results/all_samples_sex_specific_combat_analysis/"
out2 <- "all_samples_sex_specific_combat_analysis"
model <- "DEGs_with_hbb/"

# get file list
files <- list.files(paste0(out, model, "metascape_input"))
files <- files[grepl(pattern = "^[LH]\\.7d.+", x = files)]

# loop through files and create variable for each gene list
for (i in files) {
  genes <- readLines(paste0(out, model, "metascape_input/", i))
  genes <- genes[genes != ""]
  var_name <- gsub(pattern = "regulated_FDRq_0.10_LFC_0.20.txt",
                   replacement = "",
                   x = i)
  assign(x = var_name, value = genes)
}

# format in a list
list_input <- list(
  "H.7d.F vs S.7d.F down-regulated" = H.7d.F_vs_S.7d.F_down,
  "H.7d.M vs S.7d.M down-regulated" = H.7d.M_vs_S.7d.M_down,
  "L.7d.F vs S.7d.F down-regulated" = L.7d.F_vs_S.7d.F_down,
  "L.7d.M vs S.7d.M down-regulated" = L.7d.M_vs_S.7d.M_down,
  "H.7d.F vs S.7d.F up-regulated" = H.7d.F_vs_S.7d.F_up,
  "H.7d.M vs S.7d.M up-regulated" = H.7d.M_vs_S.7d.M_up,
  "L.7d.F vs S.7d.F up-regulated" = L.7d.F_vs_S.7d.F_up,
  "L.7d.M vs S.7d.M up-regulated" = L.7d.M_vs_S.7d.M_up
)
data <- UpSetR::fromList(list_input)

# You will specify the male and female groups by their group number
group_mapping <- data.frame(
  Group_Number = 1:8,  # Group numbers (1-8)
  Gene_Set_Name = sort(colnames(data))  # Sorted gene set names
)
group_mapping

# Assign shapes based on group numbers
male_groups <- c("3","4","7","8")  # Male group numbers
female_groups <- c("1","2","5","6")  # Female group numbers

# Assign colors and shapes based on specific group numbers
upregulated_groups <- c("2","4","6","8")
downregulated_groups <- c("1","3","5","7")

# Create the basic UpSet plot (no dynamic colors or shapes) to extract n_points
dummy_upset <- ComplexUpset::upset(
  data, 
  colnames(data),  # Use column names (gene sets)
  set_sizes = ComplexUpset::upset_set_size(),  # Display set sizes
  sort_sets = FALSE,
  
  # Plot the intersection matrix with points (default shapes)
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(size = 3)  # Simple points without dynamic aesthetics
  )
)

# Ensure n_points matches the actual data
n_points <- nrow(dummy_upset$data)
print(paste("Number of points being plotted:", n_points))  # Should be 300

# Create empty vectors for colors and shapes
color_vector <- rep(NA, n_points)  # To store colors
shape_vector <- rep(NA, n_points)  # To store shapes

# Assign shapes and colors based on group numbers
for (i in 1:n_points) {
  group_value <- as.character(dummy_upset$data$group[i])
  
  # Assign shapes
  if (group_value %in% male_groups) {
    shape_vector[i] <- 22  # Filled square for Male comparisons
  } else if (group_value %in% female_groups) {
    shape_vector[i] <- 21  # Circle for Female comparisons
  }
  
  # Assign colors based on group numbers, only if the point is involved in an intersection
  if (dummy_upset$data$value[i]) {  # If the point is part of an intersection
    if (group_value %in% upregulated_groups) {
      color_vector[i] <- "red"  # Red for Up-regulated
    } else if (group_value %in% downregulated_groups) {
      color_vector[i] <- "blue"  # Blue for Down-regulated
    }
  } else {
    color_vector[i] <- NA  # No fill for points not involved in intersections
  }
}

# Ensure both vectors are of correct length
if (length(shape_vector) != n_points || length(color_vector) != n_points) {
  stop("Error: shape_vector or color_vector length does not match the number of plotted points.")
}

# Print the color_vector to ensure correctness before plotting
print(unique(color_vector))  # Ensure this contains only valid colors or NA

# Reintroduce dynamic fill color for filled shapes
upset_gene <- ComplexUpset::upset(
  data, 
  colnames(data),
  set_sizes = (
    ComplexUpset::upset_set_size() +
    geom_text(aes(label = ..count..), hjust = 1.1, stat = 'count') +
    expand_limits(y = 100)
  ),
  
  # Prevent automatic sorting of sets
  sort_sets = FALSE,
  
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(
      ggplot2::aes(shape = factor(shape_vector), fill = color_vector),  # Fill shapes with dynamic colors
      size = 3,
      stroke = 0.45  # Keeps the outline black
    )
  ) + ggplot2::scale_shape_manual(
    name = "Sex",
    labels = c("Male", "Female"),  # Legend labels
    values = c(21, 22) # Circle=21, Filled square=22
    ) + ggplot2::scale_fill_identity(na.translate = FALSE),  # Use the exact colors in color_vector and remove NA fill
  
    base_annotations = list(
    'Intersection size' = (
      intersection_size(bar_number_threshold = 1, width = 0.5) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, 100)) +
      theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = 'black')
      )
    )
  ),
)

# Add title
upset_gene <- upset_gene + 
  ggtitle("7 Days Comparison, FDRq < 0.10, |LFC| > 0.20")

# Remove legend
upset_gene <- upset_gene + theme(legend.position = "None")

# Save temp without legend
pdf(file = paste0(out, model, "/upset/7_days_comparison_FDRq_0.10_LFC_0.20.pdf"), 
    height = 6, 
    width = 8)
upset_gene
```

```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = paste0(out, model, "/upset/7_days_comparison_results.tsv"),
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

# Upset: sex specific, with/without ComBat
## L.8h
```{r}
# set variables
suffix <- "_FDRq_0.10_LFC_0.20.txt"
combat_prefix <- paste0("../results/all_samples_sex_specific_combat_analysis/",
                        "DEGs_with_hbb/metascape_input/")
nc_prefix <- paste0("../results/all_samples_sex_specific_analysis/",
                        "DEGs_with_hbb/metascape_input/")

# get file list
combat_L.8h.F_vs_S.8h.F_down <- 
  readLines(paste0(combat_prefix, "L.8h.F_vs_S.8h.F_downregulated", suffix))
combat_L.8h.M_vs_S.8h.M_down <- 
  readLines(paste0(combat_prefix, "L.8h.M_vs_S.8h.M_downregulated", suffix))
nc_L.8h.F_vs_S.8h.F_down <- 
  readLines(paste0(nc_prefix, "L.8h.F_vs_S.8h.F_downregulated", suffix))
nc_L.8h.M_vs_S.8h.M_down <- 
  readLines(paste0(nc_prefix, "L.8h.M_vs_S.8h.M_downregulated", suffix))
combat_L.8h.F_vs_S.8h.F_up <- 
  readLines(paste0(combat_prefix, "L.8h.F_vs_S.8h.F_upregulated", suffix))
combat_L.8h.M_vs_S.8h.M_up <- 
  readLines(paste0(combat_prefix, "L.8h.M_vs_S.8h.M_upregulated", suffix))
nc_L.8h.F_vs_S.8h.F_up <- 
  readLines(paste0(nc_prefix, "L.8h.F_vs_S.8h.F_upregulated", suffix))
nc_L.8h.M_vs_S.8h.M_up <- 
  readLines(paste0(nc_prefix, "L.8h.M_vs_S.8h.M_upregulated", suffix))


# format in a list
list_input <- list(
  "ComBat: L.8h.F vs S.8h.F down-regulated" = combat_L.8h.F_vs_S.8h.F_down,
  "ComBat: L.8h.M vs S.8h.M down-regulated" = combat_L.8h.M_vs_S.8h.M_down,
  "Non-ComBat: L.8h.F vs S.8h.F down-regulated" = nc_L.8h.F_vs_S.8h.F_down,
  "Non-ComBat: L.8h.M vs S.8h.M down-regulated" = nc_L.8h.M_vs_S.8h.M_down,
  "ComBat: L.8h.F vs S.8h.F up-regulated" = combat_L.8h.F_vs_S.8h.F_up,
  "ComBat: L.8h.M vs S.8h.M up-regulated" = combat_L.8h.M_vs_S.8h.M_up,
  "Non-ComBat: L.8h.F vs S.8h.F up-regulated" = nc_L.8h.F_vs_S.8h.F_up,
  "Non-ComBat: L.8h.M vs S.8h.M up-regulated" = nc_L.8h.M_vs_S.8h.M_up
)
data <- UpSetR::fromList(list_input)

# You will specify the male and female groups by their group number
group_mapping <- data.frame(
  Group_Number = 1:8,  # Group numbers (1-8)
  Gene_Set_Name = sort(colnames(data))  # Sorted gene set names
)
group_mapping

# Assign shapes based on group numbers
male_groups <- c("3","4","7","8")  # Male group numbers
female_groups <- c("1","2","5","6")  # Female group numbers

# Assign colors and shapes based on specific group numbers
upregulated_groups <- c("2","4","6","8")
downregulated_groups <- c("1","3","5","7")

# Create the basic UpSet plot (no dynamic colors or shapes) to extract n_points
dummy_upset <- ComplexUpset::upset(
  data, 
  colnames(data),  # Use column names (gene sets)
  set_sizes = ComplexUpset::upset_set_size(),  # Display set sizes
  sort_sets = FALSE,
  
  # Plot the intersection matrix with points (default shapes)
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(size = 3)  # Simple points without dynamic aesthetics
  )
)

# Ensure n_points matches the actual data
n_points <- nrow(dummy_upset$data)
print(paste("Number of points being plotted:", n_points))  # Should be 300

# Create empty vectors for colors and shapes
color_vector <- rep(NA, n_points)  # To store colors
shape_vector <- rep(NA, n_points)  # To store shapes

# Assign shapes and colors based on group numbers
for (i in 1:n_points) {
  group_value <- as.character(dummy_upset$data$group[i])
  
  # Assign shapes
  if (group_value %in% male_groups) {
    shape_vector[i] <- 22  # Filled square for Male comparisons
  } else if (group_value %in% female_groups) {
    shape_vector[i] <- 21  # Circle for Female comparisons
  }
  
  # Assign colors based on group numbers, only if the point is involved in an intersection
  if (dummy_upset$data$value[i]) {  # If the point is part of an intersection
    if (group_value %in% upregulated_groups) {
      color_vector[i] <- "red"  # Red for Up-regulated
    } else if (group_value %in% downregulated_groups) {
      color_vector[i] <- "blue"  # Blue for Down-regulated
    }
  } else {
    color_vector[i] <- NA  # No fill for points not involved in intersections
  }
}

# Ensure both vectors are of correct length
if (length(shape_vector) != n_points || length(color_vector) != n_points) {
  stop("Error: shape_vector or color_vector length does not match the number of plotted points.")
}

# Print the color_vector to ensure correctness before plotting
print(unique(color_vector))  # Ensure this contains only valid colors or NA

# Reintroduce dynamic fill color for filled shapes
upset_gene <- ComplexUpset::upset(
  data, 
  colnames(data),
  set_sizes = (
    ComplexUpset::upset_set_size() +
    geom_text(aes(label = ..count..), hjust = 1.1, stat = 'count') +
    expand_limits(y = 1000)
  ),
  
  # Prevent automatic sorting of sets
  sort_sets = FALSE,
  
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(
      ggplot2::aes(shape = factor(shape_vector), fill = color_vector),  # Fill shapes with dynamic colors
      size = 3,
      stroke = 0.45  # Keeps the outline black
    )
  ) + ggplot2::scale_shape_manual(
    name = "Sex",
    labels = c("Male", "Female"),  # Legend labels
    values = c(21, 22) # Circle=21, Filled square=22
    ) + ggplot2::scale_fill_identity(na.translate = FALSE),  # Use the exact colors in color_vector and remove NA fill
  
    base_annotations = list(
    'Intersection size' = (
      intersection_size(bar_number_threshold = 1, width = 0.5) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, 500)) +
      theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = 'black')
      )
    )
  ),
)

# Add title
upset_gene <- upset_gene + 
  ggtitle("L.8h, FDRq < 0.10, |LFC| > 0.20")

# Remove legend
upset_gene <- upset_gene + theme(legend.position = "None")

# Save temp without legend
pdf(file = "../results/compare_analyses/L.8h_sex_specific_with_and_without_combat.pdf", 
    height = 6, 
    width = 8)
upset_gene
```

```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = "../results/compare_analyses/L.8h_sex_specific_with_and_without_combat.tsv",
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

## L.24h
```{r}
# set variables
suffix <- "_FDRq_0.10_LFC_0.20.txt"
combat_prefix <- paste0("../results/all_samples_sex_specific_combat_analysis/",
                        "DEGs_with_hbb/metascape_input/")
nc_prefix <- paste0("../results/all_samples_sex_specific_analysis/",
                        "DEGs_with_hbb/metascape_input/")

# get file list
combat_L.24h.F_vs_S.24h.F_down <- 
  readLines(paste0(combat_prefix, "L.24h.F_vs_S.24h.F_downregulated", suffix))
combat_L.24h.M_vs_S.24h.M_down <- 
  readLines(paste0(combat_prefix, "L.24h.M_vs_S.24h.M_downregulated", suffix))
nc_L.24h.F_vs_S.24h.F_down <- 
  readLines(paste0(nc_prefix, "L.24h.F_vs_S.24h.F_downregulated", suffix))
nc_L.24h.M_vs_S.24h.M_down <- 
  readLines(paste0(nc_prefix, "L.24h.M_vs_S.24h.M_downregulated", suffix))
combat_L.24h.F_vs_S.24h.F_up <- 
  readLines(paste0(combat_prefix, "L.24h.F_vs_S.24h.F_upregulated", suffix))
combat_L.24h.M_vs_S.24h.M_up <- 
  readLines(paste0(combat_prefix, "L.24h.M_vs_S.24h.M_upregulated", suffix))
nc_L.24h.F_vs_S.24h.F_up <- 
  readLines(paste0(nc_prefix, "L.24h.F_vs_S.24h.F_upregulated", suffix))
nc_L.24h.M_vs_S.24h.M_up <- 
  readLines(paste0(nc_prefix, "L.24h.M_vs_S.24h.M_upregulated", suffix))


# format in a list
list_input <- list(
  "ComBat: L.24h.F vs S.24h.F down-regulated" = combat_L.24h.F_vs_S.24h.F_down,
  "ComBat: L.24h.M vs S.24h.M down-regulated" = combat_L.24h.M_vs_S.24h.M_down,
  "Non-ComBat: L.24h.F vs S.24h.F down-regulated" = nc_L.24h.F_vs_S.24h.F_down,
  "Non-ComBat: L.24h.M vs S.24h.M down-regulated" = nc_L.24h.M_vs_S.24h.M_down,
  "ComBat: L.24h.F vs S.24h.F up-regulated" = combat_L.24h.F_vs_S.24h.F_up,
  "ComBat: L.24h.M vs S.24h.M up-regulated" = combat_L.24h.M_vs_S.24h.M_up,
  "Non-ComBat: L.24h.F vs S.24h.F up-regulated" = nc_L.24h.F_vs_S.24h.F_up,
  "Non-ComBat: L.24h.M vs S.24h.M up-regulated" = nc_L.24h.M_vs_S.24h.M_up
)
data <- UpSetR::fromList(list_input)

# You will specify the male and female groups by their group number
group_mapping <- data.frame(
  Group_Number = 1:8,  # Group numbers (1-8)
  Gene_Set_Name = sort(colnames(data))  # Sorted gene set names
)
group_mapping

# Assign shapes based on group numbers
male_groups <- c("3","4","7","8")  # Male group numbers
female_groups <- c("1","2","5","6")  # Female group numbers

# Assign colors and shapes based on specific group numbers
upregulated_groups <- c("2","4","6","8")
downregulated_groups <- c("1","3","5","7")

# Create the basic UpSet plot (no dynamic colors or shapes) to extract n_points
dummy_upset <- ComplexUpset::upset(
  data, 
  colnames(data),  # Use column names (gene sets)
  set_sizes = ComplexUpset::upset_set_size(),  # Display set sizes
  sort_sets = FALSE,
  
  # Plot the intersection matrix with points (default shapes)
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(size = 3)  # Simple points without dynamic aesthetics
  )
)

# Ensure n_points matches the actual data
n_points <- nrow(dummy_upset$data)
print(paste("Number of points being plotted:", n_points))  # Should be 300

# Create empty vectors for colors and shapes
color_vector <- rep(NA, n_points)  # To store colors
shape_vector <- rep(NA, n_points)  # To store shapes

# Assign shapes and colors based on group numbers
for (i in 1:n_points) {
  group_value <- as.character(dummy_upset$data$group[i])
  
  # Assign shapes
  if (group_value %in% male_groups) {
    shape_vector[i] <- 22  # Filled square for Male comparisons
  } else if (group_value %in% female_groups) {
    shape_vector[i] <- 21  # Circle for Female comparisons
  }
  
  # Assign colors based on group numbers, only if the point is involved in an intersection
  if (dummy_upset$data$value[i]) {  # If the point is part of an intersection
    if (group_value %in% upregulated_groups) {
      color_vector[i] <- "red"  # Red for Up-regulated
    } else if (group_value %in% downregulated_groups) {
      color_vector[i] <- "blue"  # Blue for Down-regulated
    }
  } else {
    color_vector[i] <- NA  # No fill for points not involved in intersections
  }
}

# Ensure both vectors are of correct length
if (length(shape_vector) != n_points || length(color_vector) != n_points) {
  stop("Error: shape_vector or color_vector length does not match the number of plotted points.")
}

# Print the color_vector to ensure correctness before plotting
print(unique(color_vector))  # Ensure this contains only valid colors or NA

# Reintroduce dynamic fill color for filled shapes
upset_gene <- ComplexUpset::upset(
  data, 
  colnames(data),
  set_sizes = (
    ComplexUpset::upset_set_size() +
    geom_text(aes(label = ..count..), hjust = 1.1, stat = 'count') +
    expand_limits(y = 1000)
  ),
  
  # Prevent automatic sorting of sets
  sort_sets = FALSE,
  
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(
      ggplot2::aes(shape = factor(shape_vector), fill = color_vector),  # Fill shapes with dynamic colors
      size = 3,
      stroke = 0.45  # Keeps the outline black
    )
  ) + ggplot2::scale_shape_manual(
    name = "Sex",
    labels = c("Male", "Female"),  # Legend labels
    values = c(21, 22) # Circle=21, Filled square=22
    ) + ggplot2::scale_fill_identity(na.translate = FALSE),  # Use the exact colors in color_vector and remove NA fill
  
    base_annotations = list(
    'Intersection size' = (
      intersection_size(bar_number_threshold = 1, width = 0.5) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, 800)) +
      theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = 'black')
      )
    )
  ),
)

# Add title
upset_gene <- upset_gene + 
  ggtitle("L.24h, FDRq < 0.10, |LFC| > 0.20")

# Remove legend
upset_gene <- upset_gene + theme(legend.position = "None")

# Save temp without legend
pdf(file = "../results/compare_analyses/L.24h_sex_specific_with_and_without_combat.pdf", 
    height = 6, 
    width = 10)
upset_gene
``` 

```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = "../results/compare_analyses/L.24h_sex_specific_with_and_without_combat.tsv",
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

## L.7d
```{r}
# set variables
suffix <- "_FDRq_0.10_LFC_0.20.txt"
combat_prefix <- paste0("../results/all_samples_sex_specific_combat_analysis/",
                        "DEGs_with_hbb/metascape_input/")
nc_prefix <- paste0("../results/all_samples_sex_specific_analysis/",
                        "DEGs_with_hbb/metascape_input/")

# get file list
combat_L.7d.F_vs_S.7d.F_down <- 
  readLines(paste0(combat_prefix, "L.7d.F_vs_S.7d.F_downregulated", suffix))
combat_L.7d.M_vs_S.7d.M_down <- 
  readLines(paste0(combat_prefix, "L.7d.M_vs_S.7d.M_downregulated", suffix))
nc_L.7d.F_vs_S.7d.F_down <- 
  readLines(paste0(nc_prefix, "L.7d.F_vs_S.7d.F_downregulated", suffix))
nc_L.7d.M_vs_S.7d.M_down <- 
  readLines(paste0(nc_prefix, "L.7d.M_vs_S.7d.M_downregulated", suffix))
combat_L.7d.F_vs_S.7d.F_up <- 
  readLines(paste0(combat_prefix, "L.7d.F_vs_S.7d.F_upregulated", suffix))
combat_L.7d.M_vs_S.7d.M_up <- 
  readLines(paste0(combat_prefix, "L.7d.M_vs_S.7d.M_upregulated", suffix))
nc_L.7d.F_vs_S.7d.F_up <- 
  readLines(paste0(nc_prefix, "L.7d.F_vs_S.7d.F_upregulated", suffix))
nc_L.7d.M_vs_S.7d.M_up <- 
  readLines(paste0(nc_prefix, "L.7d.M_vs_S.7d.M_upregulated", suffix))


# format in a list
list_input <- list(
  "ComBat: L.7d.F vs S.7d.F down-regulated" = combat_L.7d.F_vs_S.7d.F_down,
  "ComBat: L.7d.M vs S.7d.M down-regulated" = combat_L.7d.M_vs_S.7d.M_down,
  "Non-ComBat: L.7d.F vs S.7d.F down-regulated" = nc_L.7d.F_vs_S.7d.F_down,
  "Non-ComBat: L.7d.M vs S.7d.M down-regulated" = nc_L.7d.M_vs_S.7d.M_down,
  "ComBat: L.7d.F vs S.7d.F up-regulated" = combat_L.7d.F_vs_S.7d.F_up,
  "ComBat: L.7d.M vs S.7d.M up-regulated" = combat_L.7d.M_vs_S.7d.M_up,
  "Non-ComBat: L.7d.F vs S.7d.F up-regulated" = nc_L.7d.F_vs_S.7d.F_up,
  "Non-ComBat: L.7d.M vs S.7d.M up-regulated" = nc_L.7d.M_vs_S.7d.M_up
)
data <- UpSetR::fromList(list_input)

# You will specify the male and female groups by their group number
group_mapping <- data.frame(
  Group_Number = 1:8,  # Group numbers (1-8)
  Gene_Set_Name = sort(colnames(data))  # Sorted gene set names
)
group_mapping

# Assign shapes based on group numbers
male_groups <- c("3","4","7","8")  # Male group numbers
female_groups <- c("1","2","5","6")  # Female group numbers

# Assign colors and shapes based on specific group numbers
upregulated_groups <- c("2","4","6","8")
downregulated_groups <- c("1","3","5","7")

# Create the basic UpSet plot (no dynamic colors or shapes) to extract n_points
dummy_upset <- ComplexUpset::upset(
  data, 
  colnames(data),  # Use column names (gene sets)
  set_sizes = ComplexUpset::upset_set_size(),  # Display set sizes
  sort_sets = FALSE,
  
  # Plot the intersection matrix with points (default shapes)
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(size = 3)  # Simple points without dynamic aesthetics
  )
)

# Ensure n_points matches the actual data
n_points <- nrow(dummy_upset$data)
print(paste("Number of points being plotted:", n_points))  # Should be 300

# Create empty vectors for colors and shapes
color_vector <- rep(NA, n_points)  # To store colors
shape_vector <- rep(NA, n_points)  # To store shapes

# Assign shapes and colors based on group numbers
for (i in 1:n_points) {
  group_value <- as.character(dummy_upset$data$group[i])
  
  # Assign shapes
  if (group_value %in% male_groups) {
    shape_vector[i] <- 22  # Filled square for Male comparisons
  } else if (group_value %in% female_groups) {
    shape_vector[i] <- 21  # Circle for Female comparisons
  }
  
  # Assign colors based on group numbers, only if the point is involved in an intersection
  if (dummy_upset$data$value[i]) {  # If the point is part of an intersection
    if (group_value %in% upregulated_groups) {
      color_vector[i] <- "red"  # Red for Up-regulated
    } else if (group_value %in% downregulated_groups) {
      color_vector[i] <- "blue"  # Blue for Down-regulated
    }
  } else {
    color_vector[i] <- NA  # No fill for points not involved in intersections
  }
}

# Ensure both vectors are of correct length
if (length(shape_vector) != n_points || length(color_vector) != n_points) {
  stop("Error: shape_vector or color_vector length does not match the number of plotted points.")
}

# Print the color_vector to ensure correctness before plotting
print(unique(color_vector))  # Ensure this contains only valid colors or NA

# Reintroduce dynamic fill color for filled shapes
upset_gene <- ComplexUpset::upset(
  data, 
  colnames(data),
  set_sizes = (
    ComplexUpset::upset_set_size() +
    geom_text(aes(label = ..count..), hjust = 1.1, stat = 'count') +
    expand_limits(y = 100)
  ),
  
  # Prevent automatic sorting of sets
  sort_sets = FALSE,
  
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(
      ggplot2::aes(shape = factor(shape_vector), fill = color_vector),  # Fill shapes with dynamic colors
      size = 3,
      stroke = 0.45  # Keeps the outline black
    )
  ) + ggplot2::scale_shape_manual(
    name = "Sex",
    labels = c("Male", "Female"),  # Legend labels
    values = c(21, 22) # Circle=21, Filled square=22
    ) + ggplot2::scale_fill_identity(na.translate = FALSE),  # Use the exact colors in color_vector and remove NA fill
  
    base_annotations = list(
    'Intersection size' = (
      intersection_size(bar_number_threshold = 1, width = 0.5) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, 50)) +
      theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = 'black')
      )
    )
  ),
)

# Add title
upset_gene <- upset_gene + 
  ggtitle("L.7d, FDRq < 0.10, |LFC| > 0.20")

# Remove legend
upset_gene <- upset_gene + theme(legend.position = "None")

# Save temp without legend
pdf(file = "../results/compare_analyses/L.7d_sex_specific_with_and_without_combat.pdf", 
    height = 6, 
    width = 10)
upset_gene
``` 

```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = "../results/compare_analyses/L.7d_sex_specific_with_and_without_combat.tsv",
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

## H.8h
```{r}
# set variables
suffix <- "_FDRq_0.10_LFC_0.20.txt"
combat_prefix <- paste0("../results/all_samples_sex_specific_combat_analysis/",
                        "DEGs_with_hbb/metascape_input/")
nc_prefix <- paste0("../results/all_samples_sex_specific_analysis/",
                        "DEGs_with_hbb/metascape_input/")

# get file list
combat_H.8h.F_vs_S.8h.F_down <- 
  readLines(paste0(combat_prefix, "H.8h.F_vs_S.8h.F_downregulated", suffix))
combat_H.8h.M_vs_S.8h.M_down <- 
  readLines(paste0(combat_prefix, "H.8h.M_vs_S.8h.M_downregulated", suffix))
nc_H.8h.F_vs_S.8h.F_down <- 
  readLines(paste0(nc_prefix, "H.8h.F_vs_S.8h.F_downregulated", suffix))
nc_H.8h.M_vs_S.8h.M_down <- 
  readLines(paste0(nc_prefix, "H.8h.M_vs_S.8h.M_downregulated", suffix))
combat_H.8h.F_vs_S.8h.F_up <- 
  readLines(paste0(combat_prefix, "H.8h.F_vs_S.8h.F_upregulated", suffix))
combat_H.8h.M_vs_S.8h.M_up <- 
  readLines(paste0(combat_prefix, "H.8h.M_vs_S.8h.M_upregulated", suffix))
nc_H.8h.F_vs_S.8h.F_up <- 
  readLines(paste0(nc_prefix, "H.8h.F_vs_S.8h.F_upregulated", suffix))
nc_H.8h.M_vs_S.8h.M_up <- 
  readLines(paste0(nc_prefix, "H.8h.M_vs_S.8h.M_upregulated", suffix))


# format in a list
list_input <- list(
  "ComBat: H.8h.F vs S.8h.F down-regulated" = combat_H.8h.F_vs_S.8h.F_down,
  "ComBat: H.8h.M vs S.8h.M down-regulated" = combat_H.8h.M_vs_S.8h.M_down,
  "Non-ComBat: H.8h.F vs S.8h.F down-regulated" = nc_H.8h.F_vs_S.8h.F_down,
  "Non-ComBat: H.8h.M vs S.8h.M down-regulated" = nc_H.8h.M_vs_S.8h.M_down,
  "ComBat: H.8h.F vs S.8h.F up-regulated" = combat_H.8h.F_vs_S.8h.F_up,
  "ComBat: H.8h.M vs S.8h.M up-regulated" = combat_H.8h.M_vs_S.8h.M_up,
  "Non-ComBat: H.8h.F vs S.8h.F up-regulated" = nc_H.8h.F_vs_S.8h.F_up,
  "Non-ComBat: H.8h.M vs S.8h.M up-regulated" = nc_H.8h.M_vs_S.8h.M_up
)
data <- UpSetR::fromList(list_input)

# You will specify the male and female groups by their group number
group_mapping <- data.frame(
  Group_Number = 1:8,  # Group numbers (1-8)
  Gene_Set_Name = sort(colnames(data))  # Sorted gene set names
)
group_mapping

# Assign shapes based on group numbers
male_groups <- c("3","4","7","8")  # Male group numbers
female_groups <- c("1","2","5","6")  # Female group numbers

# Assign colors and shapes based on specific group numbers
upregulated_groups <- c("2","4","6","8")
downregulated_groups <- c("1","3","5","7")

# Create the basic UpSet plot (no dynamic colors or shapes) to extract n_points
dummy_upset <- ComplexUpset::upset(
  data, 
  colnames(data),  # Use column names (gene sets)
  set_sizes = ComplexUpset::upset_set_size(),  # Display set sizes
  sort_sets = FALSE,
  
  # Plot the intersection matrix with points (default shapes)
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(size = 3)  # Simple points without dynamic aesthetics
  )
)

# Ensure n_points matches the actual data
n_points <- nrow(dummy_upset$data)
print(paste("Number of points being plotted:", n_points))  # Should be 300

# Create empty vectors for colors and shapes
color_vector <- rep(NA, n_points)  # To store colors
shape_vector <- rep(NA, n_points)  # To store shapes

# Assign shapes and colors based on group numbers
for (i in 1:n_points) {
  group_value <- as.character(dummy_upset$data$group[i])
  
  # Assign shapes
  if (group_value %in% male_groups) {
    shape_vector[i] <- 22  # Filled square for Male comparisons
  } else if (group_value %in% female_groups) {
    shape_vector[i] <- 21  # Circle for Female comparisons
  }
  
  # Assign colors based on group numbers, only if the point is involved in an intersection
  if (dummy_upset$data$value[i]) {  # If the point is part of an intersection
    if (group_value %in% upregulated_groups) {
      color_vector[i] <- "red"  # Red for Up-regulated
    } else if (group_value %in% downregulated_groups) {
      color_vector[i] <- "blue"  # Blue for Down-regulated
    }
  } else {
    color_vector[i] <- NA  # No fill for points not involved in intersections
  }
}

# Ensure both vectors are of correct length
if (length(shape_vector) != n_points || length(color_vector) != n_points) {
  stop("Error: shape_vector or color_vector length does not match the number of plotted points.")
}

# Print the color_vector to ensure correctness before plotting
print(unique(color_vector))  # Ensure this contains only valid colors or NA

# Reintroduce dynamic fill color for filled shapes
upset_gene <- ComplexUpset::upset(
  data, 
  colnames(data),
  set_sizes = (
    ComplexUpset::upset_set_size() +
    geom_text(aes(label = ..count..), hjust = 1.1, stat = 'count') +
    expand_limits(y = 400)
  ),
  
  # Prevent automatic sorting of sets
  sort_sets = FALSE,
  
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(
      ggplot2::aes(shape = factor(shape_vector), fill = color_vector),  # Fill shapes with dynamic colors
      size = 3,
      stroke = 0.45  # Keeps the outline black
    )
  ) + ggplot2::scale_shape_manual(
    name = "Sex",
    labels = c("Male", "Female"),  # Legend labels
    values = c(21, 22) # Circle=21, Filled square=22
    ) + ggplot2::scale_fill_identity(na.translate = FALSE),  # Use the exact colors in color_vector and remove NA fill
  
    base_annotations = list(
    'Intersection size' = (
      intersection_size(bar_number_threshold = 1, width = 0.5) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0,200)) +
      theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = 'black')
      )
    )
  ),
)

# Add title
upset_gene <- upset_gene + 
  ggtitle("H.8h, FDRq < 0.10, |LFC| > 0.20")

# Remove legend
upset_gene <- upset_gene + theme(legend.position = "None")

# Save temp without legend
pdf(file = "../results/compare_analyses/H.8h_sex_specific_with_and_without_combat.pdf", 
    height = 6, 
    width = 10)
upset_gene
```

```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = "../results/compare_analyses/H.8h_sex_specific_with_and_without_combat.tsv",
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

## H.24h
```{r}
# set variables
suffix <- "_FDRq_0.10_LFC_0.20.txt"
combat_prefix <- paste0("../results/all_samples_sex_specific_combat_analysis/",
                        "DEGs_with_hbb/metascape_input/")
nc_prefix <- paste0("../results/all_samples_sex_specific_analysis/",
                        "DEGs_with_hbb/metascape_input/")

# get file list
combat_H.24h.F_vs_S.24h.F_down <- 
  readLines(paste0(combat_prefix, "H.24h.F_vs_S.24h.F_downregulated", suffix))
combat_H.24h.M_vs_S.24h.M_down <- 
  readLines(paste0(combat_prefix, "H.24h.M_vs_S.24h.M_downregulated", suffix))
nc_H.24h.F_vs_S.24h.F_down <- 
  readLines(paste0(nc_prefix, "H.24h.F_vs_S.24h.F_downregulated", suffix))
nc_H.24h.M_vs_S.24h.M_down <- 
  readLines(paste0(nc_prefix, "H.24h.M_vs_S.24h.M_downregulated", suffix))
combat_H.24h.F_vs_S.24h.F_up <- 
  readLines(paste0(combat_prefix, "H.24h.F_vs_S.24h.F_upregulated", suffix))
combat_H.24h.M_vs_S.24h.M_up <- 
  readLines(paste0(combat_prefix, "H.24h.M_vs_S.24h.M_upregulated", suffix))
nc_H.24h.F_vs_S.24h.F_up <- 
  readLines(paste0(nc_prefix, "H.24h.F_vs_S.24h.F_upregulated", suffix))
nc_H.24h.M_vs_S.24h.M_up <- 
  readLines(paste0(nc_prefix, "H.24h.M_vs_S.24h.M_upregulated", suffix))


# format in a list
list_input <- list(
  "ComBat: H.24h.F vs S.24h.F down-regulated" = combat_H.24h.F_vs_S.24h.F_down,
  "ComBat: H.24h.M vs S.24h.M down-regulated" = combat_H.24h.M_vs_S.24h.M_down,
  "Non-ComBat: H.24h.F vs S.24h.F down-regulated" = nc_H.24h.F_vs_S.24h.F_down,
  "Non-ComBat: H.24h.M vs S.24h.M down-regulated" = nc_H.24h.M_vs_S.24h.M_down,
  "ComBat: H.24h.F vs S.24h.F up-regulated" = combat_H.24h.F_vs_S.24h.F_up,
  "ComBat: H.24h.M vs S.24h.M up-regulated" = combat_H.24h.M_vs_S.24h.M_up,
  "Non-ComBat: H.24h.F vs S.24h.F up-regulated" = nc_H.24h.F_vs_S.24h.F_up,
  "Non-ComBat: H.24h.M vs S.24h.M up-regulated" = nc_H.24h.M_vs_S.24h.M_up
)
data <- UpSetR::fromList(list_input)

# You will specify the male and female groups by their group number
group_mapping <- data.frame(
  Group_Number = 1:8,  # Group numbers (1-8)
  Gene_Set_Name = sort(colnames(data))  # Sorted gene set names
)
group_mapping

# Assign shapes based on group numbers
male_groups <- c("3","4","7","8")  # Male group numbers
female_groups <- c("1","2","5","6")  # Female group numbers

# Assign colors and shapes based on specific group numbers
upregulated_groups <- c("2","4","6","8")
downregulated_groups <- c("1","3","5","7")

# Create the basic UpSet plot (no dynamic colors or shapes) to extract n_points
dummy_upset <- ComplexUpset::upset(
  data, 
  colnames(data),  # Use column names (gene sets)
  set_sizes = ComplexUpset::upset_set_size(),  # Display set sizes
  sort_sets = FALSE,
  
  # Plot the intersection matrix with points (default shapes)
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(size = 3)  # Simple points without dynamic aesthetics
  )
)

# Ensure n_points matches the actual data
n_points <- nrow(dummy_upset$data)
print(paste("Number of points being plotted:", n_points))  # Should be 300

# Create empty vectors for colors and shapes
color_vector <- rep(NA, n_points)  # To store colors
shape_vector <- rep(NA, n_points)  # To store shapes

# Assign shapes and colors based on group numbers
for (i in 1:n_points) {
  group_value <- as.character(dummy_upset$data$group[i])
  
  # Assign shapes
  if (group_value %in% male_groups) {
    shape_vector[i] <- 22  # Filled square for Male comparisons
  } else if (group_value %in% female_groups) {
    shape_vector[i] <- 21  # Circle for Female comparisons
  }
  
  # Assign colors based on group numbers, only if the point is involved in an intersection
  if (dummy_upset$data$value[i]) {  # If the point is part of an intersection
    if (group_value %in% upregulated_groups) {
      color_vector[i] <- "red"  # Red for Up-regulated
    } else if (group_value %in% downregulated_groups) {
      color_vector[i] <- "blue"  # Blue for Down-regulated
    }
  } else {
    color_vector[i] <- NA  # No fill for points not involved in intersections
  }
}

# Ensure both vectors are of correct length
if (length(shape_vector) != n_points || length(color_vector) != n_points) {
  stop("Error: shape_vector or color_vector length does not match the number of plotted points.")
}

# Print the color_vector to ensure correctness before plotting
print(unique(color_vector))  # Ensure this contains only valid colors or NA

# Reintroduce dynamic fill color for filled shapes
upset_gene <- ComplexUpset::upset(
  data, 
  colnames(data),
  set_sizes = (
    ComplexUpset::upset_set_size() +
    geom_text(aes(label = ..count..), hjust = 1.1, stat = 'count') +
    expand_limits(y = 600)
  ),
  
  # Prevent automatic sorting of sets
  sort_sets = FALSE,
  
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(
      ggplot2::aes(shape = factor(shape_vector), fill = color_vector),  # Fill shapes with dynamic colors
      size = 3,
      stroke = 0.45  # Keeps the outline black
    )
  ) + ggplot2::scale_shape_manual(
    name = "Sex",
    labels = c("Male", "Female"),  # Legend labels
    values = c(21, 22) # Circle=21, Filled square=22
    ) + ggplot2::scale_fill_identity(na.translate = FALSE),  # Use the exact colors in color_vector and remove NA fill
  
    base_annotations = list(
    'Intersection size' = (
      intersection_size(bar_number_threshold = 1, width = 0.5) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0,400)) +
      theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = 'black')
      )
    )
  ),
)

# Add title
upset_gene <- upset_gene + 
  ggtitle("H.24h, FDRq < 0.10, |LFC| > 0.20")

# Remove legend
upset_gene <- upset_gene + theme(legend.position = "None")

# Save temp without legend
pdf(file = "../results/compare_analyses/H.24h_sex_specific_with_and_without_combat.pdf", 
    height = 6, 
    width = 12)
upset_gene
```

```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = "../results/compare_analyses/H.24h_sex_specific_with_and_without_combat.tsv",
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

## H.7d
```{r}
# set variables
suffix <- "_FDRq_0.10_LFC_0.20.txt"
combat_prefix <- paste0("../results/all_samples_sex_specific_combat_analysis/",
                        "DEGs_with_hbb/metascape_input/")
nc_prefix <- paste0("../results/all_samples_sex_specific_analysis/",
                        "DEGs_with_hbb/metascape_input/")

# get file list
combat_H.7d.F_vs_S.7d.F_down <- 
  readLines(paste0(combat_prefix, "H.7d.F_vs_S.7d.F_downregulated", suffix))
combat_H.7d.M_vs_S.7d.M_down <- 
  readLines(paste0(combat_prefix, "H.7d.M_vs_S.7d.M_downregulated", suffix))
nc_H.7d.F_vs_S.7d.F_down <- 
  readLines(paste0(nc_prefix, "H.7d.F_vs_S.7d.F_downregulated", suffix))
nc_H.7d.M_vs_S.7d.M_down <- 
  readLines(paste0(nc_prefix, "H.7d.M_vs_S.7d.M_downregulated", suffix))
combat_H.7d.F_vs_S.7d.F_up <- 
  readLines(paste0(combat_prefix, "H.7d.F_vs_S.7d.F_upregulated", suffix))
combat_H.7d.M_vs_S.7d.M_up <- 
  readLines(paste0(combat_prefix, "H.7d.M_vs_S.7d.M_upregulated", suffix))
nc_H.7d.F_vs_S.7d.F_up <- 
  readLines(paste0(nc_prefix, "H.7d.F_vs_S.7d.F_upregulated", suffix))
nc_H.7d.M_vs_S.7d.M_up <- 
  readLines(paste0(nc_prefix, "H.7d.M_vs_S.7d.M_upregulated", suffix))


# format in a list
list_input <- list(
  "ComBat: H.7d.F vs S.7d.F down-regulated" = combat_H.7d.F_vs_S.7d.F_down,
  "ComBat: H.7d.M vs S.7d.M down-regulated" = combat_H.7d.M_vs_S.7d.M_down,
  "Non-ComBat: H.7d.F vs S.7d.F down-regulated" = nc_H.7d.F_vs_S.7d.F_down,
  "Non-ComBat: H.7d.M vs S.7d.M down-regulated" = nc_H.7d.M_vs_S.7d.M_down,
  "ComBat: H.7d.F vs S.7d.F up-regulated" = combat_H.7d.F_vs_S.7d.F_up,
  "ComBat: H.7d.M vs S.7d.M up-regulated" = combat_H.7d.M_vs_S.7d.M_up,
  "Non-ComBat: H.7d.F vs S.7d.F up-regulated" = nc_H.7d.F_vs_S.7d.F_up,
  "Non-ComBat: H.7d.M vs S.7d.M up-regulated" = nc_H.7d.M_vs_S.7d.M_up
)
data <- UpSetR::fromList(list_input)

# You will specify the male and female groups by their group number
group_mapping <- data.frame(
  Group_Number = 1:8,  # Group numbers (1-8)
  Gene_Set_Name = sort(colnames(data))  # Sorted gene set names
)
group_mapping

# Assign shapes based on group numbers
male_groups <- c("3","4","7","8")  # Male group numbers
female_groups <- c("1","2","5","6")  # Female group numbers

# Assign colors and shapes based on specific group numbers
upregulated_groups <- c("2","4","6","8")
downregulated_groups <- c("1","3","5","7")

# Create the basic UpSet plot (no dynamic colors or shapes) to extract n_points
dummy_upset <- ComplexUpset::upset(
  data, 
  colnames(data),  # Use column names (gene sets)
  set_sizes = ComplexUpset::upset_set_size(),  # Display set sizes
  sort_sets = FALSE,
  
  # Plot the intersection matrix with points (default shapes)
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(size = 3)  # Simple points without dynamic aesthetics
  )
)

# Ensure n_points matches the actual data
n_points <- nrow(dummy_upset$data)
print(paste("Number of points being plotted:", n_points))  # Should be 300

# Create empty vectors for colors and shapes
color_vector <- rep(NA, n_points)  # To store colors
shape_vector <- rep(NA, n_points)  # To store shapes

# Assign shapes and colors based on group numbers
for (i in 1:n_points) {
  group_value <- as.character(dummy_upset$data$group[i])
  
  # Assign shapes
  if (group_value %in% male_groups) {
    shape_vector[i] <- 22  # Filled square for Male comparisons
  } else if (group_value %in% female_groups) {
    shape_vector[i] <- 21  # Circle for Female comparisons
  }
  
  # Assign colors based on group numbers, only if the point is involved in an intersection
  if (dummy_upset$data$value[i]) {  # If the point is part of an intersection
    if (group_value %in% upregulated_groups) {
      color_vector[i] <- "red"  # Red for Up-regulated
    } else if (group_value %in% downregulated_groups) {
      color_vector[i] <- "blue"  # Blue for Down-regulated
    }
  } else {
    color_vector[i] <- NA  # No fill for points not involved in intersections
  }
}

# Ensure both vectors are of correct length
if (length(shape_vector) != n_points || length(color_vector) != n_points) {
  stop("Error: shape_vector or color_vector length does not match the number of plotted points.")
}

# Print the color_vector to ensure correctness before plotting
print(unique(color_vector))  # Ensure this contains only valid colors or NA

# Reintroduce dynamic fill color for filled shapes
upset_gene <- ComplexUpset::upset(
  data, 
  colnames(data),
  set_sizes = (
    ComplexUpset::upset_set_size() +
    geom_text(aes(label = ..count..), hjust = 1.1, stat = 'count') +
    expand_limits(y = 10)
  ),
  
  # Prevent automatic sorting of sets
  sort_sets = FALSE,
  
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(
      ggplot2::aes(shape = factor(shape_vector), fill = color_vector),  # Fill shapes with dynamic colors
      size = 3,
      stroke = 0.45  # Keeps the outline black
    )
  ) + ggplot2::scale_shape_manual(
    name = "Sex",
    labels = c("Male", "Female"),  # Legend labels
    values = c(21, 22) # Circle=21, Filled square=22
    ) + ggplot2::scale_fill_identity(na.translate = FALSE),  # Use the exact colors in color_vector and remove NA fill
  
    base_annotations = list(
    'Intersection size' = (
      intersection_size(bar_number_threshold = 1, width = 0.5) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0,5)) +
      theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = 'black')
      )
    )
  ),
)

# Add title
upset_gene <- upset_gene + 
  ggtitle("H.7d, FDRq < 0.10, |LFC| > 0.20")

# Remove legend
upset_gene <- upset_gene + theme(legend.position = "None")

# Save temp without legend
pdf(file = "../results/compare_analyses/H.7d_sex_specific_with_and_without_combat.pdf", 
    height = 6, 
    width = 12)
upset_gene
```

```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = "../results/compare_analyses/H.7d_sex_specific_with_and_without_combat.tsv",
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

# Upset: both sexes, with/without ComBat
## L.8h
```{r}
# set variables
suffix <- "_FDRq_0.10_LFC_0.20.txt"
combat_prefix <- paste0("../results/all_samples_both_sexes_combat_analysis/",
                        "DEGs_with_hbb/metascape_input/")
nc_prefix <- paste0("../results/all_samples_both_sexes_analysis/",
                        "DEGs_with_hbb/metascape_input/")

# get file list
combat_L.8h_vs_S.8h_down <- 
  readLines(paste0(combat_prefix, "L.8h_vs_S.8h_downregulated", suffix))
nc_L.8h_vs_S.8h_down <- 
  readLines(paste0(nc_prefix, "L.8h_vs_S.8h_downregulated", suffix))
combat_L.8h_vs_S.8h_up <- 
  readLines(paste0(combat_prefix, "L.8h_vs_S.8h_upregulated", suffix))
nc_L.8h_vs_S.8h_up <- 
  readLines(paste0(nc_prefix, "L.8h_vs_S.8h_upregulated", suffix))


# format in a list
list_input <- list(
  "ComBat: L.8h vs S.8h down-regulated" = combat_L.8h_vs_S.8h_down,
  "Non-ComBat: L.8h vs S.8h down-regulated" = nc_L.8h_vs_S.8h_down,
  "ComBat: L.8h vs S.8h up-regulated" = combat_L.8h_vs_S.8h_up,
  "Non-ComBat: L.8h vs S.8h up-regulated" = nc_L.8h_vs_S.8h_up
)
data <- UpSetR::fromList(list_input)

# You will specify the male and female groups by their group number
group_mapping <- data.frame(
  Group_Number = 1:4,  # Group numbers (1-8)
  Gene_Set_Name = sort(colnames(data))  # Sorted gene set names
)

# Assign shapes based on group numbers
male_groups <- c()  # Male group numbers
female_groups <- c("1","2","3","4")  # Female group numbers

# Assign colors and shapes based on specific group numbers
upregulated_groups <- c("2","4")
downregulated_groups <- c("1","3")

# Create the basic UpSet plot (no dynamic colors or shapes) to extract n_points
dummy_upset <- ComplexUpset::upset(
  data, 
  colnames(data),  # Use column names (gene sets)
  set_sizes = ComplexUpset::upset_set_size(),  # Display set sizes
  sort_sets = FALSE,
  
  # Plot the intersection matrix with points (default shapes)
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(size = 3)  # Simple points without dynamic aesthetics
  )
)

# Ensure n_points matches the actual data
n_points <- nrow(dummy_upset$data)
print(paste("Number of points being plotted:", n_points))  # Should be 300

# Create empty vectors for colors and shapes
color_vector <- rep(NA, n_points)  # To store colors
shape_vector <- rep(NA, n_points)  # To store shapes

# Assign shapes and colors based on group numbers
for (i in 1:n_points) {
  group_value <- as.character(dummy_upset$data$group[i])
  
  # Assign shapes
  if (group_value %in% male_groups) {
    shape_vector[i] <- 22  # Filled square for Male comparisons
  } else if (group_value %in% female_groups) {
    shape_vector[i] <- 21  # Circle for Female comparisons
  }
  
  # Assign colors based on group numbers, only if the point is involved in an intersection
  if (dummy_upset$data$value[i]) {  # If the point is part of an intersection
    if (group_value %in% upregulated_groups) {
      color_vector[i] <- "red"  # Red for Up-regulated
    } else if (group_value %in% downregulated_groups) {
      color_vector[i] <- "blue"  # Blue for Down-regulated
    }
  } else {
    color_vector[i] <- NA  # No fill for points not involved in intersections
  }
}

# Ensure both vectors are of correct length
if (length(shape_vector) != n_points || length(color_vector) != n_points) {
  stop("Error: shape_vector or color_vector length does not match the number of plotted points.")
}

# Print the color_vector to ensure correctness before plotting
print(unique(color_vector))  # Ensure this contains only valid colors or NA

# Reintroduce dynamic fill color for filled shapes
upset_gene <- ComplexUpset::upset(
  data, 
  colnames(data),
  set_sizes = (
    ComplexUpset::upset_set_size() +
    geom_text(aes(label = ..count..), hjust = 1.1, stat = 'count') +
    expand_limits(y = 1000)
  ),
  
  # Prevent automatic sorting of sets
  sort_sets = FALSE,
  
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(
      ggplot2::aes(shape = factor(shape_vector), fill = color_vector),  # Fill shapes with dynamic colors
      size = 3,
      stroke = 0.45  # Keeps the outline black
    )
  ) + ggplot2::scale_shape_manual(
    name = "Sex",
    labels = c("Male", "Female"),  # Legend labels
    values = c(21, 22) # Circle=21, Filled square=22
    ) + ggplot2::scale_fill_identity(na.translate = FALSE),  # Use the exact colors in color_vector and remove NA fill
  
    base_annotations = list(
    'Intersection size' = (
      intersection_size(bar_number_threshold = 1, width = 0.5) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, 500)) +
      theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = 'black')
      )
    )
  ),
)

# Add title
upset_gene <- upset_gene + 
  ggtitle("L.8h, FDRq < 0.10, |LFC| > 0.20")

# Remove legend
upset_gene <- upset_gene + theme(legend.position = "None")

# Save temp without legend
pdf(file = "../results/compare_analyses/L.8h_both_sexes_with_and_without_combat.pdf", 
    height = 6, 
    width = 8)
upset_gene
```

```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = "../results/compare_analyses/L.8h_sex_specific_with_and_without_combat.tsv",
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = "../results/compare_analyses/L.8h_both_sexes_with_and_without_combat.tsv",
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

## L.24h
```{r}
# set variables
suffix <- "_FDRq_0.10_LFC_0.20.txt"
combat_prefix <- paste0("../results/all_samples_both_sexes_combat_analysis/",
                        "DEGs_with_hbb/metascape_input/")
nc_prefix <- paste0("../results/all_samples_both_sexes_analysis/",
                        "DEGs_with_hbb/metascape_input/")

# get file list
combat_L.24h_vs_S.24h_down <- 
  readLines(paste0(combat_prefix, "L.24h_vs_S.24h_downregulated", suffix))
nc_L.24h_vs_S.24h_down <- 
  readLines(paste0(nc_prefix, "L.24h_vs_S.24h_downregulated", suffix))
combat_L.24h_vs_S.24h_up <- 
  readLines(paste0(combat_prefix, "L.24h_vs_S.24h_upregulated", suffix))
nc_L.24h_vs_S.24h_up <- 
  readLines(paste0(nc_prefix, "L.24h_vs_S.24h_upregulated", suffix))


# format in a list
list_input <- list(
  "ComBat: L.24h vs S.24h down-regulated" = combat_L.24h_vs_S.24h_down,
  "Non-ComBat: L.24h vs S.24h down-regulated" = nc_L.24h_vs_S.24h_down,
  "ComBat: L.24h vs S.24h up-regulated" = combat_L.24h_vs_S.24h_up,
  "Non-ComBat: L.24h vs S.24h up-regulated" = nc_L.24h_vs_S.24h_up
)
data <- UpSetR::fromList(list_input)

# You will specify the male and female groups by their group number
group_mapping <- data.frame(
  Group_Number = 1:4,  # Group numbers (1-8)
  Gene_Set_Name = sort(colnames(data))  # Sorted gene set names
)

# Assign shapes based on group numbers
male_groups <- c()  # Male group numbers
female_groups <- c("1","2","3","4")  # Female group numbers

# Assign colors and shapes based on specific group numbers
upregulated_groups <- c("2","4")
downregulated_groups <- c("1","3")

# Create the basic UpSet plot (no dynamic colors or shapes) to extract n_points
dummy_upset <- ComplexUpset::upset(
  data, 
  colnames(data),  # Use column names (gene sets)
  set_sizes = ComplexUpset::upset_set_size(),  # Display set sizes
  sort_sets = FALSE,
  
  # Plot the intersection matrix with points (default shapes)
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(size = 3)  # Simple points without dynamic aesthetics
  )
)

# Ensure n_points matches the actual data
n_points <- nrow(dummy_upset$data)
print(paste("Number of points being plotted:", n_points))  # Should be 300

# Create empty vectors for colors and shapes
color_vector <- rep(NA, n_points)  # To store colors
shape_vector <- rep(NA, n_points)  # To store shapes

# Assign shapes and colors based on group numbers
for (i in 1:n_points) {
  group_value <- as.character(dummy_upset$data$group[i])
  
  # Assign shapes
  if (group_value %in% male_groups) {
    shape_vector[i] <- 22  # Filled square for Male comparisons
  } else if (group_value %in% female_groups) {
    shape_vector[i] <- 21  # Circle for Female comparisons
  }
  
  # Assign colors based on group numbers, only if the point is involved in an intersection
  if (dummy_upset$data$value[i]) {  # If the point is part of an intersection
    if (group_value %in% upregulated_groups) {
      color_vector[i] <- "red"  # Red for Up-regulated
    } else if (group_value %in% downregulated_groups) {
      color_vector[i] <- "blue"  # Blue for Down-regulated
    }
  } else {
    color_vector[i] <- NA  # No fill for points not involved in intersections
  }
}

# Ensure both vectors are of correct length
if (length(shape_vector) != n_points || length(color_vector) != n_points) {
  stop("Error: shape_vector or color_vector length does not match the number of plotted points.")
}

# Print the color_vector to ensure correctness before plotting
print(unique(color_vector))  # Ensure this contains only valid colors or NA

# Reintroduce dynamic fill color for filled shapes
upset_gene <- ComplexUpset::upset(
  data, 
  colnames(data),
  set_sizes = (
    ComplexUpset::upset_set_size() +
    geom_text(aes(label = ..count..), hjust = 1.1, stat = 'count') +
    expand_limits(y = 1000)
  ),
  
  # Prevent automatic sorting of sets
  sort_sets = FALSE,
  
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(
      ggplot2::aes(shape = factor(shape_vector), fill = color_vector),  # Fill shapes with dynamic colors
      size = 3,
      stroke = 0.45  # Keeps the outline black
    )
  ) + ggplot2::scale_shape_manual(
    name = "Sex",
    labels = c("Male", "Female"),  # Legend labels
    values = c(21, 22) # Circle=21, Filled square=22
    ) + ggplot2::scale_fill_identity(na.translate = FALSE),  # Use the exact colors in color_vector and remove NA fill
  
    base_annotations = list(
    'Intersection size' = (
      intersection_size(bar_number_threshold = 1, width = 0.5) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, 500)) +
      theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = 'black')
      )
    )
  ),
)

# Add title
upset_gene <- upset_gene + 
  ggtitle("L.24h, FDRq < 0.10, |LFC| > 0.20")

# Remove legend
upset_gene <- upset_gene + theme(legend.position = "None")

# Save temp without legend
pdf(file = "../results/compare_analyses/L.24h_both_sexes_with_and_without_combat.pdf", 
    height = 6, 
    width = 8)
upset_gene
```

```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = "../results/compare_analyses/L.24h_both_sexes_with_and_without_combat.tsv",
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

## L.7d
```{r}
# set variables
suffix <- "_FDRq_0.10_LFC_0.20.txt"
combat_prefix <- paste0("../results/all_samples_both_sexes_combat_analysis/",
                        "DEGs_with_hbb/metascape_input/")
nc_prefix <- paste0("../results/all_samples_both_sexes_analysis/",
                        "DEGs_with_hbb/metascape_input/")

# get file list
combat_L.7d_vs_S.7d_down <- 
  readLines(paste0(combat_prefix, "L.7d_vs_S.7d_downregulated", suffix))
nc_L.7d_vs_S.7d_down <- 
  readLines(paste0(nc_prefix, "L.7d_vs_S.7d_downregulated", suffix))
combat_L.7d_vs_S.7d_up <- 
  readLines(paste0(combat_prefix, "L.7d_vs_S.7d_upregulated", suffix))
nc_L.7d_vs_S.7d_up <- 
  readLines(paste0(nc_prefix, "L.7d_vs_S.7d_upregulated", suffix))


# format in a list
list_input <- list(
  "ComBat: L.7d vs S.7d down-regulated" = combat_L.7d_vs_S.7d_down,
  "Non-ComBat: L.7d vs S.7d down-regulated" = nc_L.7d_vs_S.7d_down,
  "ComBat: L.7d vs S.7d up-regulated" = combat_L.7d_vs_S.7d_up,
  "Non-ComBat: L.7d vs S.7d up-regulated" = nc_L.7d_vs_S.7d_up
)
data <- UpSetR::fromList(list_input)

# You will specify the male and female groups by their group number
group_mapping <- data.frame(
  Group_Number = 1:4,  # Group numbers (1-8)
  Gene_Set_Name = sort(colnames(data))  # Sorted gene set names
)

# Assign shapes based on group numbers
male_groups <- c()  # Male group numbers
female_groups <- c("1","2","3","4")  # Female group numbers

# Assign colors and shapes based on specific group numbers
upregulated_groups <- c("2","4")
downregulated_groups <- c("1","3")

# Create the basic UpSet plot (no dynamic colors or shapes) to extract n_points
dummy_upset <- ComplexUpset::upset(
  data, 
  colnames(data),  # Use column names (gene sets)
  set_sizes = ComplexUpset::upset_set_size(),  # Display set sizes
  sort_sets = FALSE,
  
  # Plot the intersection matrix with points (default shapes)
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(size = 3)  # Simple points without dynamic aesthetics
  )
)

# Ensure n_points matches the actual data
n_points <- nrow(dummy_upset$data)
print(paste("Number of points being plotted:", n_points))  # Should be 300

# Create empty vectors for colors and shapes
color_vector <- rep(NA, n_points)  # To store colors
shape_vector <- rep(NA, n_points)  # To store shapes

# Assign shapes and colors based on group numbers
for (i in 1:n_points) {
  group_value <- as.character(dummy_upset$data$group[i])
  
  # Assign shapes
  if (group_value %in% male_groups) {
    shape_vector[i] <- 22  # Filled square for Male comparisons
  } else if (group_value %in% female_groups) {
    shape_vector[i] <- 21  # Circle for Female comparisons
  }
  
  # Assign colors based on group numbers, only if the point is involved in an intersection
  if (dummy_upset$data$value[i]) {  # If the point is part of an intersection
    if (group_value %in% upregulated_groups) {
      color_vector[i] <- "red"  # Red for Up-regulated
    } else if (group_value %in% downregulated_groups) {
      color_vector[i] <- "blue"  # Blue for Down-regulated
    }
  } else {
    color_vector[i] <- NA  # No fill for points not involved in intersections
  }
}

# Ensure both vectors are of correct length
if (length(shape_vector) != n_points || length(color_vector) != n_points) {
  stop("Error: shape_vector or color_vector length does not match the number of plotted points.")
}

# Print the color_vector to ensure correctness before plotting
print(unique(color_vector))  # Ensure this contains only valid colors or NA

# Reintroduce dynamic fill color for filled shapes
upset_gene <- ComplexUpset::upset(
  data, 
  colnames(data),
  set_sizes = (
    ComplexUpset::upset_set_size() +
    geom_text(aes(label = ..count..), hjust = 1.1, stat = 'count') +
    expand_limits(y = 1000)
  ),
  
  # Prevent automatic sorting of sets
  sort_sets = FALSE,
  
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(
      ggplot2::aes(shape = factor(shape_vector), fill = color_vector),  # Fill shapes with dynamic colors
      size = 3,
      stroke = 0.45  # Keeps the outline black
    )
  ) + ggplot2::scale_shape_manual(
    name = "Sex",
    labels = c("Male", "Female"),  # Legend labels
    values = c(21, 22) # Circle=21, Filled square=22
    ) + ggplot2::scale_fill_identity(na.translate = FALSE),  # Use the exact colors in color_vector and remove NA fill
  
    base_annotations = list(
    'Intersection size' = (
      intersection_size(bar_number_threshold = 1, width = 0.5) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, 500)) +
      theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = 'black')
      )
    )
  ),
)

# Add title
upset_gene <- upset_gene + 
  ggtitle("L.7d, FDRq < 0.10, |LFC| > 0.20")

# Remove legend
upset_gene <- upset_gene + theme(legend.position = "None")

# Save temp without legend
pdf(file = "../results/compare_analyses/L.7d_both_sexes_with_and_without_combat.pdf", 
    height = 6, 
    width = 8)
upset_gene
```

```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = "../results/compare_analyses/L.7d_both_sexes_with_and_without_combat.tsv",
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```


## H.24h
```{r}
# set variables
suffix <- "_FDRq_0.10_LFC_0.20.txt"
combat_prefix <- paste0("../results/all_samples_both_sexes_combat_analysis/",
                        "DEGs_with_hbb/metascape_input/")
nc_prefix <- paste0("../results/all_samples_both_sexes_analysis/",
                        "DEGs_with_hbb/metascape_input/")

# get file list
combat_H.24h_vs_S.24h_down <- 
  readLines(paste0(combat_prefix, "H.24h_vs_S.24h_downregulated", suffix))
nc_H.24h_vs_S.24h_down <- 
  readLines(paste0(nc_prefix, "H.24h_vs_S.24h_downregulated", suffix))
combat_H.24h_vs_S.24h_up <- 
  readLines(paste0(combat_prefix, "H.24h_vs_S.24h_upregulated", suffix))
nc_H.24h_vs_S.24h_up <- 
  readLines(paste0(nc_prefix, "H.24h_vs_S.24h_upregulated", suffix))


# format in a list
list_input <- list(
  "ComBat: H.24h vs S.24h down-regulated" = combat_H.24h_vs_S.24h_down,
  "Non-ComBat: H.24h vs S.24h down-regulated" = nc_H.24h_vs_S.24h_down,
  "ComBat: H.24h vs S.24h up-regulated" = combat_H.24h_vs_S.24h_up,
  "Non-ComBat: H.24h vs S.24h up-regulated" = nc_H.24h_vs_S.24h_up
)
data <- UpSetR::fromList(list_input)

# You will specify the male and female groups by their group number
group_mapping <- data.frame(
  Group_Number = 1:4,  # Group numbers (1-8)
  Gene_Set_Name = sort(colnames(data))  # Sorted gene set names
)

# Assign shapes based on group numbers
male_groups <- c()  # Male group numbers
female_groups <- c("1","2","3","4")  # Female group numbers

# Assign colors and shapes based on specific group numbers
upregulated_groups <- c("2","4")
downregulated_groups <- c("1","3")

# Create the basic UpSet plot (no dynamic colors or shapes) to extract n_points
dummy_upset <- ComplexUpset::upset(
  data, 
  colnames(data),  # Use column names (gene sets)
  set_sizes = ComplexUpset::upset_set_size(),  # Display set sizes
  sort_sets = FALSE,
  
  # Plot the intersection matrix with points (default shapes)
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(size = 3)  # Simple points without dynamic aesthetics
  )
)

# Ensure n_points matches the actual data
n_points <- nrow(dummy_upset$data)
print(paste("Number of points being plotted:", n_points))  # Should be 300

# Create empty vectors for colors and shapes
color_vector <- rep(NA, n_points)  # To store colors
shape_vector <- rep(NA, n_points)  # To store shapes

# Assign shapes and colors based on group numbers
for (i in 1:n_points) {
  group_value <- as.character(dummy_upset$data$group[i])
  
  # Assign shapes
  if (group_value %in% male_groups) {
    shape_vector[i] <- 22  # Filled square for Male comparisons
  } else if (group_value %in% female_groups) {
    shape_vector[i] <- 21  # Circle for Female comparisons
  }
  
  # Assign colors based on group numbers, only if the point is involved in an intersection
  if (dummy_upset$data$value[i]) {  # If the point is part of an intersection
    if (group_value %in% upregulated_groups) {
      color_vector[i] <- "red"  # Red for Up-regulated
    } else if (group_value %in% downregulated_groups) {
      color_vector[i] <- "blue"  # Blue for Down-regulated
    }
  } else {
    color_vector[i] <- NA  # No fill for points not involved in intersections
  }
}

# Ensure both vectors are of correct length
if (length(shape_vector) != n_points || length(color_vector) != n_points) {
  stop("Error: shape_vector or color_vector length does not match the number of plotted points.")
}

# Print the color_vector to ensure correctness before plotting
print(unique(color_vector))  # Ensure this contains only valid colors or NA

# Reintroduce dynamic fill color for filled shapes
upset_gene <- ComplexUpset::upset(
  data, 
  colnames(data),
  set_sizes = (
    ComplexUpset::upset_set_size() +
    geom_text(aes(label = ..count..), hjust = 1.1, stat = 'count') +
    expand_limits(y = 1000)
  ),
  
  # Prevent automatic sorting of sets
  sort_sets = FALSE,
  
  matrix = ComplexUpset::intersection_matrix(
    ggplot2::geom_point(
      ggplot2::aes(shape = factor(shape_vector), fill = color_vector),  # Fill shapes with dynamic colors
      size = 3,
      stroke = 0.45  # Keeps the outline black
    )
  ) + ggplot2::scale_shape_manual(
    name = "Sex",
    labels = c("Male", "Female"),  # Legend labels
    values = c(21, 22) # Circle=21, Filled square=22
    ) + ggplot2::scale_fill_identity(na.translate = FALSE),  # Use the exact colors in color_vector and remove NA fill
  
    base_annotations = list(
    'Intersection size' = (
      intersection_size(bar_number_threshold = 1, width = 0.5) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = c(0, 500)) +
      theme(
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = 'black')
      )
    )
  ),
)

# Add title
upset_gene <- upset_gene + 
  ggtitle("H.24h, FDRq < 0.10, |LFC| > 0.20")

# Remove legend
upset_gene <- upset_gene + theme(legend.position = "None")

# Save temp without legend
pdf(file = "../results/compare_analyses/H.24h_both_sexes_with_and_without_combat.pdf", 
    height = 6, 
    width = 8)
upset_gene
```

```{r}
# convert the gene sets into a binary matrix (like UpSet plot input)
upset_data <- UpSetR::fromList(list_input)

# call the function to extract the exclusive intersection genes
exclusive_intersection_results <- get_exclusive_intersection_genes(upset_data, list_input)

# convert the list to a data.frame with two columns: Total genes and gene list
exclusive_intersection_df <- do.call(rbind, lapply(names(exclusive_intersection_results), function(x) {
  data.frame(
    Intersection = x,
    Total_Genes = exclusive_intersection_results[[x]]$total_genes,
    Gene_List = exclusive_intersection_results[[x]]$gene_list,
    stringsAsFactors = FALSE
  )
}))
exclusive_intersection_df <- 
  exclusive_intersection_df[order(exclusive_intersection_df$Total_Genes, decreasing = TRUE),]

# save
write.table(x = exclusive_intersection_df,
            file = "../results/compare_analyses/H.24h_both_sexes_with_and_without_combat.tsv",
            sep = "\t",
            quote = FALSE,
            col.names = TRUE,
            row.names = FALSE)
```

# Heatmaps
## High dose
### Sex specific table
```{r}
# get files
path <- "../results/all_samples_sex_specific_analysis/DEGs_with_hbb/DEG_tables"
files <- list.files(path)
files <- files[grepl("H", files)]
files <- files[grepl("FDRq_1.00_LFC_0.00.tsv", files)]
row_names <- gsub("_FDRq_1.00_LFC_0.00.tsv", "", files)

# loop through DEG tables and create logFC master table
ssLfc <- data.frame()
for (i in files) {
  print(i)
  df <- read.delim2(file = paste0(path, "/", i))
  df <- df[,c("gene_name_unique","logFC")]
  rownames(df) <- df$gene_name_unique
  df$gene_name_unique <- NULL
  df <- t(df)
  ssLfc <- gtools::smartbind(ssLfc, df)
}

# set row names
rownames(ssLfc) <- row_names

# reorder
ssLfc <- ssLfc[c("H.8h.F_vs_S.8h.F", "H.8h.M_vs_S.8h.M",
                       "H.24h.F_vs_S.24h.F", "H.24h.M_vs_S.24h.M",
                       "H.7d.F_vs_S.7d.F", "H.7d.M_vs_S.7d.M"),]
ssLfc <- t(ssLfc)

# subset
ssLfc2 <- ssLfc[1:50,]
ssLfc2 <- as.data.frame(ssLfc2)
gene_names <- rownames(ssLfc2)
ssLfc2 <- as.data.frame(lapply(ssLfc2, as.numeric))
rownames(ssLfc2) <- gene_names

# cleanup
ssLfc <- ssLfc2
remove(ssLfc2)
```

### Both sexes table
```{r}
# get files
path <- "../results/all_samples_both_sexes_analysis/DEGs_with_hbb/DEG_tables"
files <- list.files(path)
files <- files[grepl("H", files)]
files <- files[grepl("FDRq_1.00_LFC_0.00.tsv", files)]
row_names <- gsub("_FDRq_1.00_LFC_0.00.tsv", "", files)

# loop through DEG tables and create logFC master table
bsLfc <- data.frame()
for (i in files) {
  print(i)
  df <- read.delim2(file = paste0(path, "/", i))
  df <- df[,c("gene_name_unique","logFC")]
  rownames(df) <- df$gene_name_unique
  df$gene_name_unique <- NULL
  df <- t(df)
  bsLfc <- gtools::smartbind(bsLfc, df)
}

# set row names
rownames(bsLfc) <- row_names

# reorder
bsLfc <- bsLfc[c("H.8h_vs_S.8h", "H.24h_vs_S.24h", "H.7d_vs_S.7d"),]
bsLfc <- t(bsLfc)

# check
table(gene_names %in% rownames(bsLfc))

# subset
bsLfc2 <- bsLfc[gene_names,]
bsLfc2 <- as.data.frame(bsLfc2)
gene_names2 <- rownames(bsLfc2)
bsLfc2 <- as.data.frame(lapply(bsLfc2, as.numeric))
rownames(bsLfc2) <- gene_names2

# cleanup
bsLfc <- bsLfc2
remove(bsLfc2,df)
```

### Plot
```{r}
# set heatmap colors, these are the same for both heatmaps
paletteLength <- 100
myColor <- colorRampPalette(c("blue", "white", "red"))(paletteLength)
myBreaks <- c(seq(min(bsLfc,ssLfc), 0, length.out = ceiling(paletteLength/2) + 1), 
              seq(max(bsLfc,ssLfc) / paletteLength, max(bsLfc,ssLfc), length.out = floor(paletteLength/2)))

######################### sex specific heatmap ################################

# set heatmap annotation
ss_ann_colors <- list(sex = c(female = "gold", male = "purple"),
                  time_point = c(`8 hours` = "green2", `24 hours` = "cornflowerblue",`7 days` = "tomato2"))
ss_meta <- data.frame(sex = rep(c("female", "male"), 3),
                   time_point = c("8 hours","8 hours","24 hours", "24 hours","7 days", "7 days"))
rownames(ss_meta) <- colnames(ssLfc)

# plot heatmap
h1 <- pheatmap::pheatmap(
  ssLfc,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  annotation_col = ss_meta,
  annotation_colors = ss_ann_colors,
  color = myColor,
  breaks = myBreaks)

######################### sex specific heatmap ################################

# set heatmap annotation
bs_ann_colors <- list(
  sex = c(both = "grey"),
  time_point = c(`8 hours` = "green2", `24 hours` = "cornflowerblue", `7 days` = "tomato2"))
bs_meta <- data.frame(sex = rep("both",3), time_point = c("8 hours", "24 hours", "7 days"))
rownames(bs_meta) <- colnames(bsLfc)

# plot heatmap
h2 <- pheatmap::pheatmap(
  bsLfc,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  annotation_col = bs_meta,
  annotation_colors = bs_ann_colors,
  color = myColor,
  breaks = myBreaks)
```

### Grid
```{r}
# Extract gtables directly from pheatmap objects
h1_grob <- h1$gtable
h2_grob <- h2$gtable

# save
path <- "../results/figures/high_dose_LFC_heatmap.pdf"
pdf(path, width = 10, height = 10)

# Arrange the grobs with custom widths
grid.arrange(h2_grob, h1_grob, ncol = 2, widths = c(2, 3))
```

# 3D LogFC Correaltion
## Low dose
```{r low_dose_3D}
# read tables
low.8h <- read.delim(paste0(out, model, "DEG_tables/psilo.low.8h_vs_sal.8h_FDRq_1.00.tsv"))
low.24h <- read.delim(paste0(out, model, "DEG_tables/psilo.low.24h_vs_sal.24h_FDRq_1.00.tsv"))
low.7d <- read.delim(paste0(out, model, "DEG_tables/psilo.low.7d_vs_sal.7d_FDRq_1.00.tsv"))

# extract logFC
df <- data.frame(genes = low.8h$gene_name_unique,
                 low.8h = low.8h$logFC,
                 low.24h = low.24h$logFC,
                 low.7d = low.7d$logFC)
rownames(df) <- df$genes

# direction, 1 = upreg, 2 = downreg, 3 = other
df$direction <- ""
for (gene in df$genes) {
  if((df[gene,"low.8h"] > 0) && (df[gene,"low.24h"] > 0) && (df[gene,"low.7d"] > 0)) {
    df[gene, "direction"] <- "up-up-up"
  } else if ((df[gene,"low.8h"] < 0) && (df[gene,"low.24h"] < 0) && (df[gene,"low.7d"] < 0)) {
    df[gene, "direction"] <- "down-down-down"
  } else if((df[gene,"low.8h"] < 0) && (df[gene,"low.24h"] < 0) && (df[gene,"low.7d"] > 0)) {
    df[gene, "direction"] <- "down-down-up"
  } else if((df[gene,"low.8h"] > 0) && (df[gene,"low.24h"] > 0) && (df[gene,"low.7d"] < 0)) {
    df[gene, "direction"] <- "up-up-down"
  } else if((df[gene,"low.8h"] < 0) && (df[gene,"low.24h"] > 0) && (df[gene,"low.7d"] < 0)) {
    df[gene, "direction"] <- "down-up-down"
  } else if((df[gene,"low.8h"] < 0) && (df[gene,"low.24h"] > 0) && (df[gene,"low.7d"] > 0)) {
    df[gene, "direction"] <- "down-up-up"
  } else if((df[gene,"low.8h"] < 0) && (df[gene,"low.24h"] < 0) && (df[gene,"low.7d"] > 0)) {
    df[gene, "direction"] <- "down-down-up"
  } else if((df[gene,"low.8h"] < 0) && (df[gene,"low.24h"] < 0) && (df[gene,"low.7d"] > 0)) {
    df[gene, "direction"] <- "down-down-up"
  } else if((df[gene,"low.8h"] > 0) && (df[gene,"low.24h"] < 0) && (df[gene,"low.7d"] > 0)) {
    df[gene, "direction"] <- "up-down-up"
  } else if((df[gene,"low.8h"] > 0) && (df[gene,"low.24h"] < 0) && (df[gene,"low.7d"] < 0)) {
    df[gene, "direction"] <- "up-down-down"
  } else {
    df[gene, "direction"] <- "other"
  }
}
df$direction <- factor(df$direction)

# plot
fig1 <- plot_ly(df, 
                x = ~low.8h, 
                y = ~low.24h, 
                z = ~low.7d, 
                color = ~direction, 
                colors = c("blue","purple","yellow","orange","green",
                                 "gray","pink","red"),
                text = ~genes) %>%
  add_markers() %>% 
  layout(legend=list(title=list(text='<b> Low: 8h vs 24h vs 7d </b>')))
fig1
```

## High dose
```{r high_dose_3D}
# read tables
high.8h <- read.delim(paste0(out, model, "DEG_tables/psilo.high.8h_vs_sal.8h_FDRq_1.00.tsv"))
high.24h <- read.delim(paste0(out, model, "DEG_tables/psilo.high.24h_vs_sal.24h_FDRq_1.00.tsv"))
high.7d <- read.delim(paste0(out, model, "DEG_tables/psilo.high.7d_vs_sal.7d_FDRq_1.00.tsv"))

# extract logFC
df <- data.frame(genes = high.8h$gene_name_unique,
                 high.8h = high.8h$logFC,
                 high.24h = high.24h$logFC,
                 high.7d = high.7d$logFC)
rownames(df) <- df$genes

# direction, 1 = upreg, 2 = downreg, 3 = other
df$direction <- ""
for (gene in df$genes) {
  if((df[gene,"high.8h"] > 0) && (df[gene,"high.24h"] > 0) && (df[gene,"high.7d"] > 0)) {
    df[gene, "direction"] <- "up-up-up"
  } else if ((df[gene,"high.8h"] < 0) && (df[gene,"high.24h"] < 0) && (df[gene,"high.7d"] < 0)) {
    df[gene, "direction"] <- "down-down-down"
  } else if((df[gene,"high.8h"] < 0) && (df[gene,"high.24h"] < 0) && (df[gene,"high.7d"] > 0)) {
    df[gene, "direction"] <- "down-down-up"
  } else if((df[gene,"high.8h"] > 0) && (df[gene,"high.24h"] > 0) && (df[gene,"high.7d"] < 0)) {
    df[gene, "direction"] <- "up-up-down"
  } else if((df[gene,"high.8h"] < 0) && (df[gene,"high.24h"] > 0) && (df[gene,"high.7d"] < 0)) {
    df[gene, "direction"] <- "down-up-down"
  } else if((df[gene,"high.8h"] < 0) && (df[gene,"high.24h"] > 0) && (df[gene,"high.7d"] > 0)) {
    df[gene, "direction"] <- "down-up-up"
  } else if((df[gene,"high.8h"] < 0) && (df[gene,"high.24h"] < 0) && (df[gene,"high.7d"] > 0)) {
    df[gene, "direction"] <- "down-down-up"
  } else if((df[gene,"high.8h"] < 0) && (df[gene,"high.24h"] < 0) && (df[gene,"high.7d"] > 0)) {
    df[gene, "direction"] <- "down-down-up"
  } else if((df[gene,"high.8h"] > 0) && (df[gene,"high.24h"] < 0) && (df[gene,"high.7d"] > 0)) {
    df[gene, "direction"] <- "up-down-up"
  } else if((df[gene,"high.8h"] > 0) && (df[gene,"high.24h"] < 0) && (df[gene,"high.7d"] < 0)) {
    df[gene, "direction"] <- "up-down-down"
  } else {
    df[gene, "direction"] <- "other"
  }
}
df$direction <- factor(df$direction)

# plot
fig2 <- plot_ly(df, 
                x = ~high.8h, 
                y = ~high.24h, 
                z = ~high.7d, 
                color = ~direction, 
                colors = c("blue","purple","yellow","orange","green",
                                 "gray","pink","red"),
                text = ~genes) %>%
  add_markers() %>% 
  layout(legend=list(title=list(text='<b> High: 8h vs 24h vs 7d </b>')))
fig2
```

# 2D Corrrelation
```{r}
# Function to create the correlation plot
createCorrelationPlot <- function(group1, group2, group1_file, group2_file, fdrq_cutoff) {
  
  # Read data
  data1 <- read.table(group1_file, header = TRUE)
  data2 <- read.table(group2_file, header = TRUE)
  
  # Subset data based on adjusted p-value
  data1 <- data1[data1$adj.P.Val < fdrq_cutoff,]
  data2 <- data2[data2$adj.P.Val < fdrq_cutoff,]
  
  # Find shared genes and subset
  shared_genes <- data1$gene_name_unique[data1$gene_name_unique %in% data2$gene_name_unique]
  shared_genes <- shared_genes[!duplicated(shared_genes)]
  data1 <- data1[data1$gene_name_unique %in% shared_genes,]
  data2 <- data2[data2$gene_name_unique %in% shared_genes,]
  shared <- left_join(data1, data2, by = "gene_name_unique")[, c(1, 2, 9, 15, 17, 24, 30)]
  
  # Remove rows with NA values in logFC.x or logFC.y
  shared <- shared %>%
    filter(!is.na(logFC.x) & !is.na(logFC.y))
  
  # Ensure there are enough data points
  if (nrow(shared) < 2) {
    print("Not enough data to plot")
    return(NULL)  # Not enough data to plot
  }
  
  # Spearman's correlation
  res <- cor.test(shared$logFC.x, shared$logFC.y, method = "spearman")
  p_value <- round(res$p.value, 5)
  rho_value <- round(res$estimate, 3)
  
  # Fit a manual linear model
  lm_fit <- lm(logFC.y ~ logFC.x, data = shared)
  slope <- coef(lm_fit)[2]
  intercept <- coef(lm_fit)[1]
  
  # Calculate the limits for the plot based on the range of the data
  x_limits <- c(min(shared$logFC.x, na.rm = TRUE), max(shared$logFC.x, na.rm = TRUE))
  y_limits <- c(min(shared$logFC.y, na.rm = TRUE), max(shared$logFC.y, na.rm = TRUE))
  
  # Create the plot and manually add the regression line
  p <- ggplot(data = shared, 
              aes(x = logFC.x, y = logFC.y, text = paste(gene_name_unique))) +
    # Restore the pink and blue color annotations
    annotate("rect",
             xmin = 0,
             xmax = max(shared$logFC.x),
             ymin = max(shared$logFC.y),
             ymax = 0,
             fill = "lightpink3",
             alpha = .5) +  
    annotate("rect",
             xmin = 0,
             xmax = min(shared$logFC.x),
             ymin = 0,
             ymax = min(shared$logFC.y),
             fill = "cadetblue3",
             alpha = .5) +
    geom_point(size = 2) +
    theme_bw() +
    scale_x_continuous(limits = x_limits) +  # Set x-axis limits
    scale_y_continuous(limits = y_limits) +  # Set y-axis limits
    geom_abline(slope = slope, intercept = intercept, color = "gray", 
                size = 0.8, linetype = "dashed") +  # Make the line thinner and dashed
    annotate("text",
             x = 0,
             y = 0,
             label = paste0("rho = ", rho_value),
             parse = TRUE,
             color = "red",
             size = 5) +
    labs(x = paste0(group1, " LogFC"),
         y = paste0(group2, " LogFC"))
  
  return(ggplotly(p))
}
```

```{r}
# user defined variables
path1 <- paste0("../results/all_samples_both_sexes_analysis/DEGs_with_hbb/",
                "DEG_tables/H.7d_vs_S.7d_FDRq_1.00_LFC_0.00.tsv")
path2 <- paste0("../results/all_samples_both_sexes_combat_analysis/DEGs_with_hbb/",
                "DEG_tables/H.7d_vs_S.7d_FDRq_1.00_LFC_0.00.tsv")

createCorrelationPlot(group1 = "Non-ComBat: H.7d vs S.7d",
                      group2 = "ComBat: H.7d vs S.7d",
                      group1_file = path1,
                      group2_file = path2,
                      fdrq_cutoff = 1)
```

# Line graph
## Low dose
```{r low_dose_line}
# read tables
low.8h <- read.delim(paste0(out, model, "DEG_tables/psilo.low.8h_vs_sal.8h_FDRq_1.00.tsv"))
low.24h <- read.delim(paste0(out, model, "DEG_tables/psilo.low.24h_vs_sal.24h_FDRq_1.00.tsv"))
low.7d <- read.delim(paste0(out, model, "DEG_tables/psilo.low.7d_vs_sal.7d_FDRq_1.00.tsv"))

# extract logFC
df <- data.frame(genes = low.8h$gene_name_unique,
                 low.8h = low.8h$logFC,
                 low.24h = low.24h$logFC,
                 low.7d = low.7d$logFC)
rownames(df) <- df$genes

# find absolute diff
df$diff_8h_vs_24h <- abs(df$low.8h - df$low.24h)
df$diff_24h_vs_7d <- abs(df$low.24h - df$low.7d)
df <- df[df$diff_24h_vs_7d > 0.2,]
df <- df[df$diff_8h_vs_24h > 0.2,]

# direction, 1 = upreg, 2 = downreg, 3 = other
df$direction <- ""
for (gene in df$genes) {
  if((df[gene,"low.8h"] > 0) && (df[gene,"low.24h"] > 0) && (df[gene,"low.7d"] > 0)) {
    df[gene, "direction"] <- "up-up-up"
  } else if ((df[gene,"low.8h"] < 0) && (df[gene,"low.24h"] < 0) && (df[gene,"low.7d"] < 0)) {
    df[gene, "direction"] <- "down-down-down"
  } else if((df[gene,"low.8h"] < 0) && (df[gene,"low.24h"] < 0) && (df[gene,"low.7d"] > 0)) {
    df[gene, "direction"] <- "down-down-up"
  } else if((df[gene,"low.8h"] > 0) && (df[gene,"low.24h"] > 0) && (df[gene,"low.7d"] < 0)) {
    df[gene, "direction"] <- "up-up-down"
  } else if((df[gene,"low.8h"] < 0) && (df[gene,"low.24h"] > 0) && (df[gene,"low.7d"] < 0)) {
    df[gene, "direction"] <- "down-up-down"
  } else if((df[gene,"low.8h"] < 0) && (df[gene,"low.24h"] > 0) && (df[gene,"low.7d"] > 0)) {
    df[gene, "direction"] <- "down-up-up"
  } else if((df[gene,"low.8h"] < 0) && (df[gene,"low.24h"] < 0) && (df[gene,"low.7d"] > 0)) {
    df[gene, "direction"] <- "down-down-up"
  } else if((df[gene,"low.8h"] < 0) && (df[gene,"low.24h"] < 0) && (df[gene,"low.7d"] > 0)) {
    df[gene, "direction"] <- "down-down-up"
  } else if((df[gene,"low.8h"] > 0) && (df[gene,"low.24h"] < 0) && (df[gene,"low.7d"] > 0)) {
    df[gene, "direction"] <- "up-down-up"
  } else if((df[gene,"low.8h"] > 0) && (df[gene,"low.24h"] < 0) && (df[gene,"low.7d"] < 0)) {
    df[gene, "direction"] <- "up-down-down"
  } else {
    df[gene, "direction"] <- "other"
  }
}
df$direction <- factor(df$direction)

# reformat
df <- pivot_longer(data = df, cols = 2:4, names_to = "dose")
colnames(df)[6] <- "logFC"
df$dose <- factor(df$dose, levels = c("low.8h", "low.24h", "low.7d"))

# plot
fig3 <- plot_ly(df,
                x = ~dose,
                y = ~logFC,
                type = "scatter",
                mode = "lines",
                text = df$genes,
                color = ~direction,
                colors = c("blue","purple","yellow","orange","green",
                                 "gray","pink","red"))
fig3
```

## High dose
```{r high_dose_line}
# read tables
high.8h <- read.delim(paste0(out, model, "DEG_tables/psilo.high.8h_vs_sal.8h_FDRq_1.00.tsv"))
high.24h <- read.delim(paste0(out, model, "DEG_tables/psilo.high.24h_vs_sal.24h_FDRq_1.00.tsv"))
high.7d <- read.delim(paste0(out, model, "DEG_tables/psilo.high.7d_vs_sal.7d_FDRq_1.00.tsv"))

# extract logFC
df <- data.frame(genes = high.8h$gene_name_unique,
                 high.8h = high.8h$logFC,
                 high.24h = high.24h$logFC,
                 high.7d = high.7d$logFC)
rownames(df) <- df$genes

# find absolute diff
df$diff_8h_vs_24h <- abs(df$high.8h - df$high.24h)
df$diff_24h_vs_7d <- abs(df$high.24h - df$high.7d)
df <- df[df$diff_24h_vs_7d > 0.2,]
df <- df[df$diff_8h_vs_24h > 0.2,]

# direction, 1 = upreg, 2 = downreg, 3 = other
df$direction <- ""
for (gene in df$genes) {
  if((df[gene,"high.8h"] > 0) && (df[gene,"high.24h"] > 0) && (df[gene,"high.7d"] > 0)) {
    df[gene, "direction"] <- "up-up-up"
  } else if ((df[gene,"high.8h"] < 0) && (df[gene,"high.24h"] < 0) && (df[gene,"high.7d"] < 0)) {
    df[gene, "direction"] <- "down-down-down"
  } else if((df[gene,"high.8h"] < 0) && (df[gene,"high.24h"] < 0) && (df[gene,"high.7d"] > 0)) {
    df[gene, "direction"] <- "down-down-up"
  } else if((df[gene,"high.8h"] > 0) && (df[gene,"high.24h"] > 0) && (df[gene,"high.7d"] < 0)) {
    df[gene, "direction"] <- "up-up-down"
  } else if((df[gene,"high.8h"] < 0) && (df[gene,"high.24h"] > 0) && (df[gene,"high.7d"] < 0)) {
    df[gene, "direction"] <- "down-up-down"
  } else if((df[gene,"high.8h"] < 0) && (df[gene,"high.24h"] > 0) && (df[gene,"high.7d"] > 0)) {
    df[gene, "direction"] <- "down-up-up"
  } else if((df[gene,"high.8h"] < 0) && (df[gene,"high.24h"] < 0) && (df[gene,"high.7d"] > 0)) {
    df[gene, "direction"] <- "down-down-up"
  } else if((df[gene,"high.8h"] < 0) && (df[gene,"high.24h"] < 0) && (df[gene,"high.7d"] > 0)) {
    df[gene, "direction"] <- "down-down-up"
  } else if((df[gene,"high.8h"] > 0) && (df[gene,"high.24h"] < 0) && (df[gene,"high.7d"] > 0)) {
    df[gene, "direction"] <- "up-down-up"
  } else if((df[gene,"high.8h"] > 0) && (df[gene,"high.24h"] < 0) && (df[gene,"high.7d"] < 0)) {
    df[gene, "direction"] <- "up-down-down"
  } else {
    df[gene, "direction"] <- "other"
  }
}
df$direction <- factor(df$direction)

# reformat
df <- pivot_longer(data = df, cols = 2:4, names_to = "dose")
colnames(df)[6] <- "logFC"
df$dose <- factor(df$dose, levels = c("high.8h", "high.24h", "high.7d"))

# plot
fig4 <- plot_ly(df,
                x = ~dose,
                y = ~logFC,
                type = "scatter",
                mode = "lines",
                text = df$genes,
                color = ~direction,
                colors = c("blue","purple","yellow","orange","green",
                                 "gray","pink","red"))
fig4
```

