---
title: "E. coli Mice scRNAseq"
subtitle: "Clustering and Annotation"
author: "Kennedi Todd"
date: "01/28/2025"
output:
  html_document:
    code_folding: hide
    theme: cerulean
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

# Setup
## Set working directory
```{r working_directory}
knitr::opts_knit$set(root.dir = ".")
```

## Load libraries
```{r libraries, message=FALSE, warnings=FALSE}
# load libraries
library(cowplot)     # plot_grid()
library(dotenv)      # load_dot_env()
library(dplyr)       # left_join()
library(ggplot2)     # ggplot()
library(gridExtra)   # grid.arrange()
library(parallel)    # detectCores()
library(rtracklayer) # import()
library(scCustomize) # Merge_Seurat_List()
library(Seurat)      # Read10X_h5()
library(stringr)     # str_match()
```

## Variables and functions
```{r set_variables_and_thresholds}
# variables
out <- "../../results/with_integration/pass1/"

# single cell functions
files <- list.files("../../functions", full.names = TRUE)
invisible(lapply(files, source))

# load the environment variables
load_dot_env(file = "../../refs/.env")
project_dir <- Sys.getenv("PROJECT_DIR")
```

## Load data
```{r load_data}
mouse.filtered <- readRDS("../../rObjects/seurat_obj_filtered.rds")
```

# Unwanted variation
## Cell cycle
```{r cell_cycle}
markers <- paste0(project_dir, "/refs/cell_cycle_markers.tsv")
path <- paste0(project_dir, "/results/with_integration/pass1/unwanted_variation")
mouse.filtered[["phase"]] <- cell_cycle_QC(obj = mouse.filtered,
                                           markersPath = markers,
                                           species = "mouse",
                                           outDir = path)
```

## Mitochondria
[Evaluating effects of mitochondrial expression](https://hbctraining.github.io/scRNA-seq_online/lessons/06_SC_SCT_normalization.html)
```{r mitochondria_factor}
mouse.filtered[["mito_factor"]] <- 
  mitochondria_QC(obj = mouse.filtered, outDir = path)
```

# Cluster
## SCTransform
- SCTransform method is a more accurate method of normalizing, estimating the 
variance of the raw filtered data, and identifying the most variable genes. \
- Variation in sequencing depth (total nCount_RNA per cell) is normalized using 
a regularized negative binomial model. \
- If there are other sources of uninteresting variation it can be included. \
- NOTE: By default, after normalizing, adjusting the variance, and regressing out 
uninteresting sources of variation, SCTransform will rank the genes by residual 
variance and output the 3000 most variable genes. If the dataset has larger cell 
numbers, then it may be beneficial to adjust this parameter higher using the 
variable.features.n argument. Additionally, the last line of output specifies 
“Set default assay to SCT”. \
- It is suggested to not regress out batch, and instead use a data integration 
method like Harmony \
```{r split_and_sct ransform}
# split
mouse.filtered[["RNA"]] <- split(mouse.filtered[["RNA"]], 
                                 f = mouse.filtered$sample)
# transform
mouse.filtered <- SCTransform(mouse.filtered, 
                              verbose = FALSE)
```

```{r save_seurat_sct}
saveRDS(mouse.filtered, "../../rObjects/seurat_obj_sct.rds")
```

## PCA
```{r run_pca, warning=FALSE, message=FALSE}
# run PCA on the merged object
mouse.filtered <- RunPCA(object = mouse.filtered, assay = "SCT")
```

```{r plot_pca}
# Reset idents and levels
DefaultAssay(mouse.filtered) <- "SCT"

# Plot PCA
pca1 <- DimPlot(mouse.filtered,
                reduction = "pca",
                group.by = "treatment",
                shuffle = TRUE,
                raster = FALSE)
pca1

pca2 <- DimPlot(mouse.filtered,
                reduction = "pca",
                group.by = "sample",
                shuffle = TRUE,
                raster = FALSE)
pca2

pca3 <- DimPlot(mouse.filtered,
                reduction = "pca",
                group.by = "group",
                shuffle = TRUE,
                raster = FALSE)
pca3

pca4 <- DimPlot(mouse.filtered,
                reduction = "pca",
                group.by = "group2",
                shuffle = TRUE,
                raster = FALSE)
pca4

pca5 <- DimPlot(mouse.filtered,
                reduction = "pca",
                group.by = "age",
                shuffle = TRUE,
                raster = FALSE)
pca5

pca6 <- DimPlot(mouse.filtered,
                reduction = "pca",
                group.by = "timepoint_days",
                shuffle = TRUE,
                raster = FALSE)
pca6

pca7 <- DimPlot(mouse.filtered,
                reduction = "pca",
                group.by = "sex",
                shuffle = TRUE,
                raster = FALSE)
pca7

# Plot PC elbow
e1 <- ElbowPlot(mouse.filtered,
                reduction = "pca") +
  geom_vline(xintercept = 20, linetype = "dashed", color = "red")
e1
```

```{r save_pca, echo=FALSE, eval=FALSE, message=FALSE, warning=FALSE}
# save pca1
path <- paste0(out, "clustering_QC/PCA_colored_by_treatment.pdf")
pdf(path, width = 6, height = 4)
pca1
dev.off()

path <- paste0(out, "clustering_QC/PCA_colored_by_sample.pdf")
pdf(path, width = 8, height = 4)
pca2
dev.off()

path <- paste0(out, "clustering_QC/PCA_colored_by_group.pdf")
pdf(path, width = 6, height = 4)
pca3
dev.off()

path <- paste0(out, "clustering_QC/PCA_colored_by_group2.pdf")
pdf(path, width = 6, height = 4)
pca4
dev.off()

path <- paste0(out, "clustering_QC/PCA_colored_by_age.pdf")
pdf(path, width = 6, height = 4)
pca5
dev.off()

path <- paste0(out, "clustering_QC/PCA_colored_by_timepoint.pdf")
pdf(path, width = 6, height = 4)
pca6
dev.off()

path <- paste0(out, "clustering_QC/PCA_colored_by_sex.pdf")
pdf(path, width = 6, height = 4)
pca7
dev.off()

# save elbow
path <- paste0(out, "clustering_QC/PCA_elbow.pdf")
pdf(path, width = 6, height = 6)
e1
dev.off()

# cleanup
remove(pca1,pca2,pca3,pca4,pca5,pca6,pca7,e1)
```

## Integrate
```{r harmony}
# integrate
Idents(mouse.filtered) <- "sample"
mouse.integrated <- IntegrateLayers(object = mouse.filtered,
                                    method = "HarmonyIntegration",
                                    assay = "SCT")

# re-join layers
mouse.integrated[["RNA"]] <- JoinLayers(mouse.integrated[["RNA"]])

# plot harmony
h1 <- DimPlot(mouse.integrated,
              reduction = "harmony",
              shuffle = TRUE,
              raster = FALSE)
h1

h2 <- DimPlot(mouse.integrated,
              reduction = "harmony",
              shuffle = TRUE,
              raster = FALSE,
              group.by = "treatment")
h2

h3 <- DimPlot(mouse.integrated,
              reduction = "harmony",
              shuffle = TRUE,
              raster = FALSE,
              group.by = "age")
h3

h4 <- DimPlot(mouse.integrated,
              reduction = "harmony",
              shuffle = TRUE,
              raster = FALSE,
              group.by = "group")
h4

h5 <- DimPlot(mouse.integrated,
              reduction = "harmony",
              shuffle = TRUE,
              raster = FALSE,
              group.by = "group2")
h5

h6 <- DimPlot(mouse.integrated,
              reduction = "harmony",
              shuffle = TRUE,
              raster = FALSE,
              group.by = "timepoint_days")
h6

h7 <- DimPlot(mouse.integrated,
              reduction = "harmony",
              shuffle = TRUE,
              raster = FALSE,
              group.by = "sex")
h7

# Plot elbow
mouse.integrated@reductions$harmony@stdev <-
  apply(mouse.integrated@reductions$harmony@cell.embeddings, 2, sd)
e1 <- ElbowPlot(mouse.integrated,
                reduction = "harmony") +
  geom_vline(xintercept = 20, linetype = "dashed", color = "red")
e1
```

```{r save_harmony_plots, echo=FALSE, eval=FALSE}
# save
path <- paste0(out, "clustering_QC/harmony_colored_by_sample.pdf")
pdf(path, width = 8, height = 4)
h1
dev.off()

path <- paste0(out, "clustering_QC/harmony_colored_by_treatment.pdf")
pdf(path, width = 6, height = 4)
h2
dev.off()

path <- paste0(out, "clustering_QC/harmony_colored_by_age.pdf")
pdf(path, width = 6, height = 4)
h3
dev.off()

path <- paste0(out, "clustering_QC/harmony_colored_by_group.pdf")
pdf(path, width = 6, height = 4)
h4
dev.off()

path <- paste0(out, "clustering_QC/harmony_colored_by_group2.pdf")
pdf(path, width = 6, height = 4)
h5
dev.off()

path <- paste0(out, "clustering_QC/harmony_colored_by_timepoint.pdf")
pdf(path, width = 6, height = 4)
h6
dev.off()

path <- paste0(out, "clustering_QC/harmony_colored_by_sex.pdf")
pdf(path, width = 6, height = 4)
h7
dev.off()

path <- paste0(out, "clustering_QC/harmony_elbow.pdf")
pdf(path, width = 6, height = 4)
e1
dev.off()

# cleanup
remove(h1,h2,h3,h4,h5,h6,h7,e1)
```

## UMAP
To overcome the extensive technical noise in the expression of any single gene 
for scRNA-seq data, Seurat assigns cells to clusters based on their PCA scores 
derived from the expression of the integrated most variable genes, with each PC 
essentially representing a “metagene” that combines information across a 
correlated gene set. Determining how many PCs to include in the clustering step 
is therefore important to ensure that we are capturing the majority of the 
variation, or cell types, present in our dataset.
```{r run_UMAP, message=FALSE, warning=FALSE}
# run UMAP
mouse.integrated <- RunUMAP(mouse.integrated,
                            dims = 1:20,
                            reduction = "harmony",
                            n.components = 3) # set to 3 to use with VR

# plot UMAP
DimPlot(mouse.integrated,
        shuffle = TRUE)

# save
remove(mouse.filtered)
```

## Clusters
Seurat uses a graph-based clustering approach, which embeds cells in a graph 
structure, using a K-nearest neighbor (KNN) graph (by default), with edges drawn 
between cells with similar gene expression patterns. Then, it attempts to 
partition this graph into highly interconnected ‘quasi-cliques’ or ‘communities’ 
[Seurat - Guided Clustering Tutorial]. \

We will use the FindClusters() function to perform the graph-based clustering. 
The resolution is an important argument that sets the “granularity” of the 
downstream clustering and will need to be optimized for every individual experiment. 
For datasets of 3,000 - 5,000 cells, the resolution set between 0.4-1.4 generally 
yields good clustering. Increased resolution values lead to a greater number of 
clusters, which is often required for larger datasets. \

The FindClusters() function allows us to enter a series of resolutions and will 
calculate the “granularity” of the clustering. This is very helpful for testing 
which resolution works for moving forward without having to run the function for 
each resolution. \

```{r find_neighbors, message=FALSE, warning=FALSE}
# Determine the K-nearest neighbor graph
mouse.unannotated <- FindNeighbors(object = mouse.integrated,
                                   assay = "SCT",
                                   reduction = "harmony",
                                   dims = 1:20)

# Determine the clusters for various resolutions
mouse.unannotated <- FindClusters(object = mouse.unannotated,
                                  algorithm = 1,
                                  resolution = seq(0.1, 0.3, by = 0.1))
```

## Explore resolutions
```{r explore_resolutions}
# resolution of 0.1
DimPlot(mouse.unannotated,
        group.by = "SCT_snn_res.0.1",
        label = TRUE,
        shuffle = TRUE,
        raster = FALSE)

# resolution of 0.2
DimPlot(mouse.unannotated,
        group.by = "SCT_snn_res.0.2",
        label = TRUE,
        shuffle = TRUE,
        raster = FALSE)

# resolution of 0.3
DimPlot(mouse.unannotated,
        group.by = "SCT_snn_res.0.3",
        label = TRUE,
        shuffle = TRUE,
        raster = FALSE)

# reset params
mouse.unannotated$seurat_clusters <- mouse.unannotated$SCT_snn_res.0.3
Idents(mouse.unannotated) <- "seurat_clusters"
DefaultAssay(mouse.unannotated) <- "RNA"
mouse.unannotated <- NormalizeData(mouse.unannotated)

# reset levels
cluster_freqs <- table(mouse.unannotated$seurat_clusters)
sorted_clusters <- names(sort(cluster_freqs, decreasing = TRUE))
mouse.unannotated$seurat_clusters <- 
  factor(mouse.unannotated$seurat_clusters, levels = sorted_clusters)
```

```{r save_final_object, echo=FALSE, eval=FALSE}
saveRDS(mouse.unannotated, "../../rObjects/seurat_obj_unannotated.rds")
remove(mouse.integrated)
```

## Unannotated UMAP
```{r unannotated_umap}
# umap dim 1 & 2
u1 <- DimPlot(mouse.unannotated,
             group.by = "seurat_clusters",
             raster = FALSE,
             label = TRUE,
             shuffle = TRUE)
u1

# umap dim 2 & 3
u2 <- DimPlot(mouse.unannotated,
             group.by = "seurat_clusters",
             label = TRUE,
             raster = FALSE,
             dims = c(2,3),
             shuffle = TRUE)
u2

# Extract colors
plot_data <- ggplot_build(u1)$data[[1]]
cluster_order <- levels(Idents(mouse.unannotated))
color_mapping <- unique(plot_data[, c("group", "colour")])
color_mapping$group <- factor(color_mapping$group, levels = cluster_order)
color_mapping <- color_mapping[order(color_mapping$group), ]
color_df <- data.frame(Cluster = color_mapping$group, Color = color_mapping$colour)
cluster_colors <- color_df$Color
```

```{r save_unannotated_umap, echo=FALSE, eval=FALSE}
path <- paste0(out, "UMAP/unannotated_clusters_dim1&2.pdf")
pdf(path, width = 8, height = 6)
u1
dev.off()

path <- paste0(out, "UMAP/unannotated_clusters_dim2&3.pdf")
pdf(path, width = 8, height = 6)
u2
dev.off()

remove(u1,u2,color_mapping, mouse.integrated, df, data, color_df, plot_data)
```

# Annotate Clusters
## Find markers
```{r auto_find_markers}
# set params
Idents(mouse.unannotated) <- "seurat_clusters"

# find markers
markers <- SeuratWrappers::RunPrestoAll(
  object = mouse.unannotated,
  assay = "RNA",
  slot = "counts",
  only.pos = FALSE
)

# filter based on adjusted p-value
markers <- markers[markers$p_val_adj < 0.01,]

# get top 3 and top 20 per cluster
top2 <- Reduce(rbind,
               by(markers,
                  markers["cluster"],
                  head,
                  n = 2))
top20 <- Reduce(rbind,
               by(markers,
                  markers["cluster"],
                  head,
                  n = 20))

# save
write.table(markers, 
            paste0(out, "markers/auto_find_markers_adjpval_0.01.tsv"),
            sep = "\t",
            quote = FALSE,
            row.names = FALSE)
```

## Violins
```{r marker_violins}
# auto markers
Idents(mouse.unannotated) <- "seurat_clusters"
DefaultAssay(mouse.unannotated) <- "RNA"
mouse.unannotated <- NormalizeData(mouse.unannotated)

v1 <- VlnPlot(mouse.unannotated,
              features = top2$gene,
              split.by = "seurat_clusters",
              group.by = "seurat_clusters",
              cols = cluster_colors,
              stack = TRUE,
              flip = TRUE)
v1

# Microglia, Aif1 aka Iba-1
micro <- c("Aif1","C1qa","C1qb","C1qc","Cx3cr1","Hexb","Itgam","P2ry12","Ptprc",
           "Tmem119")
v2 <- VlnPlot(mouse.unannotated,
             features = micro,
             split.by = "seurat_clusters",
             group.by = "seurat_clusters",
             cols = cluster_colors,
             stack = TRUE,
             flip = TRUE)
v2

astro <- c("Aldoc","Aqp4","Gfap","Gja1","Slc1a2","Slc1a3")
v3 <- VlnPlot(mouse.unannotated,
             features = astro,
             split.by = "seurat_clusters",
             group.by = "seurat_clusters",
             cols = cluster_colors,
             stack = TRUE,
             flip = TRUE)
v3

endo <- c("Cldn5","Flt1","Itm2a","Ly6c1","Pecam1")
v4 <- VlnPlot(mouse.unannotated,
             features = endo,
             split.by = "seurat_clusters",
             group.by = "seurat_clusters",
             cols = cluster_colors,
             stack = TRUE,
             flip = TRUE)
v4

olig <- c("Cldn11","Mag","Mbp","Mog")
v5 <- VlnPlot(mouse.unannotated,
             features = olig,
             split.by = "seurat_clusters",
             group.by = "seurat_clusters",
             cols = cluster_colors,
             stack = TRUE,
             flip = TRUE)
v5

peri <- c("Acta2","Myh11","Rgs5","Pdgfrb")
v6 <- VlnPlot(mouse.unannotated,
             features = peri,
             split.by = "seurat_clusters",
             group.by = "seurat_clusters",
             cols = cluster_colors,
             stack = TRUE,
             flip = TRUE)
v6

fibro <- c("Col1a1","Col1a2","Dcn")
v7 <- VlnPlot(mouse.unannotated,
             features = fibro,
             split.by = "seurat_clusters",
             group.by = "seurat_clusters",
             cols = cluster_colors,
             stack = TRUE,
             flip = TRUE)
v7

poly <- c("Olig1","Olig2","Tnr","Sox10","Nkx2-2")
v8 <- VlnPlot(mouse.unannotated,
             features = poly,
             split.by = "seurat_clusters",
             group.by = "seurat_clusters",
             cols = cluster_colors,
             stack = TRUE,
             flip = TRUE)
v8

neuro <- c("Eno2","Gad1","Gad2","Syt1","Meg3")
v9 <- VlnPlot(mouse.unannotated,
             features = neuro,
             split.by = "seurat_clusters",
             group.by = "seurat_clusters",
             cols = cluster_colors,
             stack = TRUE,
             flip = TRUE)
v9

v10 <- VlnPlot(mouse.unannotated,
             features =  c("Ttr","Folr1","Sox4"),
             split.by = "seurat_clusters",
             group.by = "seurat_clusters",
             cols = cluster_colors,
             stack = TRUE,
             flip = TRUE)
v10

canonical <- c("Aldoc","Gja1","C1qa","Ptprc","Cldn5","Flt1","Ttr",
               "Mog","Mag","Mbp","Syt1","Gad1","Rgs5","Pdgfrb","Olig1",
               "Tnr","Col1a1","Dcn","Stmn1","Sox11")
v11 <- VlnPlot(mouse.unannotated,
             features =  canonical,
             split.by = "seurat_clusters",
             group.by = "seurat_clusters",
             cols = cluster_colors,
             stack = TRUE,
             flip = TRUE)
v11

d1 <- DotPlot(mouse.unannotated,
              features = canonical,
              cols = "RdYlBu",
              assay = "RNA") +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
d1
```

```{r save_marker_violins}
path <- paste0(out, "markers/unannotated_auto_top2_per_cluster_markers_violin.pdf")
pdf(path, width = 12, height = 10)
v1
dev.off()

path <- paste0(out, "markers/unannotated_microglia_violin.pdf")
pdf(path, width = 10, height = 6)
v2
dev.off()

path <- paste0(out, "markers/unannotated_astrocytes_markers_violin.pdf")
pdf(path, width = 10, height = 4)
v3
dev.off()

path <- paste0(out, "markers/unannotated_endothelial_markers_violin.pdf")
pdf(path, width = 10, height = 4)
v4
dev.off()

path <- paste0(out, "markers/unannotated_oligodendrocyte_markers_violin.pdf")
pdf(path, width = 10, height = 4)
v5
dev.off()

path <- paste0(out, "markers/unannotated_pericytes_smcs_markers_violin.pdf")
pdf(path, width = 10, height = 4)
v6
dev.off()

path <- paste0(out, "markers/unannotated_fibroblast_markers_violin.pdf")
pdf(path, width = 10, height = 4)
v7
dev.off()

path <- paste0(out, "markers/unannotated_polydendrocyte_markers_violin.pdf")
pdf(path, width = 10, height = 4)
v8
dev.off()

path <- paste0(out, "markers/unannotated_neuron_markers_violin.pdf")
pdf(path, width = 10, height = 4)
v9
dev.off()

path <- paste0(out, "markers/unannotated_choroid_markers_violin.pdf")
pdf(path, width = 10, height = 4)
v10
dev.off()

path <- paste0(out, "markers/unannotated_manual_cluster_markers_violin.pdf")
pdf(path, width = 10, height = 8)
v11
dev.off()

path <- paste0(out, "markers/unannotated_manual_cluster_markers_dot_plot.pdf")
pdf(path, width = 10, height = 8)
d1
dev.off()

remove(v1,v2,v3,v4,v5,v6,v7,v8,v9)
```

## Assign identities
```{r assign_identities}
new_cluster_names <- c(
  "0" = "Astrocytes",
  "1" = "Microglia",
  "2" = "Microglia",
  "3" = "Astrocytes",
  "4" = "Endothelial",
  "5" = "Astrocytes",
  "6" = "Choroid plexus",
  "7" = "Oligodendrocytes",
  "8" = "Microglia",
  "9" = "Microglia",
  "10" = "Endothelial",
  "11" = "Neurons",
  "12" = "OPCs",
  "13" = "Oligodendrocytes",
  "14" = "Neurons",
  "15" = "Pericytes and SMCs",
  "16" = "OPCs",
  "17" = "Microglia",
  "18" = "Microglia",
  "19" = "Neurogenesis",
  "20" = "Fibroblasts",
  "21" = "Ependyma",
  "22" = "Noise"
)

mouse.annotated <- RenameIdents(mouse.unannotated, new_cluster_names)
mouse.annotated$annotated_clusters <- Idents(mouse.annotated)
Idents(mouse.annotated) <- "annotated_clusters"
```

## Annotated UMAP
```{r annotated_umap}
cluster_colors <- c("firebrick2", "gold", "chartreuse2", "darkgreen", "cyan",
                    "steelblue","blue", "violet", "purple3","tan","gray","black")
# umap dim 1 & 2
u1 <- DimPlot(mouse.annotated,
              cols = cluster_colors,
              shuffle = TRUE)
u1

# umap dim 2 & 3
u2 <- DimPlot(mouse.annotated,
              dims = c(2,3),
              shuffle = TRUE,
              cols = cluster_colors)
u2
```

```{r save_annotated_umap, echo=FALSE, eval=FALSE}
path <- paste0(out, "UMAP/annotated_clusters_dim1&2.pdf")
pdf(path, width = 8, height = 6)
u1
dev.off()

path <- paste0(out, "UMAP/annotated_clusters_dim2&3.pdf")
pdf(path, width = 8, height = 6)
u2
dev.off()

remove(u1,u2)
```

```{r save_annotated_obj, echo=FALSE, eval=FALSE}
# save seurat object
saveRDS(mouse.annotated, "../../rObjects/seurat_obj_annotated.rds")
```

## Annotated violins
```{r annotated_violin}
v1 <- VlnPlot(mouse.annotated,
             features = canonical,
             cols = cluster_colors,
             group.by = "annotated_clusters",
             split.by = "annotated_clusters",
             stack = TRUE,
             flip = TRUE)
v1

d1 <- DotPlot(mouse.annotated,
              features = canonical,
              cols = "RdYlBu",
              assay = "RNA") +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
d1
```

```{r}
path <- paste0(out, "markers/annotated_manual_cluster_markers_violin.pdf")
pdf(path, width = 8, height = 8)
v1
dev.off()

path <- paste0(out, "markers/annotated_manual_cluster_markers_dot_plot.pdf")
pdf(path, width = 8, height = 8)
d1
dev.off()
```

# Shiny App
- scDRcell is the function used to plot the first UMAP
- inpConf is an argument of scDRcell, inpConf = sc1conf
- infConf gets subset by UI column based on inpdrX which is the user's input for x-axis
- inpdrX is used to set the x-axis label
```{r shiny_app, eval=FALSE}
# create new object
library(ShinyCell)
shiny.obj <- mouse.annotated
VariableFeatures(shiny.obj) <- shiny.obj@assays$SCT@var.features

# set default params
DefaultAssay(shiny.obj) <- "RNA"

# create config
names <- colnames(shiny.obj@meta.data)
names <- names[c(32,33,2:31,34)]
sc.config <- createConfig(obj = shiny.obj,
                          meta.to.include = names)

# change wd
setwd(out)

# output shiny app folder
makeShinyApp(obj = shiny.obj, 
             scConf = sc.config, 
             gene.mapping = TRUE,
             shiny.title = "E. coli Mice")

# manual config edits
setwd("../../results/without_integration")
sc1conf <- readRDS("shinyApp/sc1conf.rds")
cluster_hex <- "firebrick2|gold|chartreuse2|darkgreen|cyan|steelblue|blue|violet|purple3|gray|chocolate4|black"
sc1conf[2,4] <- cluster_hex
saveRDS(sc1conf, "shinyApp/sc1conf.rds")
```
