---
title: "E. coli Mice scRNAseq"
subtitle: "Differential Expression"
author: "Kennedi Todd"
date: "08/21/2025"
output:
  html_document:
    code_folding: hide
    theme: cerulean
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

# Setup
## Set working directory
```{r working_directory}
knitr::opts_knit$set(root.dir = ".")
```

## Load libraries
```{r libraries, message=FALSE, warnings=FALSE}
library(DESeq2)
library(dplyr)      # %>%
library(ggplot2)    # coord_flip()
library(Seurat)     # DimPlot()
library(stringr)    # str_match()
library(ggrepel)    # geom_text_repel()
library(gtools)     # smartbind()
library(tidyr)
```

## Variables and functions
```{r set_variables_and_thresholds}
# source thresholds and output paths
source("../../refs/thresholds_and_outs.R")
out <- paste0(out, "pass3/")

# single cell functions
files <- list.files("../../functions", full.names = TRUE)
invisible(lapply(files, source))
```

## Load data
```{r load_data}
mouse.annotated <- readRDS("../../rObjects/pass3_annotated_seurat_obj.rds")
```

## Check
```{r annotated_umap}
cluster_colors <- c(
  "Astrocytes"  = "#F0E442",
  "Choroid"     = "#FDB863",
  "Endothelial" = "#E08214",
  "Fibroblasts" = "#B35806",
  "Microglia" = "#542788",
  "BAM" = "#998EC3",
  "Neurons 1" = "#0868AC",
  "Neurons 2" = "#2B8CBE",
  "Oligodendrocytes" = "#009E73",
  "OPCs" = "#7FCDBB"
)

DimPlot(mouse.annotated,
        cols = cluster_colors,
        shuffle = TRUE)
```

# QC
## Sex check
```{r sex_check}
dot <- DotPlot(mouse.annotated,
               features = c("Uty","Xist"),
               group.by = "sample")
dot
```

```{r save_sex_check, echo=FALSE, eval=FALSE}
pdf(paste0(out, "clustering_QC/sex_check_dot_plot.pdf"), width = 6, height = 16)
dot
dev.off()
```


## UMAP colored by meta
```{r split_umaps}
colors <- c("cyan2","coral","chartreuse2","blue","red","gold","forestgreen",
            "purple2")

# group
umap1 <- DimPlot(object = mouse.annotated,
                 reduction = "umap",
                 group.by = "group",
                 raster = FALSE,
                 shuffle = TRUE,
                 cols = colors)
umap1

# group2
umap2 <- DimPlot(object = mouse.annotated,
                 reduction = "umap",
                 group.by = "group2",
                 raster = FALSE,
                 shuffle = TRUE)
umap2

# sex
umap3 <- DimPlot(object = mouse.annotated,
                 reduction = "umap",
                 group.by = "sex",
                 raster = FALSE,
                 shuffle = TRUE,
                 cols = colors)
umap3

# treatment
umap4 <- DimPlot(object = mouse.annotated, 
                 reduction = "umap",
                 group.by = "treatment",
                 raster = FALSE,
                 shuffle = TRUE,
                 cols = colors)
umap4

# mito_factor
umap5 <- DimPlot(object = mouse.annotated, 
                 group.by = "mito_factor",
                 reduction = "umap",
                 raster = FALSE,
                 shuffle = TRUE,
                 cols = colors)
umap5

# phase
umap6 <- DimPlot(object = mouse.annotated, 
                 reduction = "umap",
                 group.by = "phase",
                 raster = FALSE,
                 shuffle = TRUE,
                 cols = colors)
umap6

# timepoint
umap7 <- DimPlot(object = mouse.annotated,
                 reduction = "umap",
                 group.by = "timepoint_days",
                 raster = FALSE,
                 shuffle = TRUE,
                 cols = colors)
umap7

# age
umap8 <- DimPlot(object = mouse.annotated,
                 reduction = "umap",
                 group.by = "age",
                 raster = FALSE,
                 shuffle = TRUE,
                 cols = colors)
umap8
```

```{r save_split_umap, echo=FALSE, eval=FALSE}
path <- paste0(out, "clustering_QC/umap_colored_by_")
pdf(paste0(path, "group.pdf"), width = 6, height = 4)
umap1
dev.off()

pdf(paste0(path, "group2.pdf"), width = 6, height = 4)
umap2
dev.off()

pdf(paste0(path, "sex.pdf"), width = 6, height = 4)
umap3
dev.off()

pdf(paste0(path, "treatment.pdf"), width = 6, height = 4)
umap4
dev.off()

pdf(paste0(path, "mito_factor.pdf"), width = 6, height = 4)
umap5
dev.off()

pdf(paste0(path, "cell_cycle_phase.pdf"), width = 6, height = 4)
umap6
dev.off()

pdf(paste0(path, "timepoint.pdf"), width = 6, height = 4)
umap7
dev.off()

pdf(paste0(path, "age.pdf"), width = 6, height = 4)
umap8
dev.off()

remove(umap1,umap2,umap3,umap4,umap5,umap6,umap7,umap8)
```


## UMAP split by meta
```{r split_umaps}
# group
umap1 <- DimPlot(object = mouse.annotated,
                 reduction = "umap",
                 group.by = "annotated_clusters",
                 split.by = "group",
                 raster = FALSE,
                 ncol = 4,
                 shuffle = TRUE,
                 cols = cluster_colors)
umap1
```

## Heatmap UMAP
```{r feature_plots_annotated, warning=FALSE, message=FALSE}
# UMAP percent_mt
f1 <- FeaturePlot(mouse.annotated, 
            reduction = "umap", 
            features = "percent_mt")  + 
  scale_colour_gradientn(colours = c("blue","lightblue","yellow","orange","red"))
f1

# UMAP nCount
f2 <- FeaturePlot(mouse.annotated, 
            reduction = "umap",
            features = "nCount_RNA") + 
  scale_colour_gradientn(colours = c("blue","lightblue","yellow","orange","red"))
f2

# UMAP nFeature
f3 <- FeaturePlot(mouse.annotated, 
            reduction = "umap", 
            features = "nFeature_RNA") + 
  scale_colour_gradientn(colours = c("blue","lightblue","yellow","orange","red"))
f3

# UMAP percent_ribo
f4 <- FeaturePlot(mouse.annotated, 
            reduction = "umap", 
            features = "percent_ribo") + 
  scale_colour_gradientn(colours = c("blue","lightblue","yellow","orange","red"))
f4

# UMAP cell_complexity
f5 <- FeaturePlot(mouse.annotated, 
            reduction = "umap", 
            features = "cell_complexity") + 
  scale_colour_gradientn(colours = c("blue","lightblue","yellow","orange","red"))
f5

# UMAP percent_hb
f6 <- FeaturePlot(mouse.annotated, 
            reduction = "umap", 
            features = "percent_hb") + 
  scale_colour_gradientn(colours = c("blue","lightblue","yellow","orange","red"))
f6
```

```{r save_umap_heatmap, echo=FALSE, eval=FALSE}
path <- paste0(out, "clustering_QC/umap_heatmap_")
pdf(paste0(path, "percent_mito.pdf"), width = 8, height = 6)
f1
dev.off()

pdf(paste0(path, "nCount.pdf"), width = 8, height = 6)
f2
dev.off()

pdf(paste0(path, "nFeature.pdf"), width = 8, height = 6)
f3
dev.off()

pdf(paste0(path, "percent_ribo.pdf"), width = 8, height = 6)
f4
dev.off()

pdf(paste0(path, "cell_complexity.pdf"), width = 8, height = 6)
f5
dev.off()

pdf(paste0(path, "percent_hb.pdf"), width = 8, height = 6)
f6
dev.off()

remove(f1,f2,f3,f4,f5,f6)
```

## Percent cells per cluster
```{r percent_cells}
# timepoint
b1 <- mouse.annotated@meta.data %>%
  group_by(annotated_clusters, timepoint_days) %>%
  dplyr::count() %>%
  group_by(annotated_clusters) %>%
  dplyr::mutate(percent = 100*n/sum(n)) %>%
  ungroup() %>%
  ggplot(aes(x=annotated_clusters,y=percent, fill=timepoint_days)) +
  theme_classic() +
  geom_col() +
  ggtitle("Percentage of time point per cluster") +
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))
b1

# sex
b2 <- mouse.annotated@meta.data %>%
  group_by(annotated_clusters, sex) %>%
  dplyr::count() %>%
  group_by(annotated_clusters) %>%
  dplyr::mutate(percent = 100*n/sum(n)) %>%
  ungroup() %>%
  ggplot(aes(x=annotated_clusters,y=percent, fill=sex)) +
  theme_classic() +
  geom_col() +
  ggtitle("Percentage of sex per cluster") +
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))
b2

# sample
b3 <- mouse.annotated@meta.data %>%
  group_by(annotated_clusters, sample) %>%
  dplyr::count() %>%
  group_by(annotated_clusters) %>%
  dplyr::mutate(percent = 100*n/sum(n)) %>%
  ungroup() %>%
  ggplot(aes(x=annotated_clusters,y=percent, fill=sample)) +
  theme_classic() +
  geom_col() +
  ggtitle("Percentage of sample per cluster") +
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))
b3

# phase
b4 <- mouse.annotated@meta.data %>%
  group_by(annotated_clusters, phase) %>%
  dplyr::count() %>%
  group_by(annotated_clusters) %>%
  dplyr::mutate(percent = 100*n/sum(n)) %>%
  ungroup() %>%
  ggplot(aes(x=annotated_clusters,y=percent, fill=phase)) +
  theme_classic() +
  geom_col() +
  ggtitle("Percentage of phase per cluster") +
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))
b4

# mito.factor
b5 <- mouse.annotated@meta.data %>%
  group_by(annotated_clusters, mito_factor) %>%
  dplyr::count() %>%
  group_by(annotated_clusters) %>%
  dplyr::mutate(percent = 100*n/sum(n)) %>%
  ungroup() %>%
  ggplot(aes(x=annotated_clusters,y=percent, fill=mito_factor)) +
  theme_classic() +
  geom_col() +
  ggtitle("Percentage of mito factor per cluster") +
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))
b5

# group
b6 <- mouse.annotated@meta.data %>%
  group_by(annotated_clusters, group) %>%
  dplyr::count() %>%
  group_by(annotated_clusters) %>%
  dplyr::mutate(percent = 100*n/sum(n)) %>%
  ungroup() %>%
  ggplot(aes(x=annotated_clusters,y=percent, fill=group)) +
  theme_classic() +
  geom_col() +
  ggtitle("Percentage of group per cluster") +
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))
b6

# group2
b7 <- mouse.annotated@meta.data %>%
  group_by(annotated_clusters, group2) %>%
  dplyr::count() %>%
  group_by(annotated_clusters) %>%
  dplyr::mutate(percent = 100*n/sum(n)) %>%
  ungroup() %>%
  ggplot(aes(x=annotated_clusters,y=percent, fill=group2)) +
  theme_classic() +
  geom_col() +
  ggtitle("Percentage of group2 per cluster") +
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))
b7

# age
b8 <- mouse.annotated@meta.data %>%
  group_by(annotated_clusters, age) %>%
  dplyr::count() %>%
  group_by(annotated_clusters) %>%
  dplyr::mutate(percent = 100*n/sum(n)) %>%
  ungroup() %>%
  ggplot(aes(x=annotated_clusters,y=percent, fill=age)) +
  theme_classic() +
  geom_col() +
  ggtitle("Percentage of age per cluster") +
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))
b8

# group / cluster
mouse.annotated$group <- factor(mouse.annotated$group,
                                levels = c("S.2D.Y", "E.2D.Y",
                                           "S.2D.O", "E.2D.O",
                                           "S.21D.Y", "E.21D.Y",
                                           "S.21D.O", "E.21D.O"))
b9 <- mouse.annotated@meta.data %>%
  group_by(group, annotated_clusters) %>%
  dplyr::count() %>%
  group_by(group) %>%
  dplyr::mutate(percent = 100*n/sum(n)) %>%
  ungroup() %>%
  ggplot(aes(x=group,y=percent, fill=annotated_clusters)) +
  scale_fill_manual(values = cluster_colors) +
  theme_classic() +
  geom_col() +
  ggtitle("Percentage of cell type per group") +
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5, hjust=1))
b9
```

```{r save_percent_cells,echo=FALSE,eval=FALSE}
# save
path <- paste0(out, "clustering_QC/percent_")
pdf(paste0(path, "timepoint_per_cluster.pdf"), width = 8, height = 6)
b1
dev.off()

# save
pdf(paste0(path, "sex_per_cluster.pdf"), width = 8, height = 6)
b2
dev.off()

# save
pdf(paste0(path, "sample_per_cluster.pdf"), width = 8, height = 8)
b3
dev.off()

# save
pdf(paste0(path, "phase_per_cluster.pdf"), width = 8, height = 6)
b4
dev.off()

# save
pdf(paste0(path, "mito_factor_per_cluster.pdf"), width = 8, height = 6)
b5
dev.off()

# save
pdf(paste0(path, "group_per_cluster.pdf"), width = 8, height = 6)
b6
dev.off()

# save
pdf(paste0(path, "group2_per_cluster.pdf"), width = 8, height = 6)
b7
dev.off()

# save
pdf(paste0(path, "age_per_cluster.pdf"), width = 8, height = 6)
b8
dev.off()

# save
pdf(paste0(path, "cell_types_per_group.pdf"), width = 8, height = 6)
b9
dev.off()

# cleanup
remove(b1,b2,b3,b4,b5,b6,b7,b8)
```

## Cells per sample
```{r cells_per_sample}
data <- as.data.frame(table(mouse.annotated$sample))
colnames(data) <- c("sample","frequency")

ncells <- ggplot(data, aes(x = sample, y = frequency, fill = sample)) + 
  geom_col() +
  theme_classic() +
  geom_text(aes(label = frequency), 
            position=position_dodge(width=0.9), 
            vjust=-0.25) +
  scale_y_continuous(breaks = seq(0,8000, by = 2000), limits = c(0,8000)) +
  ggtitle("Filtered: cells per sample") +
  theme(legend.position =  "none") + 
  theme(axis.text.x = element_text(angle = 45, hjust=1))
ncells
```

```{r save_cells_per_sample, echo=FALSE, eval=FALSE}
pdf(paste0(out, "clustering_QC/cells_per_sample.pdf"), width = 18, height = 4)
ncells
dev.off()
```

## Cells per cluster
```{r cells_per_cluster}
data <- as.data.frame(table(mouse.annotated$annotated_clusters))
colnames(data) <- c("cluster","frequency")

ncells <- ggplot(data, aes(x = cluster, y = frequency, fill = cluster)) + 
  geom_col() +
  theme_classic() +
  geom_text(aes(label = frequency), 
            position=position_dodge(width=0.9), 
            vjust=-0.25) +
  scale_y_continuous(breaks = seq(0,40000, by = 10000), limits = c(0,40000)) +
  ggtitle("Cells per Cluster") +
  scale_fill_manual(values = cluster_colors) +
  theme(legend.position =  "none") + 
  theme(axis.text.x = element_text(angle = 45, hjust=1))
ncells
```

```{r save_cells_per_cluster, echo=FALSE, eval=FALSE}
pdf(paste0(out, "clustering_QC/cells_per_cluster.pdf"), width = 8, height = 6)
ncells
dev.off()
```

## Cells per cluster per sample
```{r cells_per_cluster_per_sample}
# extract data
data <- data.frame(cluster = mouse.annotated$annotated_clusters,
                   sample = mouse.annotated$sample)
data_summary <- data %>%
  group_by(cluster, sample) %>%
  summarise(count = n(), .groups = 'drop') %>%
  ungroup()

# Create the stacked bar plot
ncells <- ggplot(data_summary, aes(x = cluster, y = count, fill = sample)) +
  geom_bar(stat = "identity") +
  labs(y = "Frequency", x = "Cluster", fill = "Sample") +
  theme_minimal() +
  ggtitle("Frequency of Cells in Each Cluster with Sample Distribution") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  ylim(c(0,40000))
ncells

# reformat for table output
data_summary <- data_summary %>% pivot_wider(names_from = cluster, values_from = count)
```

```{r save_cells_per_cluster_per_sample, echo=FALSE, eval=FALSE}
pdf(paste0(out, "clustering_QC/cells_per_cluster_with_sample_distribution.pdf"),
    width = 12, height = 6)
ncells
dev.off()
```

## Cells per cluster per group
```{r cells_per_cluster_per_group}
# extract data
data <- data.frame(cluster = mouse.annotated$annotated_clusters,
                   group = mouse.annotated$group)
data_summary <- data %>%
  group_by(cluster, group) %>%
  summarise(count = n(), .groups = 'drop') %>%
  ungroup()

# Create the stacked bar plot
ncells <- ggplot(data_summary, aes(x = cluster, y = count, fill = group)) +
  geom_bar(stat = "identity") +
  labs(y = "Frequency", x = "Cluster", fill = "Group") +
  theme_minimal() +
  ggtitle("Frequency of Cells in Each Cluster with Group Distribution") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  ylim(c(0,40000))
ncells

# reformat for table output
data_summary <- data_summary %>% pivot_wider(names_from = cluster, values_from = count)
```

```{r save_cells_per_cluster_per_group, echo=FALSE, eval=FALSE}
pdf(paste0(out, "clustering_QC/cells_per_cluster_with_group_distribution.pdf"),
    width = 8, height = 6)
ncells
dev.off()

write.table(data_summary, 
            file = paste0(out, "clustering_QC/cells_per_cluster_with_group_distribution.tsv"), 
            sep = "\t",quote = FALSE, row.names = FALSE)
```

# DEG Plots
## Heatmap Sex Specific
### Female vs Male
```{r compare_degs}
# set thresh
adjpval_thresh <- 0.05
lfc_thresh <- 0

# get file paths
files <- list.files(paste0(out, "DEGs_MAST_pct0.4/DEG_tables"))
files <- files[grepl(".+F_vs_.+M_DEGs.tsv", files)]

# get comparisons
comparisons <- gsub("_DEGs.tsv", "", files)
all_comparisons <- c(paste0(comparisons, "_up"),
                     paste0(comparisons, "_down"))

# get cell types
cell_types <- names(cluster_colors)

# init deg.df
deg.df <- data.frame(
  matrix(0, nrow = length(all_comparisons), ncol = length(cell_types)),
  row.names = all_comparisons
)
colnames(deg.df) <- cell_types

# reorder rows
deg.df <- deg.df[c(16,8,12,4,15,7,11,3,14,6,10,2,13,5,9,1),]

# loop through files
for (i in files) {
  
  # read table
  df <- read.table(paste0(out, "DEGs_MAST_pct0.4/DEG_tables/", i),
                   sep = "\t", 
                   header = TRUE)
  
  # filter
  df <- df[!is.na(df$padj),]
  df <- df[df$padj < adjpval_thresh,]
  df <- df[df$log2FoldChange < -lfc_thresh | df$log2FoldChange > lfc_thresh,]
  
  # add direction column
  comparisonName <- gsub("_DEGs.tsv", "", i)
  direction <- df$log2FoldChange > 0 # already filtered on lfc so only matters it's positive/negative
  direction <- gsub(TRUE, paste0(comparisonName, "_up"), direction)
  direction <- gsub(FALSE, paste0(comparisonName, "_down"), direction)
  df$direction <- direction
  
  # reformat table
  df <- df %>% 
    dplyr::count(cell_type, direction) %>%
    tidyr::spread(cell_type, n)
  rownames(df) <- df$direction
  df$direction <- NULL
  
  # add to master table
  deg.df[rownames(df), colnames(df)] <- df
}

# reformat
deg.df[is.na(deg.df)] <- 0

# save
prefix <- paste0(out, "DEGs_MAST_pct0.4/DEG_comparison/",
                 "female_vs_male_DEG_sex_specific_comparison_FDRq_", 
                 format(adjpval_thresh, nsmall = 2),
                 "_LFC_", format(lfc_thresh, nsmall = 2))
write.table(df, 
            paste0(prefix, ".tsv"),
            sep = "\t", quote = FALSE, row.names = TRUE)

# set heatmap colors and names
paletteLength <- 100
myColor <- colorRampPalette(c("white","#f0eb9e","darkgreen"))(paletteLength)

meta <- data.frame(treatment = rep(c("Saline", "Saline", "E. coli", "E. coli"), 4),
                   age = c(rep("Young", 4), rep("Old", 4),
                           rep("Young", 4), rep("Old", 4)),
                   time_point = c(rep("2 days", 8), rep("21 days", 8)),
                   direction = rep(c("Down-regulated","Up-regulated"), 8))
rownames(meta) <- rownames(deg.df)

ann_colors <- list(treatment = c(`E. coli` = "pink2", Saline = "chocolate4"),
                   age = c(Young = "gold",
                           Old = "chartreuse2"),
                   time_point = c(`2 days` = "cyan", 
                                  `21 days` = "purple"),
                   direction = c(`Down-regulated` = "#0271f0",
                                 `Up-regulated` = "tomato"))

# save
pdf(paste0(prefix, ".pdf"), width = 10, height = 10)

# plot
pheatmap::pheatmap(deg.df,
                   main = paste0("Female vs Male, FDRq < ", format(adjpval_thresh, nsmall = 2), 
                                 ", |LFC| > ", format(lfc_thresh, nsmall = 2)),
                   treeheight_row = 0,
                   treeheight_col = 0,
                   color = myColor,
                   cluster_rows = FALSE,
                   annotation_row = meta,
                   annotation_colors = ann_colors,
                   display_numbers = round(deg.df, digits = 0),
                   fontsize_number = 12,
                   number_color = "black")
```

### Old vs Young
```{r old_vs_young_degs, eval=FALSE}
# set thresh
adjpval_thresh <- 0.05
lfc_thresh <- 0

# get file paths
files <- list.files(paste0(out, "DEGs_MAST_pct0.4/DEG_tables"))
files <- files[grepl(".+O.+_vs_.+Y.+_DEGs.tsv", files)]

# get comparisons
comparisons <- gsub("_DEGs.tsv", "", files)
all_comparisons <- c(paste0(comparisons, "_up"),
                     paste0(comparisons, "_down"))

# get cell types
cell_types <- names(cluster_colors)

# init deg.df
deg.df <- data.frame(
  matrix(0, nrow = length(all_comparisons), ncol = length(cell_types)),
  row.names = all_comparisons
)
colnames(deg.df) <- cell_types

# reorder rows
deg.df <- deg.df[c(15,7,11,3,16,8,12,4,13,5,9,1,14,6,10,2),]

# loop through files
for (i in files) {
  
  # read table
  df <- read.table(paste0(out, "DEGs_MAST_pct0.4/DEG_tables/", i),
                   sep = "\t", 
                   header = TRUE)
  
  # filter
  df <- df[!is.na(df$padj),]
  df <- df[df$padj < adjpval_thresh,]
  df <- df[df$log2FoldChange < -lfc_thresh | df$log2FoldChange > lfc_thresh,]
  
  # add direction column
  comparisonName <- gsub("_DEGs.tsv", "", i)
  direction <- df$log2FoldChange > 0 # already filtered on lfc so only matters it's positive/negative
  direction <- gsub(TRUE, paste0(comparisonName, "_up"), direction)
  direction <- gsub(FALSE, paste0(comparisonName, "_down"), direction)
  df$direction <- direction
  
  # reformat table
  df <- df %>% 
    dplyr::count(cell_type, direction) %>%
    tidyr::spread(cell_type, n)
  rownames(df) <- df$direction
  df$direction <- NULL
  
  # add to master table
  deg.df[rownames(df), colnames(df)] <- df
}

# reformat
deg.df[is.na(deg.df)] <- 0

# save
prefix <- paste0(out, "DEGs_MAST_pct0.4/DEG_comparison/",
                 "old_vs_young_DEG_sex_specific_comparison_FDRq_", 
                 format(adjpval_thresh, nsmall = 2),
                 "_LFC_", format(lfc_thresh, nsmall = 2))
write.table(df, 
            paste0(prefix, ".tsv"),
            sep = "\t", quote = FALSE, row.names = TRUE)

# set heatmap colors and names
paletteLength <- 100
myColor <- colorRampPalette(c("white","#f0eb9e","darkgreen"))(paletteLength)

meta <- data.frame(treatment = rep(c("Saline", "Saline", "E. coli", "E. coli"), 4),
                   sex = c(rep("Female", 4), rep("Male", 4),
                           rep("Female", 4), rep("Male", 4)),
                   time_point = c(rep("2 days", 8), rep("21 days", 8)),
                   direction = rep(c("Down-regulated","Up-regulated"), 8))
rownames(meta) <- rownames(deg.df)

ann_colors <- list(treatment = c(`E. coli` = "pink2", Saline = "chocolate4"),
                   sex = c(Female = "gray",
                           Male = "orange"),
                   time_point = c(`2 days` = "cyan", 
                                  `21 days` = "purple"),
                   direction = c(`Down-regulated` = "#0271f0",
                                 `Up-regulated` = "tomato"))

# save
pdf(paste0(prefix, ".pdf"), width = 10, height = 10)

# plot
pheatmap::pheatmap(deg.df,
                   main = paste0("Old vs Young, FDRq < ", format(adjpval_thresh, nsmall = 2), 
                                 ", |LFC| > ", format(lfc_thresh, nsmall = 2)),
                   treeheight_row = 0,
                   treeheight_col = 0,
                   color = myColor,
                   cluster_rows = FALSE,
                   annotation_row = meta,
                   annotation_colors = ann_colors,
                   display_numbers = round(deg.df, digits = 0),
                   fontsize_number = 12,
                   number_color = "black")
```

### 21 vs 2 days
```{r old_vs_young_degs, eval=FALSE}
# set thresh
adjpval_thresh <- 0.05
lfc_thresh <- 0

# get file paths
files <- list.files(paste0(out, "DEGs_MAST_pct0.4/DEG_tables"))
files <- files[grepl(".+21D.+[MF]_vs_.+2D.+[MF]_DEGs.tsv", files)]

# get comparisons
comparisons <- gsub("_DEGs.tsv", "", files)
all_comparisons <- c(paste0(comparisons, "_up"),
                     paste0(comparisons, "_down"))

# get cell types
cell_types <- names(cluster_colors)

# init deg.df
deg.df <- data.frame(
  matrix(0, nrow = length(all_comparisons), ncol = length(cell_types)),
  row.names = all_comparisons
)
colnames(deg.df) <- cell_types

# reorder rows
deg.df <- deg.df[c(15,7,11,3,16,8,12,4,13,5,9,1,14,6,10,2),]

# loop through files
for (i in files) {
  
  # read table
  df <- read.table(paste0(out, "DEGs_MAST_pct0.4/DEG_tables/", i),
                   sep = "\t", 
                   header = TRUE)
  
  # filter
  df <- df[!is.na(df$padj),]
  df <- df[df$padj < adjpval_thresh,]
  df <- df[df$log2FoldChange < -lfc_thresh | df$log2FoldChange > lfc_thresh,]
  
  # add direction column
  comparisonName <- gsub("_DEGs.tsv", "", i)
  direction <- df$log2FoldChange > 0 # already filtered on lfc so only matters it's positive/negative
  direction <- gsub(TRUE, paste0(comparisonName, "_up"), direction)
  direction <- gsub(FALSE, paste0(comparisonName, "_down"), direction)
  df$direction <- direction
  
  # reformat table
  df <- df %>% 
    dplyr::count(cell_type, direction) %>%
    tidyr::spread(cell_type, n)
  rownames(df) <- df$direction
  df$direction <- NULL
  
  # add to master table
  deg.df[rownames(df), colnames(df)] <- df
}

# reformat
deg.df[is.na(deg.df)] <- 0

# save
prefix <- paste0(out, "DEGs_MAST_pct0.4/DEG_comparison/",
                 "21_vs_2_days_DEG_sex_specific_comparison_FDRq_", 
                 format(adjpval_thresh, nsmall = 2),
                 "_LFC_", format(lfc_thresh, nsmall = 2))
write.table(df, 
            paste0(prefix, ".tsv"),
            sep = "\t", quote = FALSE, row.names = TRUE)

# set heatmap colors and names
paletteLength <- 100
myColor <- colorRampPalette(c("white","#f0eb9e","darkgreen"))(paletteLength)

meta <- data.frame(treatment = rep(c("Saline", "Saline", "E. coli", "E. coli"), 4),
                   sex = c(rep("Female", 4), rep("Male", 4),
                           rep("Female", 4), rep("Male", 4)),
                   age = c(rep("Young", 8), rep("Old", 8)),
                   direction = rep(c("Down-regulated","Up-regulated"), 8))
rownames(meta) <- rownames(deg.df)

ann_colors <- list(treatment = c(`E. coli` = "pink2", Saline = "chocolate4"),
                   sex = c(Female = "gray",
                           Male = "orange"),
                   age = c(Young = "gold",
                           Old = "chartreuse2"),
                   direction = c(`Down-regulated` = "#0271f0",
                                 `Up-regulated` = "tomato"))

# save
pdf(paste0(prefix, ".pdf"), width = 10, height = 10)

# plot
pheatmap::pheatmap(deg.df,
                   main = paste0("21 vs 2 Days, FDRq < ", format(adjpval_thresh, nsmall = 2), 
                                 ", |LFC| > ", format(lfc_thresh, nsmall = 2)),
                   treeheight_row = 0,
                   treeheight_col = 0,
                   color = myColor,
                   cluster_rows = FALSE,
                   annotation_row = meta,
                   annotation_colors = ann_colors,
                   display_numbers = round(deg.df, digits = 0),
                   fontsize_number = 12,
                   number_color = "black")
```


### Ecoli vs saline
```{r old_vs_young_degs, eval=FALSE}
# set thresh
adjpval_thresh <- 0.01
lfc_thresh <- 0.2

# get file paths
files <- list.files(paste0(out, "DEGs_MAST_pct0.4/DEG_tables"))
files <- files[grepl("E.+[MF]_vs_S.+[MF]_DEGs.tsv", files)]

# get comparisons
comparisons <- gsub("_DEGs.tsv", "", files)
all_comparisons <- c(paste0(comparisons, "_up"),
                     paste0(comparisons, "_down"))

# get cell types
cell_types <- names(cluster_colors)

# init deg.df
deg.df <- data.frame(
  matrix(0, nrow = length(all_comparisons), ncol = length(cell_types)),
  row.names = all_comparisons
)
colnames(deg.df) <- cell_types

# reorder rows
deg.df <- deg.df[c(15,7,16,8,13,5,14,6,11,3,12,4,9,1,10,2),]

# loop through files
for (i in files) {
  
  # read table
  df <- read.table(paste0(out, "DEGs_MAST_pct0.4/DEG_tables/", i),
                   sep = "\t", 
                   header = TRUE)
  
  # filter
  df <- df[!is.na(df$p_val_adj),]
  df <- df[df$p_val_adj < adjpval_thresh,]
  df <- df[df$avg_log2FC < -lfc_thresh | df$avg_log2FC > lfc_thresh,]
  
  # add direction column
  comparisonName <- gsub("_DEGs.tsv", "", i)
  direction <- df$avg_log2FC > 0 # already filtered on lfc so only matters it's positive/negative
  direction <- gsub(TRUE, paste0(comparisonName, "_up"), direction)
  direction <- gsub(FALSE, paste0(comparisonName, "_down"), direction)
  df$direction <- direction
  
  # reformat table
  df <- df %>% 
    dplyr::count(cluster, direction) %>%
    tidyr::spread(cluster, n)
  rownames(df) <- df$direction
  df$direction <- NULL
  
  # add to master table
  deg.df[rownames(df), colnames(df)] <- df
}

# reformat
deg.df[is.na(deg.df)] <- 0

# save
prefix <- paste0(out, "DEGs_MAST_pct0.4/DEG_comparison/",
                 "ecoli_vs_saline_DEG_sex_specific_comparison_FDRq_", 
                 format(adjpval_thresh, nsmall = 2),
                 "_LFC_", format(lfc_thresh, nsmall = 2))
write.table(df, 
            paste0(prefix, ".tsv"),
            sep = "\t", quote = FALSE, row.names = TRUE)

# set heatmap colors and names
paletteLength <- 100
myColor <- colorRampPalette(c("white","#f0eb9e","darkgreen"))(paletteLength)

meta <- data.frame(time_point =  c(rep("2 days", 8), rep("21 days", 8)),
                   sex = rep(c("Female","Female","Male","Male"), 4),
                   age = c(rep("Young", 4), rep("Old", 4),
                           rep("Young", 4), rep("Old", 4)),
                   direction = rep(c("Down-regulated","Up-regulated"), 8))
rownames(meta) <- rownames(deg.df)

ann_colors <- list(time_point = c(`2 days` = "cyan", 
                                  `21 days` = "purple"),
                   sex = c(Female = "gray",
                           Male = "orange"),
                   age = c(Young = "gold",
                           Old = "chartreuse2"),
                   direction = c(`Down-regulated` = "#0271f0",
                                 `Up-regulated` = "tomato"))

# save
pdf(paste0(prefix, ".pdf"), width = 10, height = 10)

# plot
pheatmap::pheatmap(deg.df,
                   main = paste0("E. coli vs Saline, FDRq < ", format(adjpval_thresh, nsmall = 2), 
                                 ", |LFC| > ", format(lfc_thresh, nsmall = 2)),
                   treeheight_row = 0,
                   treeheight_col = 0,
                   color = myColor,
                   cluster_rows = FALSE,
                   annotation_row = meta,
                   annotation_colors = ann_colors,
                   display_numbers = round(deg.df, digits = 0),
                   fontsize_number = 12,
                   number_color = "black")
```

## Heatmap Both Sexes
### Old vs Young
```{r old_vs_young_degs, eval=FALSE}
# set thresh
adjpval_thresh <- 0.01
lfc_thresh <- 0

# get file paths
files <- list.files(paste0(out, "DEGs_MAST_pct0.4/DEG_tables"))
files <- files[grepl(".+O_vs_.+Y_DEGs.tsv", files)]

# get comparisons
comparisons <- gsub("_DEGs.tsv", "", files)
all_comparisons <- c(paste0(comparisons, "_up"),
                     paste0(comparisons, "_down"))

# get cell types
cell_types <- names(cluster_colors)

# init deg.df
deg.df <- data.frame(
  matrix(0, nrow = length(all_comparisons), ncol = length(cell_types)),
  row.names = all_comparisons
)
colnames(deg.df) <- cell_types

# reorder rows
deg.df <- deg.df[c(8,4,6,2,7,3,5,1),]

# loop through files
for (i in files) {
  
  # read table
  df <- read.table(paste0(out, "DEGs_MAST_pct0.4/DEG_tables/", i),
                   sep = "\t", 
                   header = TRUE)
  
  # filter
  df <- df[!is.na(df$p_val_adj),]
  df <- df[df$p_val_adj < adjpval_thresh,]
  df <- df[df$avg_log2FC < -lfc_thresh | df$avg_log2FC > lfc_thresh,]
  
  # add direction column
  comparisonName <- gsub("_DEGs.tsv", "", i)
  direction <- df$avg_log2FC > 0 # already filtered on lfc so only matters it's positive/negative
  direction <- gsub(TRUE, paste0(comparisonName, "_up"), direction)
  direction <- gsub(FALSE, paste0(comparisonName, "_down"), direction)
  df$direction <- direction
  
  # reformat table
  df <- df %>% 
    dplyr::count(cluster, direction) %>%
    tidyr::spread(cluster, n)
  rownames(df) <- df$direction
  df$direction <- NULL
  
  # add to master table
  deg.df[rownames(df), colnames(df)] <- df
}

# reformat
deg.df[is.na(deg.df)] <- 0

# save
prefix <- paste0(out, "DEGs_MAST_pct0.4/DEG_comparison/",
                 "old_vs_young_DEG_both_sexes_comparison_FDRq_", 
                 format(adjpval_thresh, nsmall = 2),
                 "_LFC_", format(lfc_thresh, nsmall = 2))
write.table(df, 
            paste0(prefix, ".tsv"),
            sep = "\t", quote = FALSE, row.names = TRUE)

# set heatmap colors and names
paletteLength <- 100
myColor <- colorRampPalette(c("white","#f0eb9e","darkgreen"))(paletteLength)

meta <- data.frame(treatment = rep(c("Saline", "Saline", "E. coli", "E. coli"), 2),
                   time_point = c(rep("2 days", 4), rep("21 days", 4)),
                   direction = rep(c("Down-regulated","Up-regulated"), 4))
rownames(meta) <- rownames(deg.df)

ann_colors <- list(treatment = c(`E. coli` = "pink2", Saline = "chocolate4"),
                   time_point = c(`2 days` = "cyan", 
                                  `21 days` = "purple"),
                   direction = c(`Down-regulated` = "#0271f0",
                                 `Up-regulated` = "tomato"))

# save
pdf(paste0(prefix, ".pdf"), width = 10, height = 6)

# plot
pheatmap::pheatmap(deg.df,
                   main = paste0("Old vs Young, FDRq < ", format(adjpval_thresh, nsmall = 2), 
                                 ", |LFC| > ", format(lfc_thresh, nsmall = 2)),
                   treeheight_row = 0,
                   treeheight_col = 0,
                   color = myColor,
                   cluster_rows = FALSE,
                   annotation_row = meta,
                   annotation_colors = ann_colors,
                   display_numbers = round(deg.df, digits = 0),
                   fontsize_number = 12,
                   number_color = "black")
```

### 21 vs 2 days
```{r old_vs_young_degs, eval=FALSE}
# set thresh
adjpval_thresh <- 0.01
lfc_thresh <- 0.2

# get file paths
files <- list.files(paste0(out, "DEGs_MAST_pct0.4/DEG_tables"))
files <- files[grepl(".+21D.+[OY]_vs_.+2D.+[OY]_DEGs.tsv", files)]

# get comparisons
comparisons <- gsub("_DEGs.tsv", "", files)
all_comparisons <- c(paste0(comparisons, "_up"),
                     paste0(comparisons, "_down"))

# get cell types
cell_types <- names(cluster_colors)

# init deg.df
deg.df <- data.frame(
  matrix(0, nrow = length(all_comparisons), ncol = length(cell_types)),
  row.names = all_comparisons
)
colnames(deg.df) <- cell_types

# reorder rows
deg.df <- deg.df[c(8,4,6,2,7,3,5,1),]

# loop through files
for (i in files) {
  
  # read table
  df <- read.table(paste0(out, "DEGs_MAST_pct0.4/DEG_tables/", i),
                   sep = "\t", 
                   header = TRUE)
  
  # filter
  df <- df[!is.na(df$padj),]
  df <- df[df$padj < adjpval_thresh,]
  df <- df[df$log2FoldChange < -lfc_thresh | df$log2FoldChange > lfc_thresh,]
  
  # add direction column
  comparisonName <- gsub("_DEGs.tsv", "", i)
  direction <- df$log2FoldChange > 0 # already filtered on lfc so only matters it's positive/negative
  direction <- gsub(TRUE, paste0(comparisonName, "_up"), direction)
  direction <- gsub(FALSE, paste0(comparisonName, "_down"), direction)
  df$direction <- direction
  
  # reformat table
  df <- df %>% 
    dplyr::count(cell_type, direction) %>%
    tidyr::spread(cell_type, n)
  rownames(df) <- df$direction
  df$direction <- NULL
  
  # add to master table
  deg.df[rownames(df), colnames(df)] <- df
}

# reformat
deg.df[is.na(deg.df)] <- 0

# save
prefix <- paste0(out, "DEGs_MAST_pct0.4/DEG_comparison/",
                 "21_vs_2_days_DEG_both_sexes_comparison_FDRq_", 
                 format(adjpval_thresh, nsmall = 2),
                 "_LFC_", format(lfc_thresh, nsmall = 2))
write.table(df, 
            paste0(prefix, ".tsv"),
            sep = "\t", quote = FALSE, row.names = TRUE)

# set heatmap colors and names
paletteLength <- 100
myColor <- colorRampPalette(c("white","#f0eb9e","darkgreen"))(paletteLength)

meta <- data.frame(treatment = rep(c("Saline", "Saline", "E. coli", "E. coli"), 2),
                   age = c(rep("Young", 4), rep("Old", 4)),
                   direction = rep(c("Down-regulated","Up-regulated"), 4))
rownames(meta) <- rownames(deg.df)

ann_colors <- list(treatment = c(`E. coli` = "pink2", Saline = "chocolate4"),
                   sex = c(Female = "gray",
                           Male = "orange"),
                   age = c(Young = "gold",
                           Old = "chartreuse2"),
                   direction = c(`Down-regulated` = "#0271f0",
                                 `Up-regulated` = "tomato"))

# save
pdf(paste0(prefix, ".pdf"), width = 10, height = 6)

# plot
pheatmap::pheatmap(deg.df,
                   main = paste0("21 vs 2 Days, FDRq < ", format(adjpval_thresh, nsmall = 2), 
                                 ", |LFC| > ", format(lfc_thresh, nsmall = 2)),
                   treeheight_row = 0,
                   treeheight_col = 0,
                   color = myColor,
                   cluster_rows = FALSE,
                   annotation_row = meta,
                   annotation_colors = ann_colors,
                   display_numbers = round(deg.df, digits = 0),
                   fontsize_number = 12,
                   number_color = "black")
```


### Ecoli vs saline
```{r both_sexes_ecoli_vs_saline_degs, eval=FALSE}
# set thresh
adjpval_thresh <- 0.01
lfc_thresh <- 0.2

# get file paths
files <- list.files(paste0(out, "DEGs_MAST_pct0.4/DEG_tables"))
files <- files[grepl("E.+[OY]_vs_S.+[OY]_DEGs.tsv", files)]

# get comparisons
comparisons <- gsub("_DEGs.tsv", "", files)
all_comparisons <- c(paste0(comparisons, "_up"),
                     paste0(comparisons, "_down"))

# get cell types
cell_types <- names(cluster_colors)

# init deg.df
deg.df <- data.frame(
  matrix(0, nrow = length(all_comparisons), ncol = length(cell_types)),
  row.names = all_comparisons
)
colnames(deg.df) <- cell_types

# reorder rows
deg.df <- deg.df[c(8,4,7,3,6,2,5,1),]

# loop through files
for (i in files) {
  
  # read table
  df <- read.table(paste0(out, "DEGs_MAST_pct0.4/DEG_tables/", i),
                   sep = "\t", 
                   header = TRUE)
  
  # filter
  df <- df[!is.na(df$p_val_adj),]
  df <- df[df$p_val_adj < adjpval_thresh,]
  df <- df[df$avg_log2FC < -lfc_thresh | df$avg_log2FC > lfc_thresh,]
  
  # add direction column
  comparisonName <- gsub("_DEGs.tsv", "", i)
  direction <- df$avg_log2FC > 0 # already filtered on lfc so only matters it's positive/negative
  direction <- gsub(TRUE, paste0(comparisonName, "_up"), direction)
  direction <- gsub(FALSE, paste0(comparisonName, "_down"), direction)
  df$direction <- direction
  
  # reformat table
  df <- df %>% 
    dplyr::count(cluster, direction) %>%
    tidyr::spread(cluster, n)
  rownames(df) <- df$direction
  df$direction <- NULL
  
  # add to master table
  deg.df[rownames(df), colnames(df)] <- df
}

# reformat
deg.df[is.na(deg.df)] <- 0

# save
prefix <- paste0(out, "DEGs_MAST_pct0.4/DEG_comparison/",
                 "ecoli_vs_saline_DEG_both_sexes_comparison_FDRq_", 
                 format(adjpval_thresh, nsmall = 2),
                 "_LFC_", format(lfc_thresh, nsmall = 2))
write.table(df, 
            paste0(prefix, ".tsv"),
            sep = "\t", quote = FALSE, row.names = TRUE)

# set heatmap colors and names
paletteLength <- 100
myColor <- colorRampPalette(c("white","#f0eb9e","darkgreen"))(paletteLength)

meta <- data.frame(time_point =  c(rep("2 days", 4), rep("21 days", 4)),
                   age = c(rep("Young", 2), rep("Old", 2),
                           rep("Young", 2), rep("Old", 2)),
                   direction = rep(c("Down-regulated","Up-regulated"), 4))
rownames(meta) <- rownames(deg.df)

ann_colors <- list(time_point = c(`2 days` = "cyan", 
                                  `21 days` = "purple"),
                   age = c(Young = "gold",
                           Old = "chartreuse2"),
                   direction = c(`Down-regulated` = "#0271f0",
                                 `Up-regulated` = "tomato"))

# save
pdf(paste0(prefix, ".pdf"), width = 10, height = 6)

# plot
pheatmap::pheatmap(deg.df,
                   main = paste0("E. coli vs Saline, FDRq < ", format(adjpval_thresh, nsmall = 2), 
                                 ", |LFC| > ", format(lfc_thresh, nsmall = 2)),
                   treeheight_row = 0,
                   treeheight_col = 0,
                   color = myColor,
                   cluster_rows = FALSE,
                   annotation_row = meta,
                   annotation_colors = ann_colors,
                   display_numbers = round(deg.df, digits = 0),
                   fontsize_number = 12,
                   number_color = "black")
```

## Box and wisker
### S.21D.O vs Y
```{r}
# set thresh
adjpval_thresh <- 0.05
lfc_thresh <- 0

# file paths
files <- c("S.21D.O_vs_S.21D.Y_DEGs.tsv",
           "S.21D.O.M_vs_S.21D.Y.M_DEGs.tsv",
           "S.21D.O.F_vs_S.21D.Y.F_DEGs.tsv")

# get df with all DEGs
master_df <- data.frame()
for (file in files) {
  
  # read table
  df <- read.table(paste0(out, "DEGs_MAST_pct0.4/DEG_tables/", file),
                   sep = "\t", 
                   header = TRUE)
  
  # get rid of percent columns
  df <- df[,-c(6,7)]
  
  # add comparison column
  df$comparison <- gsub("_DEGs.tsv", "", file)
  
  # add to master
  master_df <- rbind(master_df, df)
}

# filter by significance
master_df <- master_df[master_df$p_val_adj < adjpval_thresh,]

# add direction column
direction <- master_df$avg_log2FC > 0
direction <- gsub(TRUE, "up", direction)
direction <- gsub(FALSE, "down", direction)
master_df$direction <- direction

# add levels
master_df$cluster <- factor(master_df$cluster, levels = rev(c("Microglia",
                                                          "Astrocytes",
                                                          "Endothelial",
                                                          "Oligodendrocytes",
                                                          "OPCs",
                                                          "Neurons 1",
                                                          "BAM",
                                                          "Neurons 2",
                                                          "Fibroblasts")))

# reformat for plot
plot_df <- master_df %>%
  # count genes per cluster, direction, and comparison
  group_by(cluster, comparison, direction) %>%
  summarise(n = n(), .groups = "drop") %>%
  # ensure both up and down exist for every cluster within each comparison
  complete(cluster, comparison, direction = c("up", "down"), fill = list(n = 0)) %>%
  # create signed values for plotting
  mutate(n_plot = ifelse(direction == "down", -n, n))

# save
pdf("../../../figures/saline_DEGs.pdf", width = 8, height = 6)

# plot
ggplot(plot_df, aes(x = cluster, y = n_plot, fill = direction)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = n), 
            vjust = 0.5, 
            hjust = ifelse(plot_df$n_plot > 0, -0.2, 1.2),
            size = 2.) +
  scale_fill_manual(values = c("down" = "blue", "up" = "red")) +
  coord_flip() + 
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  labs(title = paste0("Number of Differentially Expressed Genes\nFDRq < ",
                      format(adjpval_thresh, nsmall = 2)),
       y = "", x = "") +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.key.size = unit(0.3, "cm"),
    legend.text = element_text(size = 7),
    legend.spacing.x = unit(0.1, "cm"),
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    plot.title = element_text(size = 8),
    plot.margin = margin(.1, .35, .01, 0, "cm")
  ) +
  scale_y_continuous(limits = c(min(plot_df$n_plot) * 1.3, 
                                max(plot_df$n_plot) * 1.3)) +
  facet_wrap(~comparison)
```


## Volcano
```{r volcano, message=FALSE, eval=FALSE}
# set thresh
adjpval_thresh <- 0.01
lfc_thresh <- 0.2

# get file paths
files <- list.files(paste0(out, "DEGs_MAST_pct0.4/DEG_tables"), full.names = TRUE)
files <- files[grepl(".+_DEGs.tsv", files)]

for (file in files) {
  
  # read DEG file
  treatment_vs_control <- read.delim(file)
  
  # filter NA values
  treatment_vs_control <- treatment_vs_control[!is.na(treatment_vs_control$p_val_adj),]
  
  # extract title for plot later
  p.comparison <- str_match(file, "DEG_tables/(.+)_DEGs.tsv")[,2]
  
  # assign colors
  color_values <- vector()
  max <- nrow(treatment_vs_control)
  for(row in 1:max){
    if (treatment_vs_control$p_val_adj[row] < adjpval_thresh){
      if (treatment_vs_control$avg_log2FC [row] > lfc_thresh){
        color_values <- c(color_values, 1) # 1 when logFC > thresh and FDRq < thresh
      } else if (treatment_vs_control$avg_log2FC[row] < -lfc_thresh){
        color_values <- c(color_values, 2) # 2 when logFC < thresh and FDRq < thresh
      } else {
        color_values <- c(color_values, 3) # 2 when FDRq < thresh and LFC not met
      }
    }
    else{
      color_values <- c(color_values, 3) # 3 when FDRq >= thresh
    }
  }
  treatment_vs_control$color_adjpval_thresh <- factor(color_values)
  
  # loop through clusters
  for (j in unique(treatment_vs_control$cluster)) {
    
    # subset cluster
    data <- subset(treatment_vs_control, cluster == j)
    
    # plot only if there are DEGs with padj < thresh
    num <- subset(data, p_val_adj < adjpval_thresh)
    num <- nrow(num)
    if(num != 0) {
        
      # subset genes to label
      up <- data[data$color_adjpval_thresh == 1,]
      up.sig <- up[order(up$p_val_adj),][1:15,]
      up.lfc <- up[order(up$avg_log2FC, decreasing = TRUE),][1:15,]
      up30 <- rbind(up.sig,up.lfc)
      up30 <- unique(up30)
      up30 <- up30[rowSums(is.na(up30)) != ncol(up30), ]
      down <- data[data$color_adjpval_thresh == 2,]
      down.sig <- down[order(down$p_val_adj),][1:15,]
      down.lfc <- down[order(down$avg_log2FC, decreasing = FALSE),][1:15,]
      down30 <- rbind(down.sig,down.lfc)
      down30 <- unique(down30)
      down30 <- down30[rowSums(is.na(down30)) != ncol(down30), ]
      
      # set manual colors
      if (!1 %in% unique(data$color_adjpval_thresh)) {
        my_colors <- c("blue","gray")
      } else if (!2 %in% unique(data$color_adjpval_thresh)) {
        my_colors <- c("red","gray")
      } else if (!1 %in% unique(data$color_adjpval_thresh) && !2 %in% unique(data$color_adjpval_thresh)) {
        my_colors <- c("gray")
      } else {
        my_colors <- c("red","blue","gray")
      }
      
      # set significance threshold
      hadjpval <- (-log10(max(
        data$p_val[data$p_val_adj < adjpval_thresh], 
        na.rm=TRUE)))

      # plot
      p <-
        ggplot(data = data, 
               aes(x = avg_log2FC,     # x-axis is logFC
                   y = -log10(p_val),  # y-axis will be -log10 of P.Value
                   color = color_adjpval_thresh)) +  # color is based on factored color column
        geom_point(alpha = 0.8, size = 2) +  # create scatterplot, alpha makes points transparent
        theme_bw() +  # set color theme
        theme(legend.position = "none") +  # no legend
        scale_color_manual(values = my_colors) +
        labs(
          title = "",
          x = expression(log[2](FC)),
          y = expression(-log[10] ~ "(" ~ italic("p") ~ "-value)")
        ) +
        theme(axis.title.x = element_text(size = 10),
              axis.text.x = element_text(size = 10)) +
        theme(axis.title.y = element_text(size = 10),
              axis.text.y = element_text(size = 10)) +
        geom_hline(yintercept = hadjpval, colour = "#000000", linetype = "dashed") +
        geom_vline(xintercept = lfc_thresh, colour = "#000000", linetype = "dashed") +
        geom_vline(xintercept = -lfc_thresh, colour = "#000000", linetype = "dashed") +
        ggtitle(paste0(j, "\n", p.comparison, ", q < ", format(adjpval_thresh, nsmall = 2), 
                       ", |LFC| > ", format(lfc_thresh, nsmall = 2))) +
        geom_text_repel(data = up30,
                        aes(x = avg_log2FC, y= -log10(p_val), label = gene), 
                        color = "maroon", 
                        fontface="italic",
                        max.overlaps = getOption("ggrepel.max.overlaps", default = 30)
                        ) +
        geom_text_repel(data = down30,
                        aes(x = avg_log2FC, y= -log10(p_val), label = gene), 
                        color = "navyblue", 
                        fontface="italic",
                        max.overlaps = getOption("ggrepel.max.overlaps", default = 30)
                        )
      p
      
      # create dir if doesn't exist
      prefix <- paste0(out, "DEGs_MAST_pct0.4/volcano/", p.comparison)
      if (!dir.exists(prefix)) {
        dir.create(prefix, recursive = TRUE)
      }
      
      # save
      j_reformat <- gsub(" / ", "_", j)
      j_reformat <- gsub(" ", "_", j_reformat)
      plot_out <- paste0(prefix, "/", p.comparison, "_", tolower(j_reformat),
                    "_FDRq_", format(adjpval_thresh, nsmall = 2), "_LFC_", 
                    format(lfc_thresh, nsmall = 2), "_volcano.pdf")
      pdf(plot_out, height = 8, width = 8)
      print(p)
      dev.off()
      
      print(paste("comparison =", p.comparison, ", cell type =", j))
      
    }
  } # end loop through clusters
} # end loop through variables
```

## Metascape input
```{r metascape_sex_specific}
# set thresh
adjpval_thresh <- 0.01
lfc_thresh <- 0.2

# get file paths
files <- list.files(paste0(out, "DEGs_MAST_pct0.4/DEG_tables"), full.names = TRUE)
files <- files[grepl(".+_DEGs.tsv", files)]

# loop through files
for (i in 1:length(files)) {
  
  # read table
  data <- read.table(files[i], header = TRUE, sep = "\t")
  
  # filter
  data <- data[!is.na(data$p_val_adj),]
  data <- data[data$p_val_adj < adjpval_thresh,]
  data <- data[data$avg_log2FC < -lfc_thresh | data$avg_log2FC > lfc_thresh,]
  
  # separate into up and down-regulated
  up.df <- data[data$avg_log2FC > 0,]
  down.df <- data[data$avg_log2FC < 0,]
  
  cell_types <- unique(data$cluster)
  
  # create dir if doesn't exist
  p.comparison <- str_match(files[i], "DEG_tables/(.+)_DEGs.tsv")[,2]
  prefix <- paste0(out, "DEGs_MAST_pct0.4/metascape_input/", p.comparison)
  if (!dir.exists(prefix)) {
    dir.create(prefix, recursive = TRUE)
  }
  
  # loop through cell_type
  for (j in cell_types) {
    
    # subset based on cell_type
    print(paste0(p.comparison, ": ", j))
    up <- up.df[up.df$cluster == j,]
    up <- up$gene
    down <- down.df[down.df$cluster == j,]
    down <- down$gene
    
    # file name
    prefix <- paste0(out, "DEGs_MAST_pct0.4/metascape_input/", p.comparison,
                     "/", p.comparison, "_", gsub(" ", "_", tolower(j)), 
                     "_FDRq_", format(adjpval_thresh, nsmall = 2), 
                     "_LFC_", format(lfc_thresh, nsmall = 2))
     
    # save
    if(length(up) > 10) {
      write.table(x = up,
                  file = paste0(prefix, "_up.txt"),
                  quote = FALSE,
                  row.names = FALSE,
                  col.names = FALSE)      
    }
    if (length(down) > 10) {
       write.table(x = down,
                file = paste0(prefix, "_down.txt"),
                quote = FALSE,
                row.names = FALSE,
                col.names = FALSE)   
    }
  } # end j for loop
} # end i for loop
```

# Shiny app
```{r shiny_app, eval=FALSE}
# create new object
library(ShinyCell)
shiny.obj <- mouse.annotated
VariableFeatures(shiny.obj) <- shiny.obj@assays$SCT@var.features

# set default params
DefaultAssay(shiny.obj) <- "RNA"

# create config
colNames <- colnames(shiny.obj@meta.data)
colNames <- colNames[c(28,29,2:24)]
sc.config <- createConfig(obj = shiny.obj,
                          meta.to.include = colNames)

# change wd
setwd(out)

# output shiny app folder
makeShinyApp(obj = shiny.obj, 
             scConf = sc.config, 
             gene.mapping = TRUE,
             shiny.title = "E. coli Mice scRNAseq")

# manual config edits
setwd(out)
sc1conf <- readRDS("shinyApp/sc1conf.rds")
cluster_hex <- "#F0E442|#FDB863|#E08214|#B35806|#542788|#998EC3|#0868AC|#2B8CBE|#009E73|#7FCDBB"
sc1conf[2,4] <- cluster_hex
saveRDS(sc1conf, "shinyApp/sc1conf.rds")

# cleanup
remove(sc.config,sc1conf,shiny.obj)
```
